### 2015-08-13 article/main.tex
include article and talk on the public repo

% compile using latex (xelatex seems to crop the figure strangley)
\documentclass{article} % For LaTeX2e
\usepackage[dvips]{graphicx}
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[outdir=./]{epstopdf}

\usepackage{graphics,graphicx}
\usepackage{amsmath,amssymb,amsfonts,comment}
\definecolor{LinkColor}{rgb}{0,0,0}

\newcommand{\p}{\text{P}}
\newcommand{\word}[1]{\texttt{#1}}
\urldef{\googlegroup}\url{https://groups.google.com/forum/#\!forum/word2vec-toolkit}
\urldef{\blogpost}\url{https://blog.lateral.io/2015/06/the-unknown-perils-of-mining-wikipedia/ }

\title{Corpus Experiments for Word Embeddings}

 \author{
 	Benjamin Wilson\\
	Lateral GmbH\\
	\texttt{benjamin@lateral.io}
	\And
	Adriaan Schakel\\
	\texttt{adriaan.schakel@gmail.com}
 }

\date{\today}
\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\graphicspath{{../outputs/}}
\maketitle

\begin{abstract}
	We introduce an experimental approach to studying the properties of word embeddings.
	Controlled experiments, acheived by modification of the training corpus, permit the demonstration of direct, causal relationships between word properties and word vector direction and length.
	This approach is demonstrated in the case of the word2vec CBOW model with experiments that independently vary word frequency and word co-occurrence noise.
\end{abstract}

\section{Introduction}
Word embeddings are typically trained with samples from the word co-occurrence distributions.
Recall that the co-occurrence distribution of a word $w$ gives the probability $\p(w'|w)$ that a word $w'$ occurs nearby, given that $w$ occurred.
Samples of the co-occurrence distributions are obtained by scanning a short window over the text.
The non-central (or ``context'') words are treated as samples of the co-occurrence distribution the middle (or ``current'') word. 
In this way, the word vectors of the word embedding are determined by the co-occurrence distributions and frequencies of the words in the vocabulary.

In this short note, we perform experiments that illustrate the effect of these two determining pieces of information (co-occurrence distribution and word frequency) as they are varied independently of one another.
This is acheived through the introduction of new ``words'' into the training corpus with varying frequencies and varying levels of noise in their co-occurrence distributions.
The frequency and co-occurrence distributions of these introduced words are modelled on existing words in the corpus.

We illustrate our approach in case of the word2vec CBOW model.
However, these experiments (and other besides) could might equally well be performed for word embedding methods such as word2vec skip-gram, GloVe and SENNA.  \cite{}
In the case of word2vec, we show that word vector length depends directly and linearly with word frequency.
We show furthermore that word vector length depends directly and linearly with the level of noise in the co-occurrence distribution of the word.
In both cases, the coefficient of linearity depends upon the word.

In the word2vec similarity and word relationship tasks, normalised word vectors are used.
Word vector length is thus disregarded.
In the word2vec forum\footnote{\googlegroup}, it has been observed that normalisation improves performance on theses tasks since word vector length is related to word frequency.
It remained unclear, however, whether word frequency was \textit{directly} related to word vector lengthor rather via correlated factors such as word significance.
Our experimental approach demonstrates that a direct relationship indeed exists.
It is shown moreover that the normalisation of the word vectors also prevents word co-occurrence noise from affecting the performance of the word similarity and word relationship tasks.

\section{Model and corpus}
Our training data is built from the Wikipedia XML datadump from October 2013.
In order to remove from the training data the bulk of robot-generated pages on Wikipedia, only pages with at least 20 monthly page views are retained\footnote{For further justification of this rationale, and to obtain the dataset, see \blogpost}.
Stubs and disambiguation pages are also removed, leaving 463 thousand pages with a total of 482 million words.

This base corpus is then modified as described in section \ref{experimental_design}.
Punctuation and numbers were removed and the corpus lower-cased prior to the modifications.
For recognisability, the tokens introduced into the corpus during the modification are uppercased.

For simplicity, only the CBOW method with a single set of hyperparameters is considered.
Specifically, a $100$-dimensional CBOW model is trained using negative sampling with 5 negative samples, a window size of $10$, a minimum word occurrence of $200$, and $10$ passes through the corpus.
Subsampling was not used so that influence of word frequency could be more clearly discerned.
Identical experimental results were obtained using hierarchical softmax, but these are omitted for succinctness.

The most recent revision of word2vec was used\footnote{SVN revision 42, see \url{http://word2vec.googlecode.com/svn/trunk/}}.
The source code for performing the experiments is made available on GitHub\footnote{\url{https://github.com/benjaminwilson/word2vec-norm-experiments}}.

\section{Experimental design and corpus modifications}\label{experimental_design}
Two controlled experiments are performed:
\begin{itemize}
	\item \emph{Word frequency variation} compare the lengths of the word vectors for words with varying frequencies, but with the same co-occurrence distribution;
	\item \emph{Co-occurrence noise variation} compare the lengths of the word vectors for words with of the same frequency, but with varying levels of noise in their co-occurrence distributions.
\end{itemize}

To perform these experiments, tokens are introduced into the corpus described in subsections \ref{WFVE} and \ref{CNVE}.
These tokens are introduced via a replacement procedure.
To illustrate this procedure, consider a word, e.g. \word{cat}.
For each occurrence of this word, a sample $i$, $1 \leqslant i \leqslant n$ is drawn from a truncated geometric distribution, and that occurrence of \word{cat} is replaced with \word{CAT\_i}.
Thus the word \word{cat} is replaced throughout the corpus by a family of tokens with varying frequencies but approximately the same co-occurrence distribution as \word{cat}.

The geometric distribution is truncated in order to limit the number of tokens introduced into the corpus.
For any ratio $0 < \lambda < 1$ and maximum value $n > 0$, the truncated geometric distribution is given by
$$ P_{\lambda, n} (i) = \frac{\lambda^{i-1} \cdot (1-\lambda)}{(1 - \lambda^n)}, \qquad 1 \leqslant i \leqslant n.$$ 
This is the distribution over $i = 1, 2, \dots n$ for which the probabilities decay exponentially base $\lambda$.
Of course, other distributions might equally have been chosen for the experiments.

\subsection{Word frequency variation experiment}\label{WFVE}
Firstly, a small set of words from the unmodified corpus is chosen uniformly at random from the vocabulary.
In order that the introduced tokens do not have too low a frequency, only words which occur at least 10 thousand times are chosen.
To this set of words we add the word \word{the}, in order to include a high-frequency stopword.
The replacement procedure is performed for each of these words, using a geometric rate of decay of $\lambda = 0.5$, and maximum value $n=20$.

Secondly, in order to have a meaningless word for comparison, an entirely new word \word{VOID} is interspersed uniformly at random throughout the corpus so that its global frequency is $0.005$.
The replacement procedure is then performed for this word with the same values for $\lambda$ and $n$.

Figure \ref{fig:word-frequency-experiment-text} shows the effect of these modifications on sample text, in the case where only the word \word{cat} is chosen.

\subsection{Co-occurrence noise variation experiment}\label{CNVE}
Here, and throughout, the noise distribution is taken to be the global word frequency distribution.
Thus noise can be added to the cooccurrence distribution of a word by inspersing occurrences of that word uniformly at randomly throughout the corpus.
A small set of words is chosen from the unmodified corpus in the same manner as subsection \ref{WFVE}.
In order that the two experiments can be carried out simulataneously, the two sets of words are chosen to be disjoint.

For each of the chosen words, the replacement procedure is performed using a geometric rate of decay of $\lambda = 5/6$, and truncating at $n=8$.
Subsequently, for every replacement token (e.g. \word{CAT\_i}) random occurrences of this token are interspersed uniformly at random throughout the corpus, such that the frequency of the replacement token is restored to that of the original token \word{cat}.
For example, if the original token \word{cat} occurred $1000$ times, then after the replacement procedure, \word{CAT\_2} occurs approximately $694$ times, so a further approximately $306 (=1000 - 694)$ random occurrences of \word{CAT\_2} are interspersed throughout the corpus.
Thus token \word{cat} is removed from the corpus, and a family of tokens \word{CAT\_i}, $1 \leqslant i \leqslant 8$ are introduced.
These tokens have the same frequency as \word{cat}, but their cooccurrence distributions, while based on that of \word{cat}, have an increasing amount of noise.
To be precise, the proportion of the occurrences of \word{CAT\_i} that arose by replacing \word{cat} is $P_{5/6}(i)$, and the proportion arising from its random intersperal throughout the corpus $(1 - P_{5/6}(i))$.

Figure \ref{fig:cooccurrence-noise-experiment-text} illustrates the effect of this modification, independently of the modifications of subsection \ref{WFVE}, in the case where the only word chosen is \word{cat}.
Note that for the experiments, \emph{both} modification procedures are performed, but with different sets of words.

\begin{figure*}
	%\input{../outputs/word-frequency-experiment-counts.tex}
\begin{center}
\begin{tabular}{c | c}
word & \# occurrences \\
\hline
\word{developments} & 14347 \\
\word{specific} & 77919 \\
\word{across} & 101617 \\
\word{path} & 24898 \\
\word{file} & 31184 \\
\word{from} & 2611891 \\
\word{admiral} & 13122 \\
\word{eyes} & 27621 \\
\end{tabular}
\end{center}
\label{fig:word-frequency-counts}
\caption{Occurrence counts for words chosen for the word frequency experiment. }
\end{figure*}

\begin{figure*}
	%\input{../outputs/word-frequency-experiment-counts.tex}
\begin{center}
\begin{tabular}{c | c}
word & \# occurrences \\
\hline
\word{sun} & 49723 \\
\word{jets} & 10520 \\
\word{holding} & 31744 \\
\word{incorporated} & 24964 \\
\word{initiative} & 15635 \\
\word{jury} & 14575 \\
\word{afternoon} & 10537 \\
\word{plants} & 42808 \\
\end{tabular}
\end{center}
\label{fig:cooccurrence-noise-counts}
\caption{Occurrence counts for words chosen for the cooccurrence noise experiment. }
\end{figure*}

% text after modifications for word frequency variation experiment for the word cat
\begin{figure}[t]
\texttt{
a pedigreed \textcolor{red}{CAT\_2} is one whose ancestry is recorded by a \textcolor{red}{CAT\_5} fancier organization
a purebred \textcolor{red}{CAT\_7} is one whose ancestry contains only individuals of the same \textcolor{red}{VOID\_4} breed
the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_1} is not owned by \textcolor{red}{VOID\_7} any one individual
the domestic \textcolor{red}{CAT\_1} was first classified as felis catus
the domesticated \textcolor{red}{CAT\_1} and its closest wild ancestor are both diploid organisms that possess  chromosomes and roughly  genes
the \textcolor{red}{CAT\_5} skull is unusual among mammals in having very large eye sockets and a powerful and specialized jaw
when it overpowers its prey a \textcolor{red}{CAT\_3} delivers a lethal neck bite with its two long canine teeth
another unusual feature is that the \textcolor{red}{CAT\_1} cannot produce taurine
this is partly the result of \textcolor{red}{CAT\_1} eyes having a tapetum lucidum which reflects any light that passes through the retina back into the eye
although wildcats are solitary the social behavior of domestic cats is much more variable and ranges from widely dispersed individuals to feral \textcolor{red}{CAT\_1} colonies that form around a food source based on groups of cooperating females
within such groups one \textcolor{red}{CAT\_1} is usually dominant over the others
each \textcolor{red}{CAT\_1} in a colony holds a distinct territory with sexually active males having the largest territories which are about  times larger than those of female cats and may overlap with several females territories
these territories are marked by urine spraying by rubbing objects at head height with secretions from facial glands and by defecation
between these territories are neutral areas where cats watch and greet one another without territorial conflicts
outside these neutral areas territory holders usually chase away stranger \textcolor{red}{VOID\_2} cats at first by staring hissing and growling and if that does not work by short but noisy and violent attacks
despite some cats cohabiting in colonies they do not have a social survival strategy or a pack mentality and always hunt alone
the domestic dog canis lupus familiaris or canis familiaris is a domesticated canid which has been selectively bred for millennia for various behaviors sensory capabilities and physical attributes
although initially thought to have originated as a manmade variant of an extant canid species variously supposed as being the dhole golden jackal or gray wolf extensive genetic studies undertaken during the s indicate that dogs diverged from a now-extinct canid in eurasia  years ago being the oldest domesticated animals their long association with people has allowed dogs to be uniquely attuned to human behavior as well as thrive on a starch-rich diet which would be inadequate for other canid species
dogs perform many roles for people such as hunting herding pulling loads protection assisting police and military companionship and more recently aiding handicapped individuals in  carl linnaeus listed among the types of quadrupeds familiar to him the latin word for \textcolor{red}{VOID\_3} dog canis among the species within this genus linnaeus listed the red fox as canis vulpes wolves canis lupus and the domestic dog canis canis in later editions linnaeus dropped canis canis and greatly expanded his list of the canis genus of quadrupeds and by  included alongside the foxes wolves and jackals and many more terms that are now listed as synonyms for domestic dog including aegyptius hairless dog aquaticus water dog and mustelinus literally badger dog among these were two that later experts have been widely used for domestic dogs as a species: canis domesticus and most predominantly canis familiaris the common or familiar dog
}
\caption{Text modified for the word frequency experiment, where the word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
\label{fig:word-frequency-experiment-text}
\end{figure}

% original text after modifications for the noise experiment with n=3 lambda 5/ 6 for cat
\begin{figure}[t]
	\texttt{
a pedigreed \textcolor{red}{CAT\_1} is \textcolor{red}{CAT\_1} one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization
a purebred \textcolor{red}{CAT\_3} is one whose ancestry contains only individuals of the same breed
the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_3} is not owned by any one individual
\textcolor{red}{CAT\_1} the domestic \textcolor{red}{CAT\_3} was first classified as felis catus
the domesticated \textcolor{red}{CAT\_1} and its closest wild ancestor are both diploid organisms \textcolor{red}{CAT\_3} that possess  chromosomes and roughly  genes
the \textcolor{red}{CAT\_2} skull is unusual among mammals in having very large eye sockets and a powerful and specialized jaw
when it overpowers its \textcolor{red}{CAT\_2} prey a \textcolor{red}{CAT\_1} delivers a lethal neck bite with \textcolor{red}{CAT\_3} its two \textcolor{red}{CAT\_2} long canine teeth
another unusual feature \textcolor{red}{CAT\_1} is that the \textcolor{red}{CAT\_2} cannot produce taurine
this is partly the result of \textcolor{red}{CAT\_3} eyes having a tapetum lucidum which reflects any light that passes through the retina back into the eye
although wildcats are solitary the social behavior of domestic cats is much more variable and \textcolor{red}{CAT\_1} ranges from widely dispersed individuals to feral \textcolor{red}{CAT\_2} colonies that form around a food source based on groups of cooperating females
within such groups one \textcolor{red}{CAT\_2} is usually dominant over the others
each \textcolor{red}{CAT\_1} in \textcolor{red}{CAT\_3} a colony holds a distinct \textcolor{red}{CAT\_3} territory with sexually active males having the largest territories which are about  times larger than those of female cats and may overlap with several females territories
these territories \textcolor{red}{CAT\_3} are marked by urine spraying by rubbing objects \textcolor{red}{CAT\_1} at head height with secretions from facial glands and by defecation
between these territories are neutral areas where cats \textcolor{red}{CAT\_1} watch and greet \textcolor{red}{CAT\_2} one another without territorial conflicts
\textcolor{red}{CAT\_1} outside these neutral areas territory holders usually chase away stranger cats at first by staring hissing and growling and if that does not work by short but \textcolor{red}{CAT\_2} noisy and violent attacks
despite some cats cohabiting in colonies they do \textcolor{red}{CAT\_3} not have a social survival strategy \textcolor{red}{CAT\_2} or a pack mentality and always hunt alone
the domestic dog canis lupus familiaris or canis familiaris is a domesticated canid which has been selectively bred for millennia for various behaviors sensory capabilities and physical attributes
although initially thought to have originated as a manmade variant of an extant canid species variously supposed as being the \textcolor{red}{CAT\_1} dhole golden jackal or gray wolf \textcolor{red}{CAT\_3} extensive genetic studies undertaken during the s indicate that dogs \textcolor{red}{CAT\_1} diverged from a now-extinct canid in eurasia  years ago being the oldest domesticated animals their long association with people has \textcolor{red}{CAT\_3} allowed dogs to be \textcolor{red}{CAT\_1} uniquely attuned to human behavior as well as thrive on \textcolor{red}{CAT\_2} a starch-rich diet which would be inadequate \textcolor{red}{CAT\_2} for other \textcolor{red}{CAT\_1} canid species
dogs perform many roles for people such \textcolor{red}{CAT\_3} as hunting herding pulling loads protection assisting \textcolor{red}{CAT\_2} police and military companionship and more recently aiding handicapped \textcolor{red}{CAT\_2} individuals in  carl linnaeus listed among the types of quadrupeds familiar to him the latin word for \textcolor{red}{CAT\_1} dog canis among \textcolor{red}{CAT\_1} the species within this genus linnaeus listed the red fox as canis vulpes wolves canis lupus and the domestic dog canis canis in later editions linnaeus dropped canis canis and greatly expanded his list of the canis genus of quadrupeds and by  included alongside the foxes wolves and jackals and \textcolor{red}{CAT\_2} many more terms that are now listed as synonyms for domestic dog including aegyptius \textcolor{red}{CAT\_1} hairless dog aquaticus water dog and mustelinus literally badger dog among these were two that later experts have been widely used for domestic dogs as a species: canis domesticus and most predominantly \textcolor{red}{CAT\_2} canis familiaris the common or familiar dog
}
\caption{Text modified for the co-occurrence noise experiment, where the word \word{cat} was chosen, $\lambda = 5/6$ and $n=3$.}
\label{fig:cooccurrence-noise-experiment-text}
\end{figure}

\section{Results}

\subsection{Word frequency experiment}
Figure \ref{fig:frequency-norm-graph} illustrates the complex global relationship between vector length and word frequency.
Figure \ref{fig:word-frequency-experiment-graph}, on the other hand, shows this relationship for individual words.
Each line corresponds to a single word, and the points on each line indicate the frequency and vector length for the tokens derived from the corresponding word.
For example, the five points on the line corresponding to \word{admiral} are labelled, from right to left, by the tokens \word{ADMIRAL\_1}, \word{ADMIRAL\_2}, \dots, \word{ADMIRAL\_5}.
The number of points on the line is determined by the frequency of the original word.
For example, the frequency of the \word{admiral} can be halved at most $5$ times and remain about the minimum frequency cut-off of word2vec.

Because all the points on a line share a co-occurrence distribution (only the frequency varies), this figure demonstrates conclusively that length does indeed depend on frequency directly, and not only via some correlation between word frequency and semantic content.
Notice, moreover, that the relative positions of the norms of the word vectors associated to the test words is independent of the frequency band.
Therefore these lines trace lines along which a frequency-adjusted norm would be constant.

\begin{figure*}
\includegraphics[scale=0.5]{occurrence-histogram}
\caption{
	Word frequency histogram.
	Here and elsewhere, frequency bands are delimited by powers of $2$.
}
\label{fig:frequency-histogram}
\end{figure*}

\begin{figure*}
\includegraphics[scale=0.3]{frequency-norm-scatterplot}
\caption{ The global relationship between word frequency and vector length.  }
\label{fig:frequency-norm-graph}
\end{figure*}

\begin{figure*}
\includegraphics[scale=0.5]{word-frequency-experiment-heatmap}
\caption{
Heatmap of the dot product of the normalised vectors associated to the tokens introduced into the corpus in the co-occurrence noise experiment (four such words chosen at random).
The largely red blocks demonstrate that the direction of the word vector does not change when noise is added to the co-occurrence distribution.
}

\label{fig:cooccurrence-noise-heatmap}
\end{figure*}
\begin{sidewaysfigure*}
	\centering{\includegraphics[scale=0.6]{word-frequency-experiment-graph}}
\caption{
	Vector length vs. frequency for the words chosen for the word frequency experiment.
	For each word, tokens of varying frequency but with the co-occurrence distribution of that word were introduced into the corpus, as described in section \ref{WFVE}.
}
\label{fig:word-frequency-experiment-graph}
\end{sidewaysfigure*}

\subsection{Co-occurrence noise experiment}
Consider the effect on a word vector when increasing levels of noise added to the co-occurrence distribution of the word.
This is affected by randomly dispersing an increasing proporition $0 \leqslant p \leqslant 1$ of the pre-existing occurrences of that word throughout the corpus.
The word vector depends on $p$, and thus $p$ parameterises a path in the feature space.
This path begins (with $p=0$) at the original word vector, and ends (with $p=1$) at the vector associated with the meaningless vector \word{VOID}.

For each word (e.g. \word{jury}) in the co-occurrence noise experiment, the vectors associated with the tokens \word{JURY\_i} mark points along this word's path for increasing values of $p$.
Figure \ref{fig:cooccurrence-noise-heatmap} shows that these points at all co-linear.
Thus the path is a straight line.

Figure \ref{fig:cooccurrence-noise-graph} illustrates the relationship between the proportion $p$ of co-occurrence noise and vector length.
Remarkably, it is apparent that this relationship is linear for any individual word.
Thus the norm decreases at a constant velocity as $p$ travels from $0$ to $1$.

It is Figure \ref{fig:cooccurrence-noise-graph} that motivates the interpretation of vector length, within a sufficiently frequency band, as a measure of the absence of co-occurrence noise, that is, of the extent to which a word determines a context.

\begin{sidewaysfigure*}
\includegraphics[scale=0.6]{cooccurrence-noise-graph}
\caption{
	Vector length vs. proportion of co-occurrence noise for words chosen for the co-occurrence noise experiment.
	For each word, tokens of equal frequency but with increasing proporitions of co-occurrence noise were introduced into the corpus, as described in subsection \ref{CNVE}.
}
\label{fig:cooccurrence-noise-graph}
\end{sidewaysfigure*}

\begin{figure*}[t!]
\includegraphics[scale=0.5]{cooccurrence-noise-heatmap}
\caption{
Heatmap of the dot product of the normalised vectors associated to the tokens introduced into the corpus in the co-occurrence noise experiment (four such words chosen at random).
The largely red blocks demonstrate that the direction of the word vector does not change when noise is added to the co-occurrence distribution.
}
\label{fig:cooccurrence-noise-heatmap}
\end{figure*}

\section{Questions}
\begin{enumerate}
\item{ Why is the meaningless vector not at zero, and could the performance on the compositionality tasks be improved by accounting for this?  Notice that the meaningless vector is only updated as part of the context vector.  Would it be zero if we were performing skipgram? Does this have something to do with the lack of bias terms in the word2vec softmax?}
\item{Our arguments from the results subsection on the co-occurrence noise experiment indicate that the meaningless vector should indeed be taken as the centre of the space for word similarity tasks.  Check again whether this works experimentally.}
\item{What is the distribution of norms for the vectors relative to the meaningless vector?  Do I need to consider a different meaningless vector for each frequency band?  This is the measure that should represent meaning.}

\end{enumerate}

\footnotesize
\bibliographystyle{unsrt}
\bibliography{main}

\end{document}

### 2015-08-13 talk/main.tex
include article and talk on the public repo

% $Header: /Users/joseph/Documents/LaTeX/beamer/solutions/generic-talks/generic-ornate-15min-45min.en.tex,v 90e850259b8b 2007/01/28 20:48:30 tantau $
\documentclass{beamer}

\mode<presentation>
{
  \usetheme{AnnArbor}
  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}

\usepackage[english]{babel}

\usepackage[latin1]{inputenc}

\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage[outdir=./]{epstopdf}
\usepackage{color,soul}

\newcommand{\source}[1]{\tiny \textcolor{gray}{Source: #1}}
\newcommand{\word}[1]{\texttt{#1}}
\newcommand{\wordvec}[1]{\textbf{v}[#1]}
\newcommand{\softmax}{\text{softmax}}
\newcommand{\cbow}{\text{CBOW}}
\newcommand{\skipgram}{\text{skipgram}}
\newcommand{\currentword}{{w_0}}
\newcommand{\context}{\mathcal{C}}
\newcommand{\vocabulary}{W}
\newcommand{\p}{P}
\newcommand{\crossentropy}{H}
\newcommand{\likelihood}{\mathcal L}
\newcommand{\correctoutput}{y}
\newcommand{\prediction}{\hat{y}}
\newcommand{\target}{y}
\newcommand{\sigmoid}{\sigma}
\newcommand{\transpose}[1]{#1^\top}
\newcommand{\smparam}{\beta}
\newcommand{\hsparam}{\theta}
\newcommand{\nsparam}{\gamma}

\title[Corpus Experiments]{Corpus Experiments for Word Embeddings}

%\subtitle
%{Presentation Subtitle} % (optional)

\author
{Benjamin Wilson (with Adriaan Schakel)}

\date[Berlin ML Learning Group] % (optional)
{Berlin ML Learning Group, July 28 2015}

\begin{document}
\graphicspath{{../outputs/}}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

\section{Word Embeddings}

\begin{frame}{Word Embeddings}

\begin{itemize}
\item
associate words with points in space
\item
spatially encode word meaning and relationships between words
\item
learned from input texts
\item
e.g. word2vec, GloVe, ...
\end{itemize}

\end{frame}

\begin{frame}{Pictorially}
\begin{itemize}
\item spatial distance corresponds to word similarity
\item words are close together $\Leftrightarrow$ their "meanings" are similar
\end{itemize}
\begin{center}
    \includegraphics[height=140px]{semantic_similarity}\\
\end{center}
\end{frame}

\begin{frame}{Co-occurrence distributions}
\begin{itemize}
	\item Suppose we read the word $\word{cat}$.  What is the probability $\p ( w | \word{cat})$ that we'll read the word $w$ nearby?
\begin{center}
\includegraphics[height=30px]{dh}
\end{center}
\item $\p ( \cdot | \word{cat})$ is the \textit{co-occurrence distribution} of the word $\word{cat}$.
\item Distributional Hypothesis: the meaning of $\word{cat}$ is captured by the co-occurrence distribution.
\item Word embeddings are trained by sampling from the co-occurrence distribution.
\end{itemize}
\end{frame}

\section{Corpus Experimentalism}

\begin{frame}{Some Questions about Word Embeddings}
How would the word vector change if ...
\begin{enumerate}
	\item ... the word were less frequent? more frequent?
	\item ... the word were less informative? i.e. if $\p ( \cdot | \word{cat})$ were noisier? less noisy?
\end{enumerate}
\end{frame}

\begin{frame}{Different Approaches}
We could attempt to answer such questions by either:
\begin{enumerate}
	\item reasoning about the mechanism (\textbf{theoretical}).
	\item studying the data (\textbf{observational}).
	\item studying how the output changes when the input is varied in a controlled manner (\textbf{experimental}).
\end{enumerate}
\end{frame}

\section{Word Frequency Experiment}

\begin{frame}{Observing word frequency}
\begin{itemize}
	\item word frequency is encoded in vector length
		\begin{center}
		\includegraphics[height=80px]{frequency-norm-scatterplot} \\
		\end{center}
	\item but word frequency is related to many other things (e.g. significance)
	\item so difficult to determine \textbf{observationally} if vector length is related to word frequency \textbf{directly}
	\item for this, an experimental approach is required
\end{itemize}
\end{frame}

\begin{frame}{Word Frequency Experiment}
\begin{itemize}
	\item Modify the training corpus, replacing all occurrences of the word $\word{cat}$ with a token chosen from the distribution:
	\begin{center}
	\includegraphics[height=120px]{truncated-geometric-distribution} \\
	\end{center}
	\item Thus $\word{cat}$ is replaced by a family of tokens with varying frequencies, but the same co-occurrence distribution as $\word{cat}$.
	\item Train the word embedding on the modified corpus
	\item Study the word vectors of $\word{CAT\_i}$ for $i = 1, 2, \dots$.
\end{itemize}
\end{frame}

\begin{frame}{Word Frequency Experiment Example}
	\par
	{\small
\texttt{a pedigreed \textcolor{red}{CAT\_2} is one whose ancestry is recorded by a \textcolor{red}{CAT\_5} fancier organization
a purebred \textcolor{red}{CAT\_2} is one whose ancestry contains only individuals of the same breed
the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_1} is not owned by any one individual
the domestic \textcolor{red}{CAT\_1} was first classified as felis catus
the domesticated \textcolor{red}{CAT\_1} and its closest wild ancestor are both diploid organisms that possess  chromosomes and roughly  genes
...
dogs perform many roles for people such as hunting herding pulling loads protection assisting police and military companionship and more recently aiding handicapped individuals in  carl linnaeus listed among the types of quadrupeds familiar to him the latin word for dog canis among the species within this genus linnaeus listed the red fox as canis vulpes wolves canis lupus and the domestic dog canis canis in later editions linnaeus dropped canis canis and greatly expanded his list of the canis genus of quadrupeds and by  included alongside the foxes wolves and jackals and many more terms that are now listed as synonyms for domestic dog including aegyptius hairless dog aquaticus water dog and mustelinus literally badger dog among these were two that later experts have been widely used for domestic dogs as a species: canis domesticus and most predominantly canis familiaris the common or familiar dog} \par}
\end{frame}

\begin{frame}{Word Frequency Experiment Results (word2vec)}
	\begin{center}
	\includegraphics[height=200px]{word-frequency-experiment-heatmap} \\
	\end{center}
\end{frame}

\begin{frame}{Word Frequency Experiment Results (word2vec)}
	\begin{center}
	\includegraphics[height=150px]{word-frequency-experiment-graph} \\
	\end{center}
\end{frame}

\begin{frame}{Word Frequency Experiment Results (word2vec)}
	\begin{itemize}
		\item Vector direction is independent of frequency.
		\item Vector length depends directly and linearly on word frequency (for the word vectors, ``syn0'').
	\end{itemize}
\end{frame}

\section{Co-occurrence Noise Experiment}

\begin{frame}{Co-occurrence Noise Experiment}
\begin{itemize}
	\item Objective: vary the level of noise in the co-occurrence distribution, without varying the frequency.
	\item Our noise distribution is the global word frequency distibution.
	\item Thus we can introduce noise into the co-occurrence distribution of a word by randomly displacing some of its original occurrences throughout the corpus.
\end{itemize}
\end{frame}

\begin{frame}{Co-occurrence Noise Experiment Recipe}
How, exactly?
\begin{enumerate}
	\item Modify corpus as before, replacing the word $\word{cat}$ by tokens $\word{CAT\_i}$ for $i = 1, 2, \dots$ with varying frequencies.
	\item Introduce random occurrences of $\word{CAT\_i}$ into the corpus, such that its frequency is raised to that of $\word{cat}$.
\end{enumerate}
\end{frame}

\begin{frame}{Co-occurrence Noise Experiment Example}
	\par
	{\small
		\texttt{ a pedigreed \textcolor{red}{CAT\_1} is \textcolor{red}{\textit{CAT\_3}} one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization
a purebred \textcolor{red}{CAT\_3} is one whose ancestry contains only individuals of the same breed
the semiferal \textcolor{red}{CAT\_2} a mostly outdoor \textcolor{red}{CAT\_2} is not owned by any one individual
\textcolor{red}{\textit{CAT\_2}} the domestic \textcolor{red}{CAT\_3} was first classified as felis catus
the domesticated \textcolor{red}{CAT\_1} and its closest wild ancestor are both diploid organisms \textcolor{red}{\textit{CAT\_3}} that possess  chromosomes and roughly  genes
...
the domestic dog canis lupus familiaris or canis familiaris is a domesticated canid which has been selectively bred for millennia for various behaviors sensory capabilities and physical attributes
although initially thought to have originated as a manmade variant of an extant canid species variously supposed as being the \textcolor{red}{\textit{CAT\_1}} dhole golden jackal or gray wolf \textcolor{red}{\textit{CAT\_3}} extensive genetic studies undertaken during the s indicate that dogs \textcolor{red}{\textit{CAT\_2}} diverged from a now-extinct canid in eurasia
} \par}
\end{frame}

\begin{frame}{Co-occurrence Noise Experiment Results (word2vec)}
	\begin{center}
	\includegraphics[height=200px]{cooccurrence-noise-heatmap} \\
	\end{center}
\end{frame}

\begin{frame}{Co-occurrence Noise Experiment Results (word2vec)}
	\begin{center}
	\includegraphics[height=150px]{cooccurrence-noise-graph} \\
	\end{center}
\end{frame}

\begin{frame}{Co-occurrence Noise Experiment Results (word2vec)}
	\begin{itemize}
		\item Vector direction is remarkably insensitive to co-occurrence noise.
		\item Vector length depends directly, linearly, on the level of co-occurrence noise.
		\item This suggests that, within any frequency band, vector length can be taken as a measure of signal strength.
	\end{itemize}
\end{frame}

\section{Appendix: word2vec}
\begin{frame}{Learning from text}
\begin{itemize}
\item word2vec learns from input text
\item considers each word \textcolor{red}{$\currentword$} in turn, along with its context \textcolor{red}{$\context$}
\item context = neighbouring words (here, for simplicity, 2 words forward and back)
\begin{center}
\includegraphics[height=45px]{context} \\
\vspace{0.5cm}
\begin{tabular}{c | c | c}
sample \# & $\currentword$ & context $\context$ \\ \hline
	1 & \word{once} & \{\word{upon}, \word{a}\} \\
 & $\cdots$ &  \\
4 & \word{time} & \{\word{upon}, \word{a}, \word{in}, \word{a}\} \\
 & $\cdots$ &  
\end{tabular}

\end{center}
\end{itemize}
\end{frame}

\begin{frame}{Two approaches: CBOW and Skip-gram}
word2vec can learn the word vectors via two distinct learning tasks, \textcolor{red}{CBOW} and \textcolor{red}{Skip-gram}.
\begin{itemize}
\item CBOW: predict the current word $\currentword$ given only $\context$
\item Skip-gram: predict words from $\context$ given $\currentword$
\item Skip-gram produces better word vectors for infrequent words
\item CBOW is faster by a factor of window size -- more appropriate for larger corpora
\item We will speak only of CBOW (life is short).
\end{itemize}
\end{frame}

\begin{frame}{CBOW learning task}
\begin{itemize}
	\item Given only the current context $\context$, e.g. $$\context = \{\word{upon}, \word{a}, \word{in}, \word{a}\}$$   predict which of all possible words is the current word $\currentword$, e.g. $\currentword =$ \word{time}.
	\item multiclass classification on the vocabulary $\textcolor{red}{\vocabulary}$
	\item output is $\textcolor{red}{\prediction} = \prediction (\context) = \p ( \cdot | \context)$ is a probability distribution on $\vocabulary$, e.g.
		\begin{center}
			\includegraphics[height=40px]{prediction}
		\end{center}
	\item train so that $\prediction$ approximates target distribution $\textcolor{red}{\target}$ -- ``one-hot'' on the current word, e.g.
		\begin{center}
			\includegraphics[height=40px]{target}
		\end{center}
\end{itemize}
\end{frame}

\begin{frame}{Training CBOW with softmax regression}
	\textbf{Model}: $$\prediction = \p ( \cdot | \context ; \alpha, \smparam) = \softmax_\smparam (\sum_{w \in \context} \alpha_w), $$ where $\alpha$, $\smparam$ are families of parameter vectors.  Pictorially:
\begin{center}
\includegraphics[height=130px]{cbow}
\end{center}
In practice, hierarchical softmax or negative sampling is used in place of softmax.
\end{frame}

\begin{frame}{Stochastic Gradient Descent}
\begin{itemize}
\item learn the model parameters (here, the linear transforms)
\item minimize the difference between output distribution $\prediction$ and target distribution $\target$, measured using the cross-entropy $\textcolor{red}{\crossentropy}$:
	$$ \crossentropy (\target, \prediction) = - \sum_{w \in \vocabulary} \target_w \log \prediction_w $$
\item given $\target$ is one-hot, same as maximizing the probability of the correct outcome 
	$$\prediction_\currentword = \p (\currentword | \context ; \alpha, \smparam).$$
\item use stochastic gradient descent: for each (current word, context) pair, update all the parameters once.
\end{itemize}
\end{frame}

\begin{frame}{Word Representation}
\textit{Post-training}, associate every word $w \in \vocabulary$ with a vector $\wordvec w$:
\begin{itemize}
\item $\wordvec w$ is the vector of synaptic strengthes connecting the input layer unit $w$ to the hidden layer
\item more meaningfully, $\wordvec w$ is the hidden-layer representation of the single-word context $\context = \{w\}$.
\item vectors are (artifically) normed to unit length (Euclidean norm), post-training.
\end{itemize}
\end{frame}

\end{document}

### 2015-08-13 markup_wordcounts.py

@@ -1,12 +1,14 @@
 import sys
 
 print r"""\begin{center}
-\begin{tabular}{c | c}
+\begin{tabular}{l | r}
 word & \# occurrences \\
 \hline"""
+wcs = []
 for line in sys.stdin:
-    word, count = line.split(',')
-    print r'\word{%s} & %s \\' % (word, count.strip())
+    word, count = line.strip().split(',')
+    wcs.append((word, int(count)))
+for word, count in sorted(wcs, lambda word, count: count):
+    print r'\word{%s} & %s \\' % (word, count)
 print r"""\end{tabular}
 \end{center}"""
-

### 2015-08-13 modify_corpus.py

@@ -24,6 +24,7 @@ print('Total words in corpus : %i' % total_words)
 frequent_words = [word for word in counts if counts[word] > experiment_word_occurrence_min]
 random.seed(random_seed)
 words_experiment_1, words_experiment_2 = [random.sample(frequent_words, number_of_experiment_words) for _ in range(2)]
+words_experiment_1.append('the') # should be unicode?
 
 with file(filenames['word_freq_experiment_words'], 'w') as f:
     for word in words_experiment_1:

### 2015-08-13 choose_experiment_words.py
makefile for experiments

@@ -1,0 +1,33 @@
+"""
+Modifies an input text for the experiments according to the parameters defined in parameters.py
+Assumes the filenames from filenames.sh
+Writes out files listing the words chosen.
+"""
+from __future__ import print_function
+import os
+from parameters import *
+from functions import *
+
+directory = os.path.dirname(os.path.realpath(__file__))
+
+filenames = dict()
+execfile(os.path.join(directory, 'filenames.sh'), filenames)
+
+counts = dict()
+with file(filenames['word_counts']) as f:
+    for line in f:
+        word, count = line.strip().split(',')
+        counts[word] = int(count)
+
+frequent_words = [word for word in counts if counts[word] > experiment_word_occurrence_min]
+random.seed(random_seed)
+words_experiment_1, words_experiment_2 = [random.sample(frequent_words, number_of_experiment_words) for _ in range(2)]
+words_experiment_1.append('the')
+
+with file(filenames['word_freq_experiment_words'], 'w') as f:
+    for word in words_experiment_1:
+        print('%s,%i' % (word, counts[word]), file=f)
+
+with file(filenames['coocc_noise_experiment_words'], 'w') as f:
+    for word in words_experiment_2:
+        print('%s,%i' % (word, counts[word]), file=f)

### 2015-08-13 count_words.py

@@ -1,0 +1,6 @@
+import sys
+from functions import *
+
+word_counts = count_words(sys.stdin)
+for word, count in word_counts.iteritems():
+    print '%s,%i' % (word, count)

### 2015-08-13 filenames.sh

@@ -1,5 +1,6 @@
-corpus_unmodified="corpus-unmodified.txt"
-corpus_modified="corpus-modified.txt"
-vectors_binary="vectors.bin"
-word_freq_experiment_words="word_freq_experiment_words"
-coocc_noise_experiment_words="coocc_noise_experiment_words"
+corpus_unmodified="outputs/corpus-unmodified.txt"
+corpus_modified="outputs/corpus-modified.txt"
+word_counts="outputs/word_counts_unmodified_corpus.csv"
+vectors_binary="outputs/vectors.bin"
+word_freq_experiment_words="outputs/word_freq_experiment_words"
+coocc_noise_experiment_words="outputs/coocc_noise_experiment_words"

### 2015-08-13 functions.py

@@ -12,7 +12,7 @@ def count_words(f):
     """
     counts = defaultdict(lambda: 0)
     for line in f:
-        for word in line.split(' '):
+        for word in line.strip().split(' '):
                 counts[word] += 1
     return counts
 

### 2015-08-13 modify_corpus.py

@@ -1,7 +1,6 @@
 """
 Modifies an input text for the experiments according to the parameters defined in parameters.py
 Assumes the filenames from filenames.sh
-Writes out files listing the words chosen.
 Requires sufficient diskspace to write out the modified text at intermediate steps.
 """
 from __future__ import print_function
@@ -10,35 +9,37 @@ from parameters import *
 from functions import *
 
 directory = os.path.dirname(os.path.realpath(__file__))
-
 filenames = dict()
 execfile(os.path.join(directory, 'filenames.sh'), filenames)
 
-intermediate_file = 'delete.me'
+def read_words(filename):
+    words = []
+    with file(filename) as f:
+        for line in f:
+            word = line.split(',')[0]
+            words.append(word)
 
-with file(filenames['corpus_unmodified']) as f:
-    counts = count_words(f)
+wf_experiment_words = read_words(filenames['word_frequency_experiment_words']) #FIXME change names
+cn_experiment_words = read_words(filenames['coocc_noise_experiment_words'])
+
+counts = dict()
+with file(filenames['word_counts']) as f:
+    for line in f:
+        word, count = line.strip().split(',')
+        counts[word] = int(count)
 total_words = sum(counts.values())
-print('Total words in corpus : %i' % total_words)
 
-frequent_words = [word for word in counts if counts[word] > experiment_word_occurrence_min]
-random.seed(random_seed)
-words_experiment_1, words_experiment_2 = [random.sample(frequent_words, number_of_experiment_words) for _ in range(2)]
-words_experiment_1.append('the') # should be unicode?
-
-with file(filenames['word_freq_experiment_words'], 'w') as f:
-    for word in words_experiment_1:
-        print('%s,%i' % (word, counts[word]), file=f)
-
-with file(filenames['coocc_noise_experiment_words'], 'w') as f:
-    for word in words_experiment_2:
-        print('%s,%i' % (word, counts[word]), file=f)
+# intersperse the meaningless token throughout the corpus
+intermediate_file = 'delete.me'
+with open(filenames['corpus_unmodified']) as f_in, open(intermediate_file, 'w') as f_out:
+    intersperse_words({meaningless_token: meaningless_token_frequency}, f_in, f_out)
+word_frequency_experiment_words.append(meaningless_token)
 
 # perform the replacement procedures for the word frequency and the noise cooccurrence experiments
 word_samplers = {}
[…]

### 2015-08-13 parameters.py

@@ -1,6 +1,9 @@
 random_seed = 3 # fix random seed for reproducibility
 number_of_experiment_words = 8
 
+meaningless_token = 'VOID'
+meaningless_token_frequency = 0.005 # frequency before the replacement procedure
+
 experiment_word_occurrence_min = 10000
 
 word_freq_experiment_ratio = 1./2

### 2015-08-13 build_images.py
first attempt at turning notebook into build_images

@@ -1,0 +1,154 @@
+import numpy as np
+import pandas as pd
+import matplotlib
+import matplotlib.pyplot as plt
+import random
+
+from parameters import *
+from functions import *
+
+filenames = dict()
+execfile('filenames.sh', filenames)
+
+matplotlib.rcParams.update({'font.size': 14})
+
+word_freq_experiment_words = read_words(filenames['word_freq_experiment_words'])
+coocc_noise_experiment_words = read_words(filenames['coocc_noise_experiment_words'])
+
+vectors_syn0 = load_word2vec_binary(filenames['vectors_binary'])
+norms_syn0 = np.sqrt((vectors_syn0 ** 2).sum(axis=1))
+
+vectors_syn1neg = load_word2vec_binary(filenames['vectors_binary'] + '.syn1neg')
+norms_syn1neg = np.sqrt((vectors_syn1neg ** 2).sum(axis=1))
+
+vocab = list(vectors_syn0.index)
+
+# Calculate frequencies in the modified corpus
+counts = read_word_counts(filenames['word_counts_modified_corpus'])
+
+stats = pd.DataFrame({'occurrences': new_counts, 'L2_norm_syn0': norms_syn0, 'L2_norm_syn1neg': norms_syn1neg}).dropna()
+stats['occurrences_band'] = np.floor(np.log2(stats.occurrences)).astype(int)
+stats['log2_frequency'] = np.log2(stats.occurrences * 1. / total_words)
+stats.L2_norm_syn0.name = 'L2 norm (syn0)'
+stats.L2_norm_syn1neg.name = 'L2 norm (syn1neg)'
+
+band_counts = stats.occurrences_band.dropna().value_counts().sort_index()
+band_counts.index.name = 'log2 (# occurrences)'
+
+_ = band_counts.plot(kind='bar')
+plt.title('Number of words in each occurrence count band', y=1.08)
+plt.tight_layout()
+plt.savefig('occurrence-histogram.eps')
+
+band_gb = stats.groupby('occurrences_band')
+means = band_gb.L2_norm_syn0.mean()
+errors = band_gb.L2_norm_syn0.std()
+ax = means.plot(yerr=errors)
+
+means = band_gb.L2_norm_syn1neg.mean()
+errors = band_gb.L2_norm_syn1neg.std()
+ax = means.plot(yerr=errors, figsize=(7,7))
+
+_ = ax.set_xlim(7, 20)
+_ = ax.set_ylim(0, 60)
+
[…]

### 2015-08-13 functions.py

@@ -16,6 +16,28 @@ def count_words(f):
                 counts[word] += 1
     return counts
 
+def read_words(filename):
+    """
+    given a CSV of word counts, returns the list of words.
+    """
+    words = []
+    with file(filename) as f:
+        for line in f:
+            word = line.split(',')[0]
+            words.append(word)
+    return words
+
+def read_word_counts(filename):
+    """
+    given a CSV of word counts, returns a word count dictionary.
+    """
+    counts = dict()
+    with file(filename) as f:
+        for line in f:
+            word, count = line.strip().split(',')
+            counts[word] = int(count)
+    return dict
+
 def build_experiment_token(word, sample):
     """
     e.g. ('cat', 3) -> CAT_3
@@ -111,4 +133,13 @@ def load_word2vec_binary(fname):
             vectors[line_no] = vector
     return pd.DataFrame(vectors, index=vocab)
 
+def cosine_similarity_heatmap(test_vecs, **kwargs):
+    test_vecs_normed = test_vecs.as_matrix() / np.sqrt((test_vecs ** 2).sum(axis=1))[:,np.newaxis]
+    mat = test_vecs_normed.dot(test_vecs_normed.transpose())    
+    plt.figure(**kwargs)
+    plt.title('Dot product of normalised word vectors')
+    plt.pcolor(mat, vmin=-1, vmax=1)
+    plt.colorbar()
+    _ = plt.yticks(np.arange(0.5, len(test_vecs.index), 1), test_vecs.index)
+    _ = plt.xticks(np.arange(0.5, len(test_vecs.index), 1), test_vecs.index, rotation=90)
 

### 2015-08-13 modify_corpus.py

@@ -12,21 +12,10 @@ directory = os.path.dirname(os.path.realpath(__file__))
 filenames = dict()
 execfile(os.path.join(directory, 'filenames.sh'), filenames)
 
-def read_words(filename):
-    words = []
-    with file(filename) as f:
-        for line in f:
-            word = line.split(',')[0]
-            words.append(word)
-
-wf_experiment_words = read_words(filenames['word_frequency_experiment_words']) #FIXME change names
+wf_experiment_words = read_words(filenames['word_freq_experiment_words'])
 cn_experiment_words = read_words(filenames['coocc_noise_experiment_words'])
 
-counts = dict()
-with file(filenames['word_counts']) as f:
-    for line in f:
-        word, count = line.strip().split(',')
-        counts[word] = int(count)
+counts = read_word_counts(filenames['word_counts_unmodified_corpus'])
 total_words = sum(counts.values())
 
 # intersperse the meaningless token throughout the corpus

### 2015-08-13 build_images.py
makefile and build_images working

@@ -1,6 +1,6 @@
 import numpy as np
 import pandas as pd
-import matplotlib
+import matplotlib; matplotlib.use('Agg') # must be set before importing pyplot
 import matplotlib.pyplot as plt
 import random
 
@@ -24,7 +24,8 @@ norms_syn1neg = np.sqrt((vectors_syn1neg ** 2).sum(axis=1))
 vocab = list(vectors_syn0.index)
 
 # Calculate frequencies in the modified corpus
-counts = read_word_counts(filenames['word_counts_modified_corpus'])
+new_counts = read_word_counts(filenames['word_counts_modified_corpus'])
+total_words = sum(new_counts.values())
 
 stats = pd.DataFrame({'occurrences': new_counts, 'L2_norm_syn0': norms_syn0, 'L2_norm_syn1neg': norms_syn1neg}).dropna()
 stats['occurrences_band'] = np.floor(np.log2(stats.occurrences)).astype(int)
@@ -38,7 +39,7 @@ band_counts.index.name = 'log2 (# occurrences)'
 _ = band_counts.plot(kind='bar')
 plt.title('Number of words in each occurrence count band', y=1.08)
 plt.tight_layout()
-plt.savefig('occurrence-histogram.eps')
+plt.savefig('outputs/occurrence-histogram.eps')
 
 band_gb = stats.groupby('occurrences_band')
 means = band_gb.L2_norm_syn0.mean()
@@ -57,12 +58,12 @@ _ = ax.set_ylabel('L2 norm')
 
 plt.title('Mean and Std. of norm as function of # occurrences', y=1.04)
 plt.legend(['syn0 vectors', 'syn1neg vectors'])
-plt.savefig('frequency-norm-graph.eps')
+plt.savefig('outputs/frequency-norm-graph.eps')
 
 # scatter plot of syn0 norm vs frequency for a sample of ordinary (non-experiment-) words
 fig = plt.figure(figsize=(16, 8))
 non_experiment_words = [word for word in stats.index if word != word.upper()]
-sample = stats.loc[random.sample(non_experiment_words, 20000)]
+sample = stats.loc[random.sample(non_experiment_words, 15000)]
 plt.scatter(sample.occurrences, sample.L2_norm_syn0, s=0.2)
 plt.xscale('log')
 plt.ylim(0, 50)
[…]

### 2015-08-13 filenames.sh

@@ -1,6 +1,7 @@
 corpus_unmodified="outputs/corpus-unmodified.txt"
 corpus_modified="outputs/corpus-modified.txt"
-word_counts="outputs/word_counts_unmodified_corpus.csv"
+word_counts_unmodified_corpus="outputs/word_counts_unmodified_corpus.csv"
+word_counts_modified_corpus="outputs/word_counts_modified_corpus.csv"
 vectors_binary="outputs/vectors.bin"
 word_freq_experiment_words="outputs/word_freq_experiment_words"
 coocc_noise_experiment_words="outputs/coocc_noise_experiment_words"

### 2015-08-13 functions.py

@@ -2,6 +2,7 @@ from __future__ import print_function
 import random
 import numpy as np
 import pandas as pd
+import matplotlib.pyplot as plt
 from collections import defaultdict
 from parameters import random_seed, experiment_word_occurrence_min
 
@@ -36,7 +37,7 @@ def read_word_counts(filename):
         for line in f:
             word, count = line.strip().split(',')
             counts[word] = int(count)
-    return dict
+    return counts
 
 def build_experiment_token(word, sample):
     """

### 2015-08-13 modify_corpus.py

@@ -22,7 +22,7 @@ total_words = sum(counts.values())
 intermediate_file = 'delete.me'
 with open(filenames['corpus_unmodified']) as f_in, open(intermediate_file, 'w') as f_out:
     intersperse_words({meaningless_token: meaningless_token_frequency}, f_in, f_out)
-word_frequency_experiment_words.append(meaningless_token)
+wf_experiment_words.append(meaningless_token)
 
 # perform the replacement procedures for the word frequency and the noise cooccurrence experiments
 word_samplers = {}

### 2015-08-13 word2vec/word2vec-normalised.c

@@ -1,638 +1,0 @@
-//  Copyright 2013 Google Inc. All Rights Reserved.
-//
-//  Licensed under the Apache License, Version 2.0 (the "License");
-//  you may not use this file except in compliance with the License.
-//  You may obtain a copy of the License at
-//
-//      http://www.apache.org/licenses/LICENSE-2.0
-//
-//  Unless required by applicable law or agreed to in writing, software
-//  distributed under the License is distributed on an "AS IS" BASIS,
-//  WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-//  See the License for the specific language governing permissions and
-//  limitations under the License.
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <math.h>
-#include <pthread.h>
-
-#define MAX_STRING 100
-#define EXP_TABLE_SIZE 1000
-#define MAX_EXP 6
-#define MAX_SENTENCE_LENGTH 1000
-#define MAX_CODE_LENGTH 40
-
-const int vocab_hash_size = 30000000;  // Maximum 30 * 0.7 = 21M words in the vocabulary
-
-typedef float real;                    // Precision of float numbers
-
-struct vocab_word {
-  long long cn;
-  int *point;
-  char *word, *code, codelen;
-};
-
-char train_file[MAX_STRING], output_file[MAX_STRING];
-char save_vocab_file[MAX_STRING], read_vocab_file[MAX_STRING];
-struct vocab_word *vocab;
-int binary = 0, cbow = 1, debug_mode = 2, window = 5, min_count = 5, num_threads = 12, min_reduce = 1;
-int *vocab_hash;
-long long vocab_max_size = 1000, vocab_size = 0, layer1_size = 100;
-long long train_words = 0, word_count_actual = 0, iter = 5, file_size = 0, classes = 0;
-real alpha = 0.025, starting_alpha, sample = 1e-3;
-real *syn0, *syn1, *syn1neg, *expTable;
-clock_t start;
-
-int hs = 0, negative = 5;
-const int table_size = 1e8;
-int *table;
-
-void InitUnigramTable() {
-  int a, i;
-  double train_words_pow = 0;
-  double d1, power = 0.75;
-  table = (int *)malloc(table_size * sizeof(int));
[…]

### 2015-08-13 build_images.py
half tidied up makefile

@@ -24,7 +24,8 @@ norms_syn1neg = np.sqrt((vectors_syn1neg ** 2).sum(axis=1))
 vocab = list(vectors_syn0.index)
 
 # Calculate frequencies in the modified corpus
-new_counts = read_word_counts(filenames['word_counts_modified_corpus'])
+with file(filenames['word_counts_modified_corpus']) as f:
+    new_counts = read_word_counts(f)
 total_words = sum(new_counts.values())
 
 stats = pd.DataFrame({'occurrences': new_counts, 'L2_norm_syn0': norms_syn0, 'L2_norm_syn1neg': norms_syn1neg}).dropna()
@@ -78,34 +79,47 @@ def set_num_plots(num_plots):
     colormap = plt.cm.gist_ncar
     plt.gca().set_color_cycle([colormap(i) for i in np.linspace(0, 0.9, num_plots)])
 
+words = word_freq_experiment_words
+fig = plt.figure(figsize=(16, 6))
+colormap = plt.cm.gist_ncar
+colorcycle = [colormap(i) for i in np.linspace(0, 0.9, len(words))]
 
-def plot_word_freq_experiment_norm_vs_freq(norms_vs_freq):
-    plt.figure(figsize=(7, 7))
-    plt.title('Norm of word vector when word frequency is varied', y=1.04)
-    plt.xlabel('log2 (# occurrences)')
-    plt.ylabel(norms_vs_freq.name)
-    set_num_plots(len(word_freq_experiment_words))
+ax_syn0 = plt.subplot(131)
+ax_syn0.set_xlabel('# occurrences')
+ax_syn0.set_ylabel('L2 norm')
+ax_syn0.set_xscale('log')
+ax_syn0.set_color_cycle(colorcycle)
+ax_syn0.set_title('syn0', y=1.04)
 
-    def plot_for_word(word, **kwargs):
-        idxs = [build_experiment_token(word, i) for i in range(1, max(word_freq_experiment_ratio, word_freq_experiment_power_max) + 1)]
-        x = np.log2(stats.loc[idxs].occurrences)
-        y = norms_vs_freq.loc[idxs]
-        plt.plot(x, y, marker='o', **kwargs)
+ax_syn1neg = plt.subplot(132, sharex=ax_syn0, sharey=ax_syn0)
+ax_syn1neg.set_xlabel('# occurrences')
+ax_syn1neg.set_color_cycle(colorcycle)
+ax_syn1neg.set_title('syn1neg', y=1.04)
+    
[…]

### 2015-08-13 choose_experiment_words.py

@@ -1,33 +1,17 @@
 """
 Modifies an input text for the experiments according to the parameters defined in parameters.py
-Assumes the filenames from filenames.sh
-Writes out files listing the words chosen.
+Writes out the words chosen to stdout.
+Takes a random seed as the first argument.
 """
-from __future__ import print_function
-import os
+import sys
 from parameters import *
 from functions import *
 
-directory = os.path.dirname(os.path.realpath(__file__))
-
-filenames = dict()
-execfile(os.path.join(directory, 'filenames.sh'), filenames)
-
-counts = dict()
-with file(filenames['word_counts']) as f:
-    for line in f:
-        word, count = line.strip().split(',')
-        counts[word] = int(count)
+random.seed(sys.argv[1])
+counts = read_word_counts(sys.stdin)
 
 frequent_words = [word for word in counts if counts[word] > experiment_word_occurrence_min]
-random.seed(random_seed)
-words_experiment_1, words_experiment_2 = [random.sample(frequent_words, number_of_experiment_words) for _ in range(2)]
-words_experiment_1.append('the')
+experiment_words = random.sample(frequent_words, number_of_experiment_words)
 
-with file(filenames['word_freq_experiment_words'], 'w') as f:
-    for word in words_experiment_1:
-        print('%s,%i' % (word, counts[word]), file=f)
-
-with file(filenames['coocc_noise_experiment_words'], 'w') as f:
-    for word in words_experiment_2:
-        print('%s,%i' % (word, counts[word]), file=f)
+for word in experiment_words:
+    print '%s,%i' % (word, counts[word])

### 2015-08-13 functions.py

@@ -28,15 +28,14 @@ def read_words(filename):
             words.append(word)
     return words
 
-def read_word_counts(filename):
+def read_word_counts(f):
     """
     given a CSV of word counts, returns a word count dictionary.
     """
     counts = dict()
-    with file(filename) as f:
-        for line in f:
-            word, count = line.strip().split(',')
-            counts[word] = int(count)
+    for line in f:
+        word, count = line.strip().split(',')
+        counts[word] = int(count)
     return counts
 
 def build_experiment_token(word, sample):

### 2015-08-13 modify_corpus.py

@@ -3,7 +3,6 @@ Modifies an input text for the experiments according to the parameters defined i
 Assumes the filenames from filenames.sh
 Requires sufficient diskspace to write out the modified text at intermediate steps.
 """
-from __future__ import print_function
 import os
 from parameters import *
 from functions import *
@@ -15,7 +14,8 @@ execfile(os.path.join(directory, 'filenames.sh'), filenames)
 wf_experiment_words = read_words(filenames['word_freq_experiment_words'])
 cn_experiment_words = read_words(filenames['coocc_noise_experiment_words'])
 
-counts = read_word_counts(filenames['word_counts_unmodified_corpus'])
+with file(filenames['word_counts_unmodified_corpus']) as f:
+    counts = read_word_counts(f)
 total_words = sum(counts.values())
 
 # intersperse the meaningless token throughout the corpus

### 2015-08-14 build_images.py
removed filenames.sh; split modify_corpus.py

@@ -3,28 +3,31 @@ import pandas as pd
 import matplotlib; matplotlib.use('Agg') # must be set before importing pyplot
 import matplotlib.pyplot as plt
 import random
+import sys
 
 from parameters import *
 from functions import *
 
-filenames = dict()
-execfile('filenames.sh', filenames)
+vectors_syn0_filename = sys.argv[1]
+vectors_syn1neg_filename = sys.argv[2]
+word_counts_filename = sys.argv[3]
+word_freq_exp_words_filename = sys.argv[4]
+coocc_noise_exp_words_filename = sys.argv[5]
 
 matplotlib.rcParams.update({'font.size': 14})
 
-word_freq_experiment_words = read_words(filenames['word_freq_experiment_words'])
-coocc_noise_experiment_words = read_words(filenames['coocc_noise_experiment_words'])
+word_freq_experiment_words = read_words(word_freq_exp_words_filename)
+coocc_noise_experiment_words = read_words(coocc_noise_exp_words_filename)
 
-vectors_syn0 = load_word2vec_binary(filenames['vectors_binary'])
+vectors_syn0 = load_word2vec_binary(vectors_syn0_filename)
 norms_syn0 = np.sqrt((vectors_syn0 ** 2).sum(axis=1))
 
-vectors_syn1neg = load_word2vec_binary(filenames['vectors_binary'] + '.syn1neg')
+vectors_syn1neg = load_word2vec_binary(vectors_syn1neg_filename)
 norms_syn1neg = np.sqrt((vectors_syn1neg ** 2).sum(axis=1))
 
 vocab = list(vectors_syn0.index)
 
-# Calculate frequencies in the modified corpus
-with file(filenames['word_counts_modified_corpus']) as f:
+with file(word_counts_filename) as f:
     new_counts = read_word_counts(f)
 total_words = sum(new_counts.values())
 

### 2015-08-14 filenames.sh

@@ -1,7 +1,0 @@
-corpus_unmodified="outputs/corpus-unmodified.txt"
-corpus_modified="outputs/corpus-modified.txt"
-word_counts_unmodified_corpus="outputs/word_counts_unmodified_corpus.csv"
-word_counts_modified_corpus="outputs/word_counts_modified_corpus.csv"
-vectors_binary="outputs/vectors.bin"
-word_freq_experiment_words="outputs/word_freq_experiment_words"
-coocc_noise_experiment_words="outputs/coocc_noise_experiment_words"

### 2015-08-14 modify_corpus.py

@@ -1,50 +1,0 @@
-"""
-Modifies an input text for the experiments according to the parameters defined in parameters.py
-Assumes the filenames from filenames.sh
-Requires sufficient diskspace to write out the modified text at intermediate steps.
-"""
-import os
-from parameters import *
-from functions import *
-
-directory = os.path.dirname(os.path.realpath(__file__))
-filenames = dict()
-execfile(os.path.join(directory, 'filenames.sh'), filenames)
-
-wf_experiment_words = read_words(filenames['word_freq_experiment_words'])
-cn_experiment_words = read_words(filenames['coocc_noise_experiment_words'])
-
-with file(filenames['word_counts_unmodified_corpus']) as f:
-    counts = read_word_counts(f)
-total_words = sum(counts.values())
-
-# intersperse the meaningless token throughout the corpus
-intermediate_file = 'delete.me'
-with open(filenames['corpus_unmodified']) as f_in, open(intermediate_file, 'w') as f_out:
-    intersperse_words({meaningless_token: meaningless_token_frequency}, f_in, f_out)
-wf_experiment_words.append(meaningless_token)
-
-# perform the replacement procedures for the word frequency and the noise cooccurrence experiments
-word_samplers = {}
-for word in wf_experiment_words:
-    word_samplers[word] = truncated_geometric_sampling(word, word_freq_experiment_ratio, word_freq_experiment_power_max)
-for word in cn_experiment_words:
-    word_samplers[word] = truncated_geometric_sampling(word, coocc_noise_experiment_ratio, coocc_noise_experiment_power_max)
-
-tmp_file = 'delete.me.2'
-with open(intermediate_file) as f_in, open(tmp_file, 'w') as f_out:
-    replace_words(word_samplers, f_in, f_out)
-os.remove(intermediate_file)
-intermediate_file = tmp_file
-
-# add noise to the cooccurrence distributions
-token_freq_dict = dict()
-for word in cn_experiment_words:
-    target_freq = counts[word] * 1. / total_words
-    for i in range(1, coocc_noise_experiment_power_max + 1):
-        current_freq = target_freq * truncated_geometric_proba(coocc_noise_experiment_ratio, i, coocc_noise_experiment_power_max)
-        token_freq_dict[build_experiment_token(word, i)] = target_freq - current_freq
-
-with open(intermediate_file) as f_in, open(filenames['corpus_modified'], 'w') as f_out:
-    intersperse_words(token_freq_dict, f_in, f_out)
-os.remove(intermediate_file)

### 2015-08-14 modify_corpus_coocc_noise_experiment.py

@@ -1,0 +1,35 @@
+"""
+Modifies an input text for the experiments according to the parameters defined in parameters.py
+Reads from stdin, writes to stdout.
+Requires sufficient diskspace to write out the modified text at intermediate steps.
+"""
+import os
+import sys
+from parameters import *
+from functions import *
+
+cn_experiment_words = read_words(sys.argv[1])
+with file(sys.argv[2]) as f:
+    counts = read_word_counts(f)
+total_words = sum(counts.values())
+
+# perform the replacement procedure
+word_samplers = {}
+for word in cn_experiment_words:
+    word_samplers[word] = truncated_geometric_sampling(word, coocc_noise_experiment_ratio, coocc_noise_experiment_power_max)
+
+intermediate_file = 'delete.me'
+with open(intermediate_file, 'w') as f_out:
+    replace_words(word_samplers, sys.stdin, f_out)
+
+# add noise to the cooccurrence distributions
+token_freq_dict = dict()
+for word in cn_experiment_words:
+    target_freq = counts[word] * 1. / total_words
+    for i in range(1, coocc_noise_experiment_power_max + 1):
+        current_freq = target_freq * truncated_geometric_proba(coocc_noise_experiment_ratio, i, coocc_noise_experiment_power_max)
+        token_freq_dict[build_experiment_token(word, i)] = target_freq - current_freq
+
+with open(intermediate_file) as f_in:
+    intersperse_words(token_freq_dict, f_in, sys.stdout)
+os.remove(intermediate_file)

### 2015-08-14 modify_corpus_word_freq_experiment.py

@@ -1,0 +1,29 @@
+"""
+Modifies an input text for the word frequency experiment.
+Reads from stdin, writes to stdout.
+Requires sufficient diskspace to write out the modified text at intermediate steps.
+"""
+import os
+import sys
+from parameters import *
+from functions import *
+
+wf_experiment_words = read_words(sys.argv[1])
+with file(sys.argv[2]) as f:
+    counts = read_word_counts(f)
+total_words = sum(counts.values())
+
+# intersperse the meaningless token throughout the corpus
+intermediate_file = 'delete.me'
+with open(intermediate_file, 'w') as f_out:
+    intersperse_words({meaningless_token: meaningless_token_frequency}, sys.stdin, f_out)
+wf_experiment_words.append(meaningless_token)
+
+# perform the replacement procedure
+word_samplers = {}
+for word in wf_experiment_words:
+    word_samplers[word] = truncated_geometric_sampling(word, word_freq_experiment_ratio, word_freq_experiment_power_max)
+
+with open(intermediate_file) as f_in:
+    replace_words(word_samplers, f_in, sys.stdout)
+os.remove(intermediate_file)

### 2015-08-14 clean_corpus.py
clean corpus is independent of pandas

@@ -1,12 +1,7 @@
-HELP_STR = """
-Cleans a two column CSV (id, text) read from stdin, removing punctuation, optionally
-lowercasing, and writing out clean UTF-8 text.  Decoding errors are ignored and
-lines whose cleaning results in other exceptions are skipped. Unescapes \\n linebreaks."""
-
-import pandas as pd
-import argparse
 import sys
+import csv
 import re
+csv.field_size_limit(sys.maxsize)
 
 """
 words consist of 1 or more alphanumeric characters where there first is not
@@ -14,33 +9,11 @@ numeric and not an underscore. alphanumeric is defined in the unicode sense.
 includes also all of greek, arabic, etc.
 """
 TOKEN_PATTERN = r'(?u)\b[^(\W|\d|_)]{1,}\w+\b'
-ESCAPED_SEQ = r'\n'
-UNESCAPE_PROG = re.compile(re.escape(ESCAPED_SEQ))
 
-if __name__ == '__main__':
-    parser = argparse.ArgumentParser(description=HELP_STR)
-    parser.add_argument(
-        '--case-sensitive', action='store_true')
-    args = parser.parse_args()
-
-    tokenizer = re.compile(TOKEN_PATTERN)
-    unescape_line_breaks = lambda unicode_text: UNESCAPE_PROG.sub(u'\n', unicode_text)
-
-
-    doc_chunks = pd.read_csv(sys.stdin,
-                             header=None,
-                             index_col=0,
-                             squeeze=True,
-                             chunksize=1000)
-    for chunk in doc_chunks:
-        chunk = chunk.str.decode('utf-8').fillna(u'')
-        chunk = chunk.apply(unescape_line_breaks)
-        for _, text in chunk.iteritems():
-            try:
-                if not args.case_sensitive:
-                    text = text.lower()
-                tokens = tokenizer.findall(text)
-                cleaned = u' '.join(tokens)
-                print cleaned.encode('utf-8')
-            except ValueError as e:
-                pass
+tokenizer = re.compile(TOKEN_PATTERN)
+reader = csv.reader(sys.stdin, delimiter=',', quotechar='"')
+for text_id, text in reader:
+    text = text.decode('utf-8').lower()
+    tokens = tokenizer.findall(text)
+    cleaned = u' '.join(tokens)
+    print cleaned.encode('utf-8')

### 2015-08-14 modify_corpus_coocc_noise_experiment.py
fix small bugs in makefile and corpus modifications

@@ -18,7 +18,7 @@ word_samplers = {}
 for word in cn_experiment_words:
     word_samplers[word] = truncated_geometric_sampling(word, coocc_noise_experiment_ratio, coocc_noise_experiment_power_max)
 
-intermediate_file = 'delete.me'
+intermediate_file = 'delete.me.word_freq_experiment'
 with open(intermediate_file, 'w') as f_out:
     replace_words(word_samplers, sys.stdin, f_out)
 

### 2015-08-14 modify_corpus_word_freq_experiment.py

@@ -14,7 +14,7 @@ with file(sys.argv[2]) as f:
 total_words = sum(counts.values())
 
 # intersperse the meaningless token throughout the corpus
-intermediate_file = 'delete.me'
+intermediate_file = 'delete.me.coocc_experiment'
 with open(intermediate_file, 'w') as f_out:
     intersperse_words({meaningless_token: meaningless_token_frequency}, sys.stdin, f_out)
 wf_experiment_words.append(meaningless_token)

### 2015-08-14 markup_wordcounts.py
fix small problems in makefile

@@ -8,7 +8,7 @@ wcs = []
 for line in sys.stdin:
     word, count = line.strip().split(',')
     wcs.append((word, int(count)))
-for word, count in sorted(wcs, lambda word, count: count):
+for word, count in sorted(wcs, lambda (word, count): count):
     print r'\word{%s} & %s \\' % (word, count)
 print r"""\end{tabular}
 \end{center}"""

### 2015-08-18 article/cooccurrence-noise-experiment.tex
generate three sample texts for the article

\texttt{\textcolor{red}{CAT\_3} the domestic \textcolor{red}{CAT\_3} was first classified as felis \textcolor{red}{CAT\_2} catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_3} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_2} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_2} is one \textcolor{red}{CAT\_2} whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_2} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated \textcolor{red}{CAT\_2} canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time \textcolor{red}{CAT\_3} at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

### 2015-08-18 article/main.tex
generate three sample texts for the article

@@
 \usepackage{hyperref}
-\usepackage{wrapfig}
+%\usepackage{wrapfig}
+\usepackage{natbib}
 \usepackage{rotating}
@@
 \urldef{\googlegroup}\url{https://groups.google.com/forum/#\!forum/word2vec-toolkit}
-\urldef{\blogpost}\url{https://blog.lateral.io/2015/06/the-unknown-perils-of-mining-wikipedia/ }
+\urldef{\blogpost}\url{https://blog.lateral.io/2015/06/the-unknown-perils-of-mining-wikipedia/}
 
@@
 
-\section{Introduction}
+\begin{section}{Introduction}
 Word embeddings are typically trained with samples from the word co-occurrence distributions.
@@
 We illustrate our approach in case of the word2vec CBOW model.
-However, these experiments (and other besides) could might equally well be performed for word embedding methods such as word2vec skip-gram, GloVe and SENNA.  \cite{}
+However, these experiments (and other besides) could might equally well be performed for word embedding methods such as word2vec skip-gram \cite{DistRepns,EfficientEstimation}, GloVe \cite{pennington2014glove} and SENNA \cite{collobert-2011}.
 In the case of word2vec, we show that word vector length depends directly and linearly with word frequency.
@@
 In the word2vec forum\footnote{\googlegroup}, it has been observed that normalisation improves performance on theses tasks since word vector length is related to word frequency.
-It remained unclear, however, whether word frequency was \textit{directly} related to word vector lengthor rather via correlated factors such as word significance.
+Figure \ref{fig:frequency-norm-graph} illustrates the global relationship between vector length and word frequency.
+It remained unclear, however, whether word frequency was \textit{directly} related to word vector length or rather via correlated factors such as word significance.
 Our experimental approach demonstrates that a direct relationship indeed exists.
@@
 
-\section{Model and corpus}
+\begin{subsection}{Previous work}
+Our experimental finding, in the context of word2vec, that word vector length decreases with co-occurrence noise is reminiscent of \cite{vecchi-baroni-zamparelli2011}, where a relationship between vector length and the ``semantic deviance'' of an adjective-noun composition was studied observationally.
+In \cite{schakel-wilson}, the interpretation of word vector length is an indication of word significance is explored in the context of the arXiv high-energy physics corpus.
+\end{subsection}
+
+\begin{figure}
+	\includegraphics[scale=0.3]{frequency-norm-scatterplot}
+	\caption{ The global relationship between word frequency and vector length.  }
+	\label{fig:frequency-norm-graph}
+\end{figure}
+\end{section}
+
+\begin{section}{Corpus and model}
 Our training data is built from the Wikipedia XML datadump from October 2013.
@@
 Identical experimental results were obtained using hierarchical softmax, but these are omitted for succinctness.
+The relatively high minimum word cut-off is chosen to ensure that word vectors receive a sufficient number of gradient updates to be meaningful.
 
@@
 
-\section{Experimental design and corpus modifications}\label{experimental_design}
-Two controlled experiments are performed:
-\begin{itemize}
-	\item \emph{Word frequency variation} compare the lengths of the word vectors for words with varying frequencies, but with the same co-occurrence distribution;
-	\item \emph{Co-occurrence noise variation} compare the lengths of the word vectors for words with of the same frequency, but with varying levels of noise in their co-occurrence distributions.
-\end{itemize}
+\begin{figure}\label{fig:frequency-histogram}
+	\includegraphics[scale=0.5]{occurrence-histogram}
+	\caption{
+	Word frequency histogram.
+	Here and elsewhere, frequency bands are delimited by powers of $2$.
+	}
+\end{figure}
 
-To perform these experiments, tokens are introduced into the corpus described in subsections \ref{WFVE} and \ref{CNVE}.
-These tokens are introduced via a replacement procedure.
+\begin{subsection}{Replacement procedure}\label{replacement-procedure}
+In the experiments, tokens are introduced into corpus via a replacement procedure.
 To illustrate this procedure, consider a word, e.g. \word{cat}.
@@
 Of course, other distributions might equally have been chosen for the experiments.
+\end{subsection}
+\end{section}
 
-\subsection{Word frequency variation experiment}\label{WFVE}
-Firstly, a small set of words from the unmodified corpus is chosen uniformly at random from the vocabulary.
+\begin{section}{Word frequency variation experiment}\label{WFVE}
+In this experiment, families of tokens are introduced into the corpus.
+The tokens in each family vary in frequency but share a common co-occurrence distribution.
+After model training, comparison of the resulting word vectors for these tokens reveals the independent effect of frequency variation on the word embedding.
+
+\begin{subsection}{Tokens derived from existing words}
+A small set of words from the unmodified corpus is chosen uniformly at random from the vocabulary.
 In order that the introduced tokens do not have too low a frequency, only words which occur at least 10 thousand times are chosen.
 To this set of words we add the word \word{the}, in order to include a high-frequency stopword.
-The replacement procedure is performed for each of these words, using a geometric rate of decay of $\lambda = 0.5$, and maximum value $n=20$.
+The replacement procedure of subsection \ref{replacement-procedure} is then performed for each of these words, using a geometric rate of decay of $\lambda = 0.5$, and maximum value $n=20$.
+This value of $\lambda$ is one of a range of values that ensure that, for each word, multiple tokens will be introduced with a number of occurrences sufficient to survive the chosen word2vec minimum occurrence cut-off of 200.
+A maximum value $n=20$ suffices for this choice of $\lambda$, since $2^{20 + \log_2{200}}$ exceeds the maximum occurrence of any token in the corpus. 
+Figure \ref{fig:word-frequency-experiment-text-cat} illustrates the effect of these modifications on a sample text, with a family of tokens \word{CAT\_i}, derived from the word \word{cat}.
 
-Secondly, in order to have a meaningless word for comparison, an entirely new word \word{VOID} is interspersed uniformly at random throughout the corpus so that its global frequency is $0.005$.
-The replacement procedure is then performed for this word with the same values for $\lambda$ and $n$.
+Figure \ref{fig:word-frequency-counts} gives the number of occurrences of the words chosen for this experiment.
 
-Figure \ref{fig:word-frequency-experiment-text} shows the effect of these modifications on sample text, in the case where only the word \word{cat} is chosen.
+\begin{figure}\label{fig:word-frequency-counts}
+	\input{word-frequency-experiment-counts.tex}
+	\caption{Occurrence counts for words chosen for the word frequency experiment. }
+\end{figure}
 
-\subsection{Co-occurrence noise variation experiment}\label{CNVE}
+\begin{figure}
+	\input{word-frequency-experiment-text-cat.tex}
+	\caption{Text modified for the word frequency experiment, where the word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
+	\label{fig:word-frequency-experiment-text-cat}
+\end{figure}
+
+\end{subsection}
+
+\begin{subsection}{Tokens derived from an introduced, meaningless word}
+	A high-frequency, purely meaningless word is included in the experiment for comparison.
+	We choose to introduce a new, entirely meaningless token \word{VOID} into the corpus, rather than choose an existing word whose meaninglessness is only supposed.
+	To achieve this, the token is interspersed uniformly at random throughout the corpus so that its global frequency is $0.005$.
+	The co-occurrence distribution of this introduced token thus coincides with the global frequency distribution.
+	The replacement procedure is then performed for \word{VOID}, using the same values for $\lambda$ and $n$ as above.
+Figure \ref{fig:word-frequency-experiment-text-void} illustrates the effect of this modifications on a sample text.
+
+\begin{figure}\label{fig:word-frequency-experiment-text-void}
+	\input{word-frequency-experiment-text-void.tex}
+	\caption{Text modified for the word frequency experiment, where the
+	word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
+\end{figure}
+\end{subsection}
+
+\subsection{Experimental results}
+Word frequency has no effect on word vector direction.
+Figure \ref{word-frequency-experiment-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus.
+The cosine similarity measures the extent to which two vectors have the same direction, taking a maximum value of $1$ and a minimum value of $0$.
+Notice that the word vectors associated to tokens derived from meaningful tokens have the same direction.
+The same is true of the word vectors associated to the \word{VOID\_i} for $i$ \textcolor{red}{below $7$ or $8$}, but that the word vectors \word{VOID\_i} gradually diverge directionally for higher values of $i$.
+This demonstrates the difficulty of learning the co-occurrence distribution of $\word{VOID}$, the global frequency distribution.
+Because of its noisiness, a higher number of samples (occurrences) is required to learn this co-occurrence distribution.
+
+It remains, therefore, to consider the effect of frequency variation on word vector length.
+Figure \ref{fig:frequency-norm-graph} illustrates the global relationship between vector length and word frequency.
+Figure \ref{fig:word-frequency-experiment-graph}, on the other hand, shows this relationship for individual words.
+Each line corresponds to a single word, and the points on each line indicate the frequency and vector length for the tokens derived from the corresponding word.
+For example, the five points on the line corresponding to \word{admiral} are labelled, from right to left, by the tokens \word{ADMIRAL\_1}, \word{ADMIRAL\_2}, \dots, \word{ADMIRAL\_5}.
+The number of points on the line is determined by the frequency of the original word.
+For example, the frequency of the \word{admiral} can be halved at most $5$ times and remain about the minimum frequency cut-off.
+
+Because all the points on a line share a co-occurrence distribution (only the frequency varies), this figure demonstrates conclusively that length does indeed depend on frequency directly, and not only via some correlation between word frequency and semantic content.
+Notice, moreover, that the relative positions of the norms of the word vectors associated to the test words is independent of the frequency band.
+Therefore these lines trace lines along which a frequency-adjusted norm would be constant.
+
+\begin{figure}
+	\includegraphics[scale=0.5]{word-frequency-experiment-heatmap}
+	\caption{
+	Heatmap of the dot product of the normalised vectors associated to the
+	tokens introduced into the corpus in the co-occurrence noise experiment
+	(four such words chosen at random).  The largely red blocks demonstrate
+	that the direction of the word vector does not change when noise is
+	added to the co-occurrence distribution.
+	}
+	\label{fig:cooccurrence-noise-heatmap}
+\end{figure}
+
+\begin{sidewaysfigure*}\label{fig:word-frequency-experiment-graph}
+	\centering{\includegraphics[scale=0.6]{word-frequency-experiment-graph}}
+	\caption{
+	Vector length vs. frequency for the words chosen for the word frequency
+	experiment.  For each word, tokens of varying frequency but with the
+	co-occurrence distribution of that word were introduced into the
+	corpus, as described in section \ref{WFVE}.
+	}
+\end{sidewaysfigure*}
+
+\end{section}
+
+\begin{section}{Co-occurrence noise variation experiment}\label{CNVE}
 Here, and throughout, the noise distribution is taken to be the global word frequency distribution.
 Thus noise can be added to the cooccurrence distribution of a word by inspersing occurrences of that word uniformly at randomly throughout the corpus.
-A small set of words is chosen from the unmodified corpus in the same manner as subsection \ref{WFVE}.
-In order that the two experiments can be carried out simulataneously, the two sets of words are chosen to be disjoint.
+A small set of words is chosen from the unmodified corpus in the same manner as section \ref{WFVE}.
 
-For each of the chosen words, the replacement procedure is performed using a geometric rate of decay of $\lambda = 5/6$, and truncating at $n=8$.
+For each of the chosen words, the replacement procedure of subsection \ref{replacement-procedure} is performed using a geometric rate of decay of $\lambda = 5/6$, and truncating at $n=8$.
 Subsequently, for every replacement token (e.g.
@@
 
-\begin{figure*}
-	%\input{../outputs/word-frequency-experiment-counts.tex}
-\begin{center}
-\begin{tabular}{c | c}
-word & \# occurrences \\
-\hline
-\word{developments} & 14347 \\
-\word{specific} & 77919 \\
-\word{across} & 101617 \\
-\word{path} & 24898 \\
-\word{file} & 31184 \\
-\word{from} & 2611891 \\
-\word{admiral} & 13122 \\
-\word{eyes} & 27621 \\
-\end{tabular}
-\end{center}
-\label{fig:word-frequency-counts}
-\caption{Occurrence counts for words chosen for the word frequency experiment. }
-\end{figure*}
-
-\begin{figure*}
-	%\input{../outputs/word-frequency-experiment-counts.tex}
-\begin{center}
-\begin{tabular}{c | c}
-word & \# occurrences \\
-\hline
-\word{sun} & 49723 \\
-\word{jets} & 10520 \\
-\word{holding} & 31744 \\
-\word{incorporated} & 24964 \\
-\word{initiative} & 15635 \\
-\word{jury} & 14575 \\
-\word{afternoon} & 10537 \\
-\word{plants} & 42808 \\
-\end{tabular}
-\end{center}
+\begin{figure}
+	\input{noise-cooccurrence-experiment-counts.tex}
 	\label{fig:cooccurrence-noise-counts}
 	\caption{Occurrence counts for words chosen for the cooccurrence noise experiment. }
-\end{figure*}
-
-% text after modifications for word frequency variation experiment for the word cat
-\begin{figure}[t]
-\texttt{
-a pedigreed \textcolor{red}{CAT\_2} is one whose ancestry is recorded by a \textcolor{red}{CAT\_5} fancier organization
-a purebred \textcolor{red}{CAT\_7} is one whose ancestry contains only individuals of the same \textcolor{red}{VOID\_4} breed
-the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_1} is not owned by \textcolor{red}{VOID\_7} any one individual
-the domestic \textcolor{red}{CAT\_1} was first classified as felis catus
-the domesticated \textcolor{red}{CAT\_1} and its closest wild ancestor are both diploid organisms that possess  chromosomes and roughly  genes
-the \textcolor{red}{CAT\_5} skull is unusual among mammals in having very large eye sockets and a powerful and specialized jaw
-when it overpowers its prey a \textcolor{red}{CAT\_3} delivers a lethal neck bite with its two long canine teeth
-another unusual feature is that the \textcolor{red}{CAT\_1} cannot produce taurine
-this is partly the result of \textcolor{red}{CAT\_1} eyes having a tapetum lucidum which reflects any light that passes through the retina back into the eye
-although wildcats are solitary the social behavior of domestic cats is much more variable and ranges from widely dispersed individuals to feral \textcolor{red}{CAT\_1} colonies that form around a food source based on groups of cooperating females
-within such groups one \textcolor{red}{CAT\_1} is usually dominant over the others
-each \textcolor{red}{CAT\_1} in a colony holds a distinct territory with sexually active males having the largest territories which are about  times larger than those of female cats and may overlap with several females territories
-these territories are marked by urine spraying by rubbing objects at head height with secretions from facial glands and by defecation
-between these territories are neutral areas where cats watch and greet one another without territorial conflicts
-outside these neutral areas territory holders usually chase away stranger \textcolor{red}{VOID\_2} cats at first by staring hissing and growling and if that does not work by short but noisy and violent attacks
-despite some cats cohabiting in colonies they do not have a social survival strategy or a pack mentality and always hunt alone
-the domestic dog canis lupus familiaris or canis familiaris is a domesticated canid which has been selectively bred for millennia for various behaviors sensory capabilities and physical attributes
-although initially thought to have originated as a manmade variant of an extant canid species variously supposed as being the dhole golden jackal or gray wolf extensive genetic studies undertaken during the s indicate that dogs diverged from a now-extinct canid in eurasia  years ago being the oldest domesticated animals their long association with people has allowed dogs to be uniquely attuned to human behavior as well as thrive on a starch-rich diet which would be inadequate for other canid species
-dogs perform many roles for people such as hunting herding pulling loads protection assisting police and military companionship and more recently aiding handicapped individuals in  carl linnaeus listed among the types of quadrupeds familiar to him the latin word for \textcolor{red}{VOID\_3} dog canis among the species within this genus linnaeus listed the red fox as canis vulpes wolves canis lupus and the domestic dog canis canis in later editions linnaeus dropped canis canis and greatly expanded his list of the canis genus of quadrupeds and by  included alongside the foxes wolves and jackals and many more terms that are now listed as synonyms for domestic dog including aegyptius hairless dog aquaticus water dog and mustelinus literally badger dog among these were two that later experts have been widely used for domestic dogs as a species: canis domesticus and most predominantly canis familiaris the common or familiar dog
-}
-\caption{Text modified for the word frequency experiment, where the word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
-\label{fig:word-frequency-experiment-text}
 \end{figure}
@@
 % original text after modifications for the noise experiment with n=3 lambda 5/ 6 for cat
-\begin{figure}[t]
-	\texttt{
-a pedigreed \textcolor{red}{CAT\_1} is \textcolor{red}{CAT\_1} one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization
-a purebred \textcolor{red}{CAT\_3} is one whose ancestry contains only individuals of the same breed
-the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_3} is not owned by any one individual
-\textcolor{red}{CAT\_1} the domestic \textcolor{red}{CAT\_3} was first classified as felis catus
-the domesticated \textcolor{red}{CAT\_1} and its closest wild ancestor are both diploid organisms \textcolor{red}{CAT\_3} that possess  chromosomes and roughly  genes
-the \textcolor{red}{CAT\_2} skull is unusual among mammals in having very large eye sockets and a powerful and specialized jaw
-when it overpowers its \textcolor{red}{CAT\_2} prey a \textcolor{red}{CAT\_1} delivers a lethal neck bite with \textcolor{red}{CAT\_3} its two \textcolor{red}{CAT\_2} long canine teeth
-another unusual feature \textcolor{red}{CAT\_1} is that the \textcolor{red}{CAT\_2} cannot produce taurine
-this is partly the result of \textcolor{red}{CAT\_3} eyes having a tapetum lucidum which reflects any light that passes through the retina back into the eye
-although wildcats are solitary the social behavior of domestic cats is much more variable and \textcolor{red}{CAT\_1} ranges from widely dispersed individuals to feral \textcolor{red}{CAT\_2} colonies that form around a food source based on groups of cooperating females
-within such groups one \textcolor{red}{CAT\_2} is usually dominant over the others
-each \textcolor{red}{CAT\_1} in \textcolor{red}{CAT\_3} a colony holds a distinct \textcolor{red}{CAT\_3} territory with sexually active males having the largest territories which are about  times larger than those of female cats and may overlap with several females territories
-these territories \textcolor{red}{CAT\_3} are marked by urine spraying by rubbing objects \textcolor{red}{CAT\_1} at head height with secretions from facial glands and by defecation
-between these territories are neutral areas where cats \textcolor{red}{CAT\_1} watch and greet \textcolor{red}{CAT\_2} one another without territorial conflicts
-\textcolor{red}{CAT\_1} outside these neutral areas territory holders usually chase away stranger cats at first by staring hissing and growling and if that does not work by short but \textcolor{red}{CAT\_2} noisy and violent attacks
-despite some cats cohabiting in colonies they do \textcolor{red}{CAT\_3} not have a social survival strategy \textcolor{red}{CAT\_2} or a pack mentality and always hunt alone
-the domestic dog canis lupus familiaris or canis familiaris is a domesticated canid which has been selectively bred for millennia for various behaviors sensory capabilities and physical attributes
-although initially thought to have originated as a manmade variant of an extant canid species variously supposed as being the \textcolor{red}{CAT\_1} dhole golden jackal or gray wolf \textcolor{red}{CAT\_3} extensive genetic studies undertaken during the s indicate that dogs \textcolor{red}{CAT\_1} diverged from a now-extinct canid in eurasia  years ago being the oldest domesticated animals their long association with people has \textcolor{red}{CAT\_3} allowed dogs to be \textcolor{red}{CAT\_1} uniquely attuned to human behavior as well as thrive on \textcolor{red}{CAT\_2} a starch-rich diet which would be inadequate \textcolor{red}{CAT\_2} for other \textcolor{red}{CAT\_1} canid species
-dogs perform many roles for people such \textcolor{red}{CAT\_3} as hunting herding pulling loads protection assisting \textcolor{red}{CAT\_2} police and military companionship and more recently aiding handicapped \textcolor{red}{CAT\_2} individuals in  carl linnaeus listed among the types of quadrupeds familiar to him the latin word for \textcolor{red}{CAT\_1} dog canis among \textcolor{red}{CAT\_1} the species within this genus linnaeus listed the red fox as canis vulpes wolves canis lupus and the domestic dog canis canis in later editions linnaeus dropped canis canis and greatly expanded his list of the canis genus of quadrupeds and by  included alongside the foxes wolves and jackals and \textcolor{red}{CAT\_2} many more terms that are now listed as synonyms for domestic dog including aegyptius \textcolor{red}{CAT\_1} hairless dog aquaticus water dog and mustelinus literally badger dog among these were two that later experts have been widely used for domestic dogs as a species: canis domesticus and most predominantly \textcolor{red}{CAT\_2} canis familiaris the common or familiar dog
-}
+\begin{figure}
+	\input{cooccurrence-noise-experiment.tex}
 	\caption{Text modified for the co-occurrence noise experiment, where the word \word{cat} was chosen, $\lambda = 5/6$ and $n=3$.}
@@
 
-\section{Results}
-
-\subsection{Word frequency experiment}
-Figure \ref{fig:frequency-norm-graph} illustrates the complex global relationship between vector length and word frequency.
-Figure \ref{fig:word-frequency-experiment-graph}, on the other hand, shows this relationship for individual words.
-Each line corresponds to a single word, and the points on each line indicate the frequency and vector length for the tokens derived from the corresponding word.
-For example, the five points on the line corresponding to \word{admiral} are labelled, from right to left, by the tokens \word{ADMIRAL\_1}, \word{ADMIRAL\_2}, \dots, \word{ADMIRAL\_5}.
-The number of points on the line is determined by the frequency of the original word.
-For example, the frequency of the \word{admiral} can be halved at most $5$ times and remain about the minimum frequency cut-off of word2vec.
-
-Because all the points on a line share a co-occurrence distribution (only the frequency varies), this figure demonstrates conclusively that length does indeed depend on frequency directly, and not only via some correlation between word frequency and semantic content.
-Notice, moreover, that the relative positions of the norms of the word vectors associated to the test words is independent of the frequency band.
-Therefore these lines trace lines along which a frequency-adjusted norm would be constant.
-
-\begin{figure*}
-\includegraphics[scale=0.5]{occurrence-histogram}
-\caption{
-	Word frequency histogram.
-	Here and elsewhere, frequency bands are delimited by powers of $2$.
-}
-\label{fig:frequency-histogram}
-\end{figure*}
-
-\begin{figure*}
-\includegraphics[scale=0.3]{frequency-norm-scatterplot}
-\caption{ The global relationship between word frequency and vector length.  }
-\label{fig:frequency-norm-graph}
-\end{figure*}
-
-\begin{figure*}
-\includegraphics[scale=0.5]{word-frequency-experiment-heatmap}
-\caption{
-Heatmap of the dot product of the normalised vectors associated to the tokens introduced into the corpus in the co-occurrence noise experiment (four such words chosen at random).
-The largely red blocks demonstrate that the direction of the word vector does not change when noise is added to the co-occurrence distribution.
-}
-
-\label{fig:cooccurrence-noise-heatmap}
-\end{figure*}
-\begin{sidewaysfigure*}
-	\centering{\includegraphics[scale=0.6]{word-frequency-experiment-graph}}
-\caption{
-	Vector length vs. frequency for the words chosen for the word frequency experiment.
-	For each word, tokens of varying frequency but with the co-occurrence distribution of that word were introduced into the corpus, as described in section \ref{WFVE}.
-}
-\label{fig:word-frequency-experiment-graph}
-\end{sidewaysfigure*}
-
-\subsection{Co-occurrence noise experiment}
+\begin{subsection}{Results}
 Consider the effect on a word vector when increasing levels of noise added to the co-occurrence distribution of the word.
@@
 
-\begin{sidewaysfigure*}
+\begin{figure}
+	\includegraphics[scale=0.5]{cooccurrence-noise-heatmap}
+	\caption{
+	Heatmap of the dot product of the normalised vectors associated to the
+	tokens introduced into the corpus in the co-occurrence noise experiment
+	(four such words chosen at random).  The largely red blocks demonstrate
+	that the direction of the word vector does not change when noise is
+	added to the co-occurrence distribution.
+	}
+	\label{fig:cooccurrence-noise-heatmap}
+\end{figure}
+
+\begin{sidewaysfigure*}\label{fig:cooccurrence-noise-graph}
 	\includegraphics[scale=0.6]{cooccurrence-noise-graph}
 	\caption{
-	Vector length vs. proportion of co-occurrence noise for words chosen for the co-occurrence noise experiment.
-	For each word, tokens of equal frequency but with increasing proporitions of co-occurrence noise were introduced into the corpus, as described in subsection \ref{CNVE}.
+	Vector length vs. proportion of co-occurrence noise for words
+	chosen for the co-occurrence noise experiment.  For each word,
+	tokens of equal frequency but with increasing proporitions of
+	co-occurrence noise were introduced into the corpus, as
+	described in section \ref{CNVE}.
 	}
-\label{fig:cooccurrence-noise-graph}
 \end{sidewaysfigure*}
 
-\begin{figure*}[t!]
-\includegraphics[scale=0.5]{cooccurrence-noise-heatmap}
-\caption{
-Heatmap of the dot product of the normalised vectors associated to the tokens introduced into the corpus in the co-occurrence noise experiment (four such words chosen at random).
-The largely red blocks demonstrate that the direction of the word vector does not change when noise is added to the co-occurrence distribution.
-}
-\label{fig:cooccurrence-noise-heatmap}
-\end{figure*}
+\end{subsection}
+
+\end{section}
 
@@
 
+\begin{section}{offcuts}
+Two controlled experiments are performed:
+\begin{itemize}
+	\item \emph{Word frequency variation} compare the lengths of the word vectors for words with varying frequencies, but with the same co-occurrence distribution;
+	\item \emph{Co-occurrence noise variation} compare the lengths of the word vectors for words with of the same frequency, but with varying levels of noise in their co-occurrence distributions.
+\end{itemize}
+\end{section}
+
 \footnotesize
-\bibliographystyle{unsrt}
+\bibliographystyle{te}
 \bibliography{main}
-
 \end{document}

### 2015-08-18 article/word-frequency-experiment-text-cat.tex
generate three sample texts for the article

\texttt{the domestic \textcolor{red}{CAT\_1} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_2} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_5} is one whose ancestry is recorded by a \textcolor{red}{CAT\_3} fancier organization\newline a purebred \textcolor{red}{CAT\_4} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_2} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

### 2015-08-18 article/word-frequency-experiment-text-void.tex
generate three sample texts for the article

\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

### 2015-08-18 markup_modifications.py

@@ -2,51 +2,64 @@
 Reads a sample text from stdin and writes to stdout LaTeX figures, with
 highlighting, showing the corpus modifications made for the two experiments
 with the modifications highlighted in red.
-"""
-import sys
+""" #FIXME
 import re
 from StringIO import StringIO
 from parameters import *
 from functions import *
 
-def markup(raw_output):
-    hl = re.sub(r'([A-Z]+_[0-9]+)', r'\\textcolor{red}{\1}', raw_output)
-    return re.sub('_', '\_', hl)
+def write_markup(raw_output, f_out):
+    """
+    Markup all modifications to a sample text for typesetting in LaTeX.
+    """
+    f_out.write(r"\texttt{")
+    output = re.sub(r'([A-Z]+_[0-9]+)', r'\\textcolor{red}{\1}', raw_output)
+    output = re.sub('_', '\_', output)
+    output = r'\newline '.join(output.strip().split('\n'))
+    f_out.write(output)
+    f_out.write(r"}")
 
-input_text = sys.stdin.read()
-counts = count_words(StringIO(input_text))
-total_words = sum(counts.values())
+# PERFORM THE REPLACEMENT PROCEDURE FOR THE WORD FREQUENCY EXPERIMENT FOR THE WORD "CAT"
 
-# intersperse the meaningless token throughout the corpus
-w_meaningless = StringIO()
-intersperse_words({meaningless_token: meaningless_token_frequency}, StringIO(input_text), w_meaningless)
+word = 'cat'
+word_samplers = {word: truncated_geometric_sampling(word, word_freq_experiment_ratio, word_freq_experiment_power_max)}
 
-# perform the replacement procedures for the word frequency experiment
-word_samplers = {}
-for word in ['cat', meaningless_token]:
-    word_samplers[word] = truncated_geometric_sampling(word, word_freq_experiment_ratio, word_freq_experiment_power_max)
+tmp_file = StringIO()
+with file('article/sample_input') as f_in:
+    replace_words(word_samplers, f_in, tmp_file)
 
-w_replacement = StringIO()
-replace_words(word_samplers, StringIO(w_meaningless.getvalue()), w_replacement)
+with file('article/word-frequency-experiment-text-cat.tex', 'w') as f_out:
+    write_markup(tmp_file.getvalue(), f_out)
 
-print r"""\begin{figure}[t]
-	\texttt{"""
[…]

### 2015-08-18 article/cooccurrence-noise-experiment.tex
separate files for article text samples, document frequency function

@@
-\texttt{\textcolor{red}{CAT\_3} the domestic \textcolor{red}{CAT\_3} was first classified as felis \textcolor{red}{CAT\_2} catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_3} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_2} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_2} is one \textcolor{red}{CAT\_2} whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_2} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated \textcolor{red}{CAT\_2} canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time \textcolor{red}{CAT\_3} at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}
+\texttt{the \textcolor{red}{CAT\_2} domestic \textcolor{red}{CAT\_3} was \textcolor{red}{CAT\_3} first \textcolor{red}{CAT\_2} classified as felis catus\newline the semiferal \textcolor{red}{CAT\_2} \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_3} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_3} is \textcolor{red}{CAT\_1} one whose ancestry is recorded by a \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} fancier \textcolor{red}{CAT\_1} organization\newline a purebred \textcolor{red}{CAT\_1} is one whose ancestry contains only \textcolor{red}{CAT\_1} individuals of the same breed\newline the \textcolor{red}{CAT\_2} \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye \textcolor{red}{CAT\_3} sockets\newline another unusual feature \textcolor{red}{CAT\_2} is that the \textcolor{red}{CAT\_2} \textcolor{red}{CAT\_3} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} is usually dominant \textcolor{red}{CAT\_1} over the others\newline ...\newline the domestic dog \textcolor{red}{CAT\_3} canis \textcolor{red}{CAT\_1} lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for \textcolor{red}{CAT\_3} people such as \textcolor{red}{CAT\_3} hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the \textcolor{red}{CAT\_2} time at \textcolor{red}{CAT\_1} which female dogs will have \textcolor{red}{CAT\_2} their first \textcolor{red}{CAT\_3} estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

### 2015-08-18 article/main.tex
separate files for article text samples, document frequency function

@@
 The relatively high minimum word cut-off is chosen to ensure that word vectors receive a sufficient number of gradient updates to be meaningful.
+This cut-off results in a vocabulary of $62025$ words (only unigrams were considered).
 
@@
 
-\begin{subsection}{Tokens derived from existing words}
+\begin{subsection}{Tokens derived from existing words}\label{WFVEexisting}
 A small set of words from the unmodified corpus is chosen uniformly at random from the vocabulary.
@@
 	\input{word-frequency-experiment-text-cat.tex}
-	\caption{Text modified for the word frequency experiment, where the word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
+	\caption{A text modified for the word frequency experiment as per subsection \ref{WFVEexisting}, where the
+	word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
 	\label{fig:word-frequency-experiment-text-cat}
@@
 
-\begin{subsection}{Tokens derived from an introduced, meaningless word}
+\begin{subsection}{Tokens derived from an introduced, meaningless word}\label{WFVEmeaningless}
 	A high-frequency, purely meaningless word is included in the experiment for comparison.
@@
 	\input{word-frequency-experiment-text-void.tex}
-	\caption{Text modified for the word frequency experiment, where the
-	word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
+	\caption{A text modified for the word frequency experiment as per subsection \ref{WFVEmeaningless}, where $\lambda=0.5$ and $n=20$.
+For illustrative purposes, the token \word{VOID} was here interspersed with a global frequency of $0.05$.}
 \end{figure}
@@
 
-% original text after modifications for the noise experiment with n=3 lambda 5/ 6 for cat
 \begin{figure}
 	\input{cooccurrence-noise-experiment.tex}
-	\caption{Text modified for the co-occurrence noise experiment, where the word \word{cat} was chosen, $\lambda = 5/6$ and $n=3$.}
+	\caption{A text modified for the co-occurrence noise experiment, where the word \word{cat} was chosen, $\lambda = 5/6$ and $n=3$.}
 \label{fig:cooccurrence-noise-experiment-text}

### 2015-08-18 article/word-frequency-experiment-text-cat.tex
separate files for article text samples, document frequency function

@@
-\texttt{the domestic \textcolor{red}{CAT\_1} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_2} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_5} is one whose ancestry is recorded by a \textcolor{red}{CAT\_3} fancier organization\newline a purebred \textcolor{red}{CAT\_4} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_2} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}
+\texttt{the domestic \textcolor{red}{CAT\_3} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_3} a mostly outdoor \textcolor{red}{CAT\_7} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_1} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_1} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_3} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_1} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_3} is usually dominant over the others}

### 2015-08-18 article/word-frequency-experiment-text-void.tex
separate files for article text samples, document frequency function

@@
-\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}
+\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned \textcolor{red}{VOID\_3} by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in \textcolor{red}{VOID\_1} having very large eye sockets\newline \textcolor{red}{VOID\_2} another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline \textcolor{red}{VOID\_1} dogs perform many roles for people such as hunting \textcolor{red}{VOID\_1} herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction\newline ...\newline rabbits are small mammals in the family leporidae\newline the male is called a buck and the female is a doe\newline the european rabbit has been introduced to many places around the world\newline more than half \textcolor{red}{VOID\_1} the worlds rabbit population resides in north america\newline rabbits are \textcolor{red}{VOID\_1} prey animals and are therefore constantly aware of their surroundings\newline rabbits are most active at dawn and dusk\newline as with other prey animals rabbits often sleep with their eyes open}

### 2015-08-18 article_generate_sample_texts.py

@@ -1,8 +1,7 @@
 """
-Reads a sample text from stdin and writes to stdout LaTeX figures, with
-highlighting, showing the corpus modifications made for the two experiments
-with the modifications highlighted in red.
-""" #FIXME
+Performs the corpus modifications on sample texts for inclusion in the article,
+highlighting in red the modifications (in LaTeX).
+"""
 import re
 from StringIO import StringIO
 from parameters import *
@@ -25,7 +24,7 @@ word = 'cat'
 word_samplers = {word: truncated_geometric_sampling(word, word_freq_experiment_ratio, word_freq_experiment_power_max)}
 
 tmp_file = StringIO()
-with file('article/sample_input') as f_in:
+with file('article/word-frequency-experiment-text-cat.input') as f_in:
     replace_words(word_samplers, f_in, tmp_file)
 
 with file('article/word-frequency-experiment-text-cat.tex', 'w') as f_out:
@@ -33,15 +32,14 @@ with file('article/word-frequency-experiment-text-cat.tex', 'w') as f_out:
 
 # PERFORM THE REPLACEMENT PROCEDURE FOR THE WORD FREQUENCY EXPERIMENT FOR THE MEANINGLESS WORD
 
-word_samplers = {meaningless_token: truncated_geometric_sampling(word, word_freq_experiment_ratio, word_freq_experiment_power_max)}
-
 # intersperse the meaningless token
 tmp_file1 = StringIO()
-with file('article/sample_input') as f_in:
-    intersperse_words({meaningless_token: meaningless_token_frequency}, f_in, tmp_file1)
+with file('article/word-frequency-experiment-text-void.input') as f_in:
+    intersperse_words({meaningless_token: 0.05}, f_in, tmp_file1)
 tmp_file1.seek(0)
 
 tmp_file2 = StringIO()
+word_samplers = {meaningless_token: truncated_geometric_sampling(meaningless_token, word_freq_experiment_ratio, word_freq_experiment_power_max)}
 replace_words(word_samplers, tmp_file1, tmp_file2)
 
 with file('article/word-frequency-experiment-text-void.tex', 'w') as f_out:
@@ -52,12 +50,12 @@ coocc_noise_experiment_power_max = 3 # better for the example, be sure to note i
 word = 'cat'
 word_samplers = {word: truncated_geometric_sampling(word, coocc_noise_experiment_ratio, coocc_noise_experiment_power_max)}
 
[…]

### 2015-08-18 functions.py

@@ -17,6 +17,19 @@ def count_words(f):
                 counts[word] += 1
     return counts
 
+def count_document_frequency(f, vocab):
+    """
+    Return a dictionary mapping each word to the number of documents it occurs
+    in, in the file 'f' specified.  Documents are separated from one another by
+    linefeeds.
+    Words are separated by ' '.
+    """
+    counts = {word: 0 for word in vocab}
+    for line in f:
+        for word in set(line.strip().split(' ')):
+                counts[word] += 1
+    return counts
+
 def read_words(filename):
     """
     given a CSV of word counts, returns the list of words.

### 2015-08-18 article_markup_wordcounts.py
fix silly error in article_markup_wordcounts.py

@@ -8,7 +8,7 @@ wcs = []
 for line in sys.stdin:
     word, count = line.strip().split(',')
     wcs.append((word, int(count)))
-for word, count in sorted(wcs, lambda (word, count): count):
+for word, count in sorted(wcs, key=lambda (w, c): c):
     print r'\word{%s} & %s \\' % (word, count)
 print r"""\end{tabular}
 \end{center}"""

### 2015-08-18 article/main.tex
article corrections; discussion and future directions

@@
 %\usepackage{wrapfig}
-\usepackage{natbib}
 \usepackage{rotating}
@@
 \newcommand{\word}[1]{\texttt{#1}}
-\urldef{\googlegroup}\url{https://groups.google.com/forum/#\!forum/word2vec-toolkit}
+\urldef{\googlegroup}\url{https://groups.google.com/forum/#!forum/word2vec-toolkit}
 \urldef{\blogpost}\url{https://blog.lateral.io/2015/06/the-unknown-perils-of-mining-wikipedia/}
@@
 	Adriaan Schakel\\
+	NNLP\\
 	\texttt{adriaan.schakel@gmail.com}
@@
 	This approach is demonstrated in the case of the word2vec CBOW model with experiments that independently vary word frequency and word co-occurrence noise.
+	It is shown in this case that word vector length depends directly and essentially linearly on both word frequency and the level of noise in the co-occurrence distribution of the word.
+	In both cases, the coefficient of linearity depends upon the word.
 \end{abstract}
@@
 Samples of the co-occurrence distributions are obtained by scanning a short window over the text.
-The non-central (or ``context'') words are treated as samples of the co-occurrence distribution the middle (or ``current'') word. 
-In this way, the word vectors of the word embedding are determined by the co-occurrence distributions and frequencies of the words in the vocabulary.
+The neighbouring words are treated as samples of the co-occurrence distribution the current word. 
+Viewed in this way, the word vectors of the word embedding are determined by the co-occurrence distributions and frequencies of the words in the corpus.
 
-In this short note, we perform experiments that illustrate the effect of these two determining pieces of information (co-occurrence distribution and word frequency) as they are varied independently of one another.
-This is acheived through the introduction of new ``words'' into the training corpus with varying frequencies and varying levels of noise in their co-occurrence distributions.
-The frequency and co-occurrence distributions of these introduced words are modelled on existing words in the corpus.
+In this short note, we perform experiments that illustrate the effect of these two determining pieces of information (the co-occurrence distribution and the word frequency) as they are varied independently of one another.
+This is acheived through the introduction of new tokens into the training corpus with varying frequencies and varying levels of noise in their co-occurrence distributions.
+The frequency and co-occurrence distributions of these introduced tokens are modelled on existing words in the corpus.
 
@@
 However, these experiments (and other besides) could might equally well be performed for word embedding methods such as word2vec skip-gram \cite{DistRepns,EfficientEstimation}, GloVe \cite{pennington2014glove} and SENNA \cite{collobert-2011}.
-In the case of word2vec, we show that word vector length depends directly and linearly with word frequency.
-We show furthermore that word vector length depends directly and linearly with the level of noise in the co-occurrence distribution of the word.
+
+In the case of word2vec, we show that word vector length depends directly on both word-frequency and the level of noise in the co-occurrence distribution of the word in an essentially linear manner.
 In both cases, the coefficient of linearity depends upon the word.
+If the co-occurrence distribution is fixed, then word vector length decreases with increasing word frequency.
+On the other hand, if word frequency is held constant, then word vector length decreases as the level of  noise in the co-occurrence distribution of the word is increased.
+
+We show furthermore that word vector direction does not vary with word frequency or with the level of co-occurrence noise.
+
+This short note is structured as follows.
+Section \ref{related-work} draws connections to related work, while Section \ref{corpus-and-model} describes the corpus and model used to illustrate the experiments.
+Section \ref{WFVE} describes controlled experiment for varying word frequency while holding the co-occurrence distribution fixed.
+Section \ref{CNVE}, in a complementary fashion, describes a controlled experiment for varying the level of noise in the cooccurrence distribution of a word, while holding the word frequency fixed.
+Section \ref{future-directions} considers further questions and possible future directions.
+\end{section}
+
+\begin{section}{Related work}\label{related-work}
+Our experimental finding (in the context of word2vec), that word vector length decreases with co-occurrence noise is reminiscent of \cite{vecchi-baroni-zamparelli2011}, where a relationship between vector length and the ``semantic deviance'' of an adjective-noun composition was studied observationally.
+In \cite{schakel-wilson}, the interpretation of word vector length as an indication of word significance is explored in the context of the arXiv high-energy physics corpus.
+
+\textcolor{red}{what about some work introducing tokens into a corpus for experiments?}
 
@@
 Word vector length is thus disregarded.
-In the word2vec forum\footnote{\googlegroup}, it has been observed that normalisation improves performance on theses tasks since word vector length is related to word frequency.
+In the word2vec forum\footnote{\googlegroup}, it has been observed that normalisation improves performance on these tasks since word vector length is related to word frequency.
 Figure \ref{fig:frequency-norm-graph} illustrates the global relationship between vector length and word frequency.
@@
 
-\begin{subsection}{Previous work}
-Our experimental finding, in the context of word2vec, that word vector length decreases with co-occurrence noise is reminiscent of \cite{vecchi-baroni-zamparelli2011}, where a relationship between vector length and the ``semantic deviance'' of an adjective-noun composition was studied observationally.
-In \cite{schakel-wilson}, the interpretation of word vector length is an indication of word significance is explored in the context of the arXiv high-energy physics corpus.
-\end{subsection}
-
-\begin{figure}
-	\includegraphics[scale=0.3]{frequency-norm-scatterplot}
-	\caption{ The global relationship between word frequency and vector length.  }
-	\label{fig:frequency-norm-graph}
-\end{figure}
+On the other hand, recent theoretical work \cite{Arora2015} has approached the problem of explaining the so-called ``compositionality'' property exhibited by some word embeddings and that has been the subject of much recent attention.
+In that work, unnormalised vectors are used in their model of the word relationship task.
+It is hoped that experimental approaches such as those described here might yield insights into the role of the word vector length in the word relationship tasks.
 \end{section}
 
-\begin{section}{Corpus and model}
+\begin{section}{Corpus and model}\label{corpus-and-model}
 Our training data is built from the Wikipedia XML datadump from October 2013.
@@
 
-This base corpus is then modified as described in section \ref{experimental_design}.
+This base corpus is then modified as described in sections \ref{WFVE} and \ref{CNVE}.
 Punctuation and numbers were removed and the corpus lower-cased prior to the modifications.
@@
 
-For simplicity, only the CBOW method with a single set of hyperparameters is considered.
-Specifically, a $100$-dimensional CBOW model is trained using negative sampling with 5 negative samples, a window size of $10$, a minimum word occurrence of $200$, and $10$ passes through the corpus.
-Subsampling was not used so that influence of word frequency could be more clearly discerned.
+For simplicity, only the word2vec CBOW word embedding with a single set of hyperparameters is considered.
+Specifically, a $100$-dimensional CBOW model is trained using negative sampling with 5 negative samples, a window size of $10$, a minimum word occurrence of $128$, and $10$ passes through the corpus.
+Subsampling was not used so that the influence of word frequency could be more clearly discerned.
 Identical experimental results were obtained using hierarchical softmax, but these are omitted for succinctness.
 The relatively high minimum word cut-off is chosen to ensure that word vectors receive a sufficient number of gradient updates to be meaningful.
-This cut-off results in a vocabulary of $62025$ words (only unigrams were considered).
+\textcolor{red}{This cut-off results in a vocabulary of $62025$ words (only unigrams were considered).}
 
@@
 
+Word vectors are then derived from the first layer of synaptic weights (``syn0'') and the second layer of weights (``syn1neg'') are discarded.
+
 \begin{figure}\label{fig:frequency-histogram}
@@
 	\caption{
-	Word frequency histogram.
-	Here and elsewhere, frequency bands are delimited by powers of $2$.
+	Number of vocabulary words by frequency band.
 	}
@@
 The tokens in each family vary in frequency but share a common co-occurrence distribution.
-After model training, comparison of the resulting word vectors for these tokens reveals the independent effect of frequency variation on the word embedding.
+After model training, comparison of the resulting word vectors associated to these tokens reveals the independent effect of frequency variation on the word embedding.
 
@@
 In order that the introduced tokens do not have too low a frequency, only words which occur at least 10 thousand times are chosen.
-To this set of words we add the word \word{the}, in order to include a high-frequency stopword.
-The replacement procedure of subsection \ref{replacement-procedure} is then performed for each of these words, using a geometric rate of decay of $\lambda = 0.5$, and maximum value $n=20$.
-This value of $\lambda$ is one of a range of values that ensure that, for each word, multiple tokens will be introduced with a number of occurrences sufficient to survive the chosen word2vec minimum occurrence cut-off of 200.
-A maximum value $n=20$ suffices for this choice of $\lambda$, since $2^{20 + \log_2{200}}$ exceeds the maximum occurrence of any token in the corpus. 
+To this set of words, the word \word{the} is added, in order to include a high-frequency stopword.
+The replacement procedure of subsection \ref{replacement-procedure} is then performed for each of these words, using a geometric rate of decay of $\lambda = 1/2$, and maximum value $n=20$.
+This value of $\lambda$ is one of a range of values that ensure that, for each word, multiple tokens will be introduced with a number of occurrences sufficient to survive the chosen word2vec minimum occurrence cut-off of 128.
+A maximum value $n=20$ suffices for this choice of $\lambda$, since $2^{20 + \log_2{128}}$ exceeds the maximum occurrence of any token in the corpus. 
 Figure \ref{fig:word-frequency-experiment-text-cat} illustrates the effect of these modifications on a sample text, with a family of tokens \word{CAT\_i}, derived from the word \word{cat}.
+Notice that the word \word{cat} has been replaced throughout with the tokens \word{CAT\_i}.
 
@@
 	We choose to introduce a new, entirely meaningless token \word{VOID} into the corpus, rather than choose an existing word whose meaninglessness is only supposed.
-	To achieve this, the token is interspersed uniformly at random throughout the corpus so that its global frequency is $0.005$.
-	The co-occurrence distribution of this introduced token thus coincides with the global frequency distribution.
+	To achieve this, the token is interspersed uniformly at random throughout the corpus so that its global frequency is $f = 0.005$.
+	The co-occurrence distribution of \word{VOID} thus coincides with the global frequency distribution.
 	The replacement procedure is then performed for \word{VOID}, using the same values for $\lambda$ and $n$ as above.
-Figure \ref{fig:word-frequency-experiment-text-void} illustrates the effect of this modifications on a sample text.
+	Figure \ref{fig:word-frequency-experiment-text-void} shows the effect of this modifications on a sample text ($f = 0.05$ is used there for illustrative purposes).
 
@@
 Because of its noisiness, a higher number of samples (occurrences) is required to learn this co-occurrence distribution.
+The same holds, but to a lesser extent, for the co-occurrence distribution of the stopword \word{the}.
 
@@
 Figure \ref{fig:frequency-norm-graph} illustrates the global relationship between vector length and word frequency.
-Figure \ref{fig:word-frequency-experiment-graph}, on the other hand, shows this relationship for individual words.
+Figure \ref{fig:word-frequency-experiment-graph}, on the other hand, shows this relationship for individual words, both for the word vectors, ``syn0'' and for the discarded second layer vectors ``syn1neg'' (included for completeness).
 Each line corresponds to a single word, and the points on each line indicate the frequency and vector length for the tokens derived from the corresponding word.
-For example, the five points on the line corresponding to \word{admiral} are labelled, from right to left, by the tokens \word{ADMIRAL\_1}, \word{ADMIRAL\_2}, \dots, \word{ADMIRAL\_5}.
+\textcolor{red}{For example, the five points on the line corresponding to \word{admiral} are labelled, from right to left, by the tokens \word{ADMIRAL\_1}, \word{ADMIRAL\_2}, \dots, \word{ADMIRAL\_5}.}
 The number of points on the line is determined by the frequency of the original word.
-For example, the frequency of the \word{admiral} can be halved at most $5$ times and remain about the minimum frequency cut-off.
+\textcolor{red}{For example, the frequency of the \word{admiral} can be halved at most $5$ times and remain about the minimum frequency cut-off.}
 
-Because all the points on a line share a co-occurrence distribution (only the frequency varies), this figure demonstrates conclusively that length does indeed depend on frequency directly, and not only via some correlation between word frequency and semantic content.
-Notice, moreover, that the relative positions of the norms of the word vectors associated to the test words is independent of the frequency band.
+Because all the points on a line share a co-occurrence distribution (only the frequency varies), this figure demonstrates conclusively that length does indeed depend on frequency directly, and not only via some correlation between word frequency and word significance.
+Moreover, this relationship is essentially linear for any particular word.
+Notice also that the relative positions of the norms of the word vectors associated to the test words is independent of the frequency band.
 Therefore these lines trace lines along which a frequency-adjusted norm would be constant.
@@
 \begin{figure}
+	\includegraphics[scale=0.3]{frequency-norm-scatterplot}
+	\caption{ The global relationship between word frequency and vector length.  }
+	\label{fig:frequency-norm-graph}
+\end{figure}
+
+\begin{figure}\label{word-frequency-experiment-heatmap}
 	\includegraphics[scale=0.5]{word-frequency-experiment-heatmap}
 	\caption{
-	Heatmap of the dot product of the normalised vectors associated to the
+	Heatmap of the cosine similarity of the vectors associated to the
 	tokens introduced into the corpus in the co-occurrence noise experiment
-	(four such words chosen at random).  The largely red blocks demonstrate
-	that the direction of the word vector does not change when noise is
-	added to the co-occurrence distribution.
+	(words other than \word{the} and \word{VOID} chosen randomly).
 	}
-	\label{fig:cooccurrence-noise-heatmap}
 \end{figure}
@@
 	corpus, as described in section \ref{WFVE}.
+	The word vectors are obtained from the first synaptic layer, ``syn0''.
+	The second layer ``syn1neg'' is included for completeness.
 	}
@@
 Here, and throughout, the noise distribution is taken to be the global word frequency distribution.
-Thus noise can be added to the cooccurrence distribution of a word by inspersing occurrences of that word uniformly at randomly throughout the corpus.
-A small set of words is chosen from the unmodified corpus in the same manner as section \ref{WFVE}.
+Thus noise can be added to the co-occurrence distribution of a word by inspersing occurrences of that word uniformly at randomly throughout the corpus.
+A small set of words is chosen from the unmodified corpus in the same manner as in section \ref{WFVE}.
 
@@
 \word{CAT\_i}) random occurrences of this token are interspersed uniformly at random throughout the corpus, such that the frequency of the replacement token is restored to that of the original token \word{cat}.
-For example, if the original token \word{cat} occurred $1000$ times, then after the replacement procedure, \word{CAT\_2} occurs approximately $694$ times, so a further approximately $306 (=1000 - 694)$ random occurrences of \word{CAT\_2} are interspersed throughout the corpus.
+For example, if the original token \word{cat} occurred $1000$ times, then after the replacement procedure, \word{CAT\_2} occurs approximately $694$ times, so a further (approximately) $306$ random occurrences of \word{CAT\_2} are interspersed throughout the corpus.
 Thus token \word{cat} is removed from the corpus, and a family of tokens \word{CAT\_i}, $1 \leqslant i \leqslant 8$ are introduced.
-These tokens have the same frequency as \word{cat}, but their cooccurrence distributions, while based on that of \word{cat}, have an increasing amount of noise.
-To be precise, the proportion of the occurrences of \word{CAT\_i} that arose by replacing \word{cat} is $P_{5/6}(i)$, and the proportion arising from its random intersperal throughout the corpus $(1 - P_{5/6}(i))$.
+These tokens all have the same frequency as \word{cat}, but their co-occurrence distributions, while based on that of \word{cat}, have an increasing amount of noise.
 
-Figure \ref{fig:cooccurrence-noise-experiment-text} illustrates the effect of this modification, independently of the modifications of subsection \ref{WFVE}, in the case where the only word chosen is \word{cat}.
-Note that for the experiments, \emph{both} modification procedures are performed, but with different sets of words.
+Figure \ref{fig:co-occurrence-noise-experiment-text} illustrates the effect of this modification, in the case where the only word chosen is \word{cat}.
+The original text in this case concerned both cats and dogs.
+Notice that the word \word{cat} has been replaced entirely in the cats section by occurrences of \word{CAT\_i} and moreover that occurrences of these same tokens appear also in the dogs section.
+These occurrences (and additionally, with probability, some occurrences from the cats section) are ``noise occurrences''.
+
+The proportion of the occurrences of \word{CAT\_i} that arose by replacing \word{cat} is $P_{5/6}(i)$, and the proportion arising from the random intersperal of this token throughout the corpus is $r_i := (1 - P_{5/6}(i))$.
+The particular value of $\lambda$ (large compared to that of section \ref{WFVE}) was chosen so that the proportion of noise $r_i$ did not increase too quickly with $i$.
+These proportions are the sampling points on our axis of variation, and this value of $\lambda$ (together with the choise of $n$) ensures a reasonable spread.
+Note that other parameter values (or indeed other distributions) could equally have been used.
 
@@
 	\input{noise-cooccurrence-experiment-counts.tex}
-	\label{fig:cooccurrence-noise-counts}
-	\caption{Occurrence counts for words chosen for the cooccurrence noise experiment. }
+	\label{fig:co-occurrence-noise-counts}
+	\caption{Occurrence counts for words chosen for the co-occurrence noise experiment. }
 \end{figure}
@@
 	\caption{A text modified for the co-occurrence noise experiment, where the word \word{cat} was chosen, $\lambda = 5/6$ and $n=3$.}
-\label{fig:cooccurrence-noise-experiment-text}
+\label{fig:co-occurrence-noise-experiment-text}
 \end{figure}
@@
 \begin{subsection}{Results}
-Consider the effect on a word vector when increasing levels of noise added to the co-occurrence distribution of the word.
-This is affected by randomly dispersing an increasing proporition $0 \leqslant p \leqslant 1$ of the pre-existing occurrences of that word throughout the corpus.
-The word vector depends on $p$, and thus $p$ parameterises a path in the feature space.
-This path begins (with $p=0$) at the original word vector, and ends (with $p=1$) at the vector associated with the meaningless vector \word{VOID}.
+Figure \ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus.
+Notice that the word vectors associated to tokens derived from the same word all share a common direction.
+We conclude that co-occurrence noise has no effect on word vector direction.
 
-For each word (e.g. \word{jury}) in the co-occurrence noise experiment, the vectors associated with the tokens \word{JURY\_i} mark points along this word's path for increasing values of $p$.
-Figure \ref{fig:cooccurrence-noise-heatmap} shows that these points at all co-linear.
-Thus the path is a straight line.
+On the other hand, Figure \ref{fig:co-occurrence-noise-graph} shows that word vector length varies directly and linearly with the proportion $r$ of noise in the co-occurrence distribution.
+This figure motivates an interpretation of vector length, within a sufficiently narrow frequency band, as a measure of the absence of co-occurrence noise, that is, of the extent to which a word determines a context.
 
-Figure \ref{fig:cooccurrence-noise-graph} illustrates the relationship between the proportion $p$ of co-occurrence noise and vector length.
-Remarkably, it is apparent that this relationship is linear for any individual word.
-Thus the norm decreases at a constant velocity as $p$ travels from $0$ to $1$.
-
-It is Figure \ref{fig:cooccurrence-noise-graph} that motivates the interpretation of vector length, within a sufficiently frequency band, as a measure of the absence of co-occurrence noise, that is, of the extent to which a word determines a context.
-
-\begin{figure}
+\begin{figure}\label{fig:co-occurrence-noise-heatmap}
 	\includegraphics[scale=0.5]{cooccurrence-noise-heatmap}
 	\caption{
-	Heatmap of the dot product of the normalised vectors associated to the
+	Heatmap of the cosine similarity of some of the vectors associated to the
 	tokens introduced into the corpus in the co-occurrence noise experiment
-	(four such words chosen at random).  The largely red blocks demonstrate
-	that the direction of the word vector does not change when noise is
-	added to the co-occurrence distribution.
+	(four such words chosen at random).  The
+	largely red blocks demonstrate that the direction of the word vector
+	does not change when noise is added to the co-occurrence distribution.
 	}
-	\label{fig:cooccurrence-noise-heatmap}
 \end{figure}
 
-\begin{sidewaysfigure*}\label{fig:cooccurrence-noise-graph}
+\begin{sidewaysfigure*}\label{fig:co-occurrence-noise-graph}
 	\includegraphics[scale=0.6]{cooccurrence-noise-graph}
@@
 	chosen for the co-occurrence noise experiment.  For each word,
-	tokens of equal frequency but with increasing proporitions of
+	tokens of equal frequency but with increasing proportions of
 	co-occurrence noise were introduced into the corpus, as
 	described in section \ref{CNVE}.
+	The word vectors are obtained from the first synaptic layer, ``syn0''.
+	The second layer ``syn1neg'' is included for completeness.
 	}
@@
 
-\section{Questions}
-\begin{enumerate}
-\item{ Why is the meaningless vector not at zero, and could the performance on the compositionality tasks be improved by accounting for this?
-Notice that the meaningless vector is only updated as part of the context vector.
-Would it be zero if we were performing skipgram?
-Does this have something to do with the lack of bias terms in the word2vec softmax?}
-\item{Our arguments from the results subsection on the co-occurrence noise experiment indicate that the meaningless vector should indeed be taken as the centre of the space for word similarity tasks.
-Check again whether this works experimentally.}
-\item{What is the distribution of norms for the vectors relative to the meaningless vector?
-Do I need to consider a different meaningless vector for each frequency band?
-This is the measure that should represent meaning.}
+\begin{section}{Discussion and future directions}\label{future-directions}
 
-\end{enumerate}
+\begin{subsection}{Controlled experiments for word embeddings}
+The principle contribution of this short note has been to demonstrate that controlled experiments can be used to gain insight into any word embedding.
+Because they are achieved via modification of the training corpus, these experiments can be carried out for any language model or word embedding.
+The experiments are moreover easy to carry out, requiring modification of code nor even knowledge of the model itself.
 
-\begin{section}{offcuts}
-Two controlled experiments are performed:
-\begin{itemize}
-	\item \emph{Word frequency variation} compare the lengths of the word vectors for words with varying frequencies, but with the same co-occurrence distribution;
-	\item \emph{Co-occurrence noise variation} compare the lengths of the word vectors for words with of the same frequency, but with varying levels of noise in their co-occurrence distributions.
-\end{itemize}
+More elaborate experiments could be carried out in the context of word embeddings.
+For instance, by introducing tokens into the corpus that mix, with varying proportions, the co-occurrence distributions of two words, the path between the word vectors in the feature space can be studied.
+The co-occurrence noise variation experiment described here would be a special case of such an experiment where one of the two words was \word{VOID}.
+
+It would, of course, be interesting to perform these experiments for other word embeddings and for different hyperparameters settings.
+\end{subsection}
+
+\begin{subsection}{Applications in other contexts}
+Further afield, it would be interesting to perform analogs of these experiments in other contexts where models are trained from co-occurrence data.
+For instance, in the context of e-Commerce, product purchase frequency and product co-occurrence noise can be varied by modifying the item-item matrix.
+\end{subsection}
+
+\begin{subsection}{Questions pertaining to word2vec}
+Questions pertaining to particularly to word2vec arise naturally from the results of the experiments.
+Figures \ref{fig:frequency-norm-graph} and \ref{fig:co-occurrence-noise-graph}, for example, demonstrate the word vectors obtained from the first synaptic layer, ``syn0'', have very different properties from those that could be obtained from the second layer, ``syn1neg'', which are typically discarded.
+These differences warrant further investigation.
+
+The co-occurrence distribution of \word{VOID} is the global frequency distribution, and in this sense pure background noise.
+Thus the word vector of \word{VOID} is a special point in the feature space.
+Figure \ref{fig:frequency-norm-graph} shows that this point is not at the origin of the feature space (i.e. is not the zero vector).
+It is the origin, however, that is implicitly used as the point of reference in the word2vec word similarity task.
+This raises the question of whether improved performance on the similarity task could be achieved by transforming the feature space (or modifying the model) such that the representation of pure noise (i.e.
+the vector for \word{VOID}) was at the origin of the transformed feature space.
+\end{subsection}
+
 \end{section}
 
+\clearpage
 \footnotesize
-\bibliographystyle{te}
 \bibliography{main}
+\bibliographystyle{plain}
 \end{document}

### 2015-08-18 build_images.py
say cosine similarity not dot product; the in heatmap once

@@ -82,7 +82,7 @@ def set_num_plots(num_plots):
     colormap = plt.cm.gist_ncar
     plt.gca().set_color_cycle([colormap(i) for i in np.linspace(0, 0.9, num_plots)])
 
-words = word_freq_experiment_words
+words = word_freq_experiment_words + [meaningless_token]
 fig = plt.figure(figsize=(16, 6))
 colormap = plt.cm.gist_ncar
 colorcycle = [colormap(i) for i in np.linspace(0, 0.9, len(words))]
@@ -114,7 +114,7 @@ set_num_plots(len(words))
 for word in words:
     plot_for_word(ax_syn1neg, word, stats.L2_norm_syn1neg)
 
-ax_syn0.set_ylim(0, 35)
+ax_syn0.set_ylim(0, 45)
 
 _ = fig.legend(lines, words, bbox_to_anchor=(0.76, 0.56), loc='center', fontsize=14, frameon=False)
 plt.tight_layout()
@@ -122,7 +122,8 @@ plt.tight_layout()
 plt.savefig('outputs/word-frequency-experiment-graph.eps')
 
 
-test_words = random.sample(word_freq_experiment_words, 4) + ['the', meaningless_token]
+test_words = random.sample([word for word in word_freq_experiment_words if word != 'the'], 3)
+test_words += ['the', meaningless_token]
 idxs = [build_experiment_token(word, i) for word in test_words for i in range(1, max(word_freq_experiment_ratio, word_freq_experiment_power_max) + 1)]
 test_vecs = vectors_syn0.loc[idxs].dropna()
 cosine_similarity_heatmap(test_vecs, figsize=(12, 10))

### 2015-08-18 functions.py

@@ -150,7 +150,7 @@ def cosine_similarity_heatmap(test_vecs, **kwargs):
     test_vecs_normed = test_vecs.as_matrix() / np.sqrt((test_vecs ** 2).sum(axis=1))[:,np.newaxis]
     mat = test_vecs_normed.dot(test_vecs_normed.transpose())    
     plt.figure(**kwargs)
-    plt.title('Dot product of normalised word vectors')
+    plt.title('Cosine similarity of word vectors')
     plt.pcolor(mat, vmin=-1, vmax=1)
     plt.colorbar()
     _ = plt.yticks(np.arange(0.5, len(test_vecs.index), 1), test_vecs.index)

### 2015-08-18 article/main.tex
spell checked

@@
-% compile using latex (xelatex seems to crop the figure strangley)
+% compile using latex (xelatex seems to crop the figure strangely)
 \documentclass{article} % For LaTeX2e
@@
 \usepackage{hyperref}
+\usepackage{mdframed}
 %\usepackage{wrapfig}
@@
 	We introduce an experimental approach to studying the properties of word embeddings.
-	Controlled experiments, acheived by modification of the training corpus, permit the demonstration of direct, causal relationships between word properties and word vector direction and length.
+	Controlled experiments, achieved by modification of the training corpus, permit the demonstration of direct, causal relationships between word properties and word vector direction and length.
 	This approach is demonstrated in the case of the word2vec CBOW model with experiments that independently vary word frequency and word co-occurrence noise.
@@
 In this short note, we perform experiments that illustrate the effect of these two determining pieces of information (the co-occurrence distribution and the word frequency) as they are varied independently of one another.
-This is acheived through the introduction of new tokens into the training corpus with varying frequencies and varying levels of noise in their co-occurrence distributions.
-The frequency and co-occurrence distributions of these introduced tokens are modelled on existing words in the corpus.
+This is achieved through the introduction of new tokens into the training corpus with varying frequencies and varying levels of noise in their co-occurrence distributions.
+The frequency and co-occurrence distributions of these introduced tokens are modeled on existing words in the corpus.
 
@@
 Section \ref{WFVE} describes controlled experiment for varying word frequency while holding the co-occurrence distribution fixed.
-Section \ref{CNVE}, in a complementary fashion, describes a controlled experiment for varying the level of noise in the cooccurrence distribution of a word, while holding the word frequency fixed.
+Section \ref{CNVE}, in a complementary fashion, describes a controlled experiment for varying the level of noise in the co-occurrence distribution of a word, while holding the word frequency fixed.
 Section \ref{future-directions} considers further questions and possible future directions.
@@
 Specifically, a $100$-dimensional CBOW model is trained using negative sampling with 5 negative samples, a window size of $10$, a minimum word occurrence of $128$, and $10$ passes through the corpus.
-Subsampling was not used so that the influence of word frequency could be more clearly discerned.
+Sub-sampling was not used so that the influence of word frequency could be more clearly discerned.
 Identical experimental results were obtained using hierarchical softmax, but these are omitted for succinctness.
@@
 
-\begin{figure}
+\begin{figure}\label{fig:word-frequency-experiment-text-cat}
+	\begin{mdframed}
 	\input{word-frequency-experiment-text-cat.tex}
+	\end{mdframed}
 	\caption{A text modified for the word frequency experiment as per subsection \ref{WFVEexisting}, where the
 	word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
-	\label{fig:word-frequency-experiment-text-cat}
 \end{figure}
@@
 \begin{figure}\label{fig:word-frequency-experiment-text-void}
+	\begin{mdframed}
 	\input{word-frequency-experiment-text-void.tex}
+	\end{mdframed}
 	\caption{A text modified for the word frequency experiment as per subsection \ref{WFVEmeaningless}, where $\lambda=0.5$ and $n=20$.
@@
 Figure \ref{word-frequency-experiment-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus.
-The cosine similarity measures the extent to which two vectors have the same direction, taking a maximum value of $1$ and a minimum value of $0$.
+The cosine similarity measures the extent to which two vectors have the same direction, taking a maximum value of $1$ and a minimum value of $-1$.
 Notice that the word vectors associated to tokens derived from meaningful tokens have the same direction.
@@
 Each line corresponds to a single word, and the points on each line indicate the frequency and vector length for the tokens derived from the corresponding word.
-\textcolor{red}{For example, the five points on the line corresponding to \word{admiral} are labelled, from right to left, by the tokens \word{ADMIRAL\_1}, \word{ADMIRAL\_2}, \dots, \word{ADMIRAL\_5}.}
+\textcolor{red}{For example, the five points on the line corresponding to \word{admiral} are labeled, from right to left, by the tokens \word{ADMIRAL\_1}, \word{ADMIRAL\_2}, \dots, \word{ADMIRAL\_5}.}
 The number of points on the line is determined by the frequency of the original word.
@@
 Here, and throughout, the noise distribution is taken to be the global word frequency distribution.
-Thus noise can be added to the co-occurrence distribution of a word by inspersing occurrences of that word uniformly at randomly throughout the corpus.
+Thus noise can be added to the co-occurrence distribution of a word by interspersing occurrences of that word uniformly at randomly throughout the corpus.
 A small set of words is chosen from the unmodified corpus in the same manner as in section \ref{WFVE}.
@@
 
-The proportion of the occurrences of \word{CAT\_i} that arose by replacing \word{cat} is $P_{5/6}(i)$, and the proportion arising from the random intersperal of this token throughout the corpus is $r_i := (1 - P_{5/6}(i))$.
+The proportion of the occurrences of \word{CAT\_i} that arose by replacing \word{cat} is $P_{5/6, 8}(i)$, and the proportion arising from the random interspersal of this token throughout the corpus is $r_i := (1 - P_{5/6, 8}(i))$.
 The particular value of $\lambda$ (large compared to that of section \ref{WFVE}) was chosen so that the proportion of noise $r_i$ did not increase too quickly with $i$.
-These proportions are the sampling points on our axis of variation, and this value of $\lambda$ (together with the choise of $n$) ensures a reasonable spread.
+These proportions are the sampling points on our axis of variation, and this value of $\lambda$ (together with the choice of $n$) ensures a reasonable spread.
 Note that other parameter values (or indeed other distributions) could equally have been used.
@@
 \begin{figure}
+	\begin{mdframed}
 	\input{cooccurrence-noise-experiment.tex}
+	\end{mdframed}
 	\caption{A text modified for the co-occurrence noise experiment, where the word \word{cat} was chosen, $\lambda = 5/6$ and $n=3$.}

### 2015-08-20 article/main.tex
tables not histograms; disregard poorly learned vecs; rm notebook

@@
 \usepackage{amsmath,amssymb,amsfonts,comment}
-\definecolor{LinkColor}{rgb}{0,0,0}
+%\definecolor{LinkColor}{rgb}{0,0,0}
 
@@
 	Adriaan Schakel\\
-	NNLP\\
+	\textcolor{red}{NNLP?}\\
 	\texttt{adriaan.schakel@gmail.com}
@@
 	We introduce an experimental approach to studying the properties of word embeddings.
-	Controlled experiments, achieved by modification of the training corpus, permit the demonstration of direct, causal relationships between word properties and word vector direction and length.
+	Controlled experiments, achieved through modifications of the training corpus, permit the demonstration of direct, causal relationships between word properties and word vector direction and length.
 	This approach is demonstrated in the case of the word2vec CBOW model with experiments that independently vary word frequency and word co-occurrence noise.
-	It is shown in this case that word vector length depends directly and essentially linearly on both word frequency and the level of noise in the co-occurrence distribution of the word.
+	In this case, it is shown that word vector length depends directly and linearly on both word frequency and the level of noise in the co-occurrence distribution of the word.
 	In both cases, the coefficient of linearity depends upon the word.
@@
 
-In this short note, we perform experiments that illustrate the effect of these two determining pieces of information (the co-occurrence distribution and the word frequency) as they are varied independently of one another.
+We perform experiments that illustrate the effect of these two determining pieces of information (the co-occurrence distribution and the word frequency) as they are varied independently of one another.
 This is achieved through the introduction of new tokens into the training corpus with varying frequencies and varying levels of noise in their co-occurrence distributions.
@@
 We illustrate our approach in case of the word2vec CBOW model.
-However, these experiments (and other besides) could might equally well be performed for word embedding methods such as word2vec skip-gram \cite{DistRepns,EfficientEstimation}, GloVe \cite{pennington2014glove} and SENNA \cite{collobert-2011}.
+However, these experiments (and other besides) could equally be carried out for word embedding methods such as word2vec skip-gram \cite{DistRepns,EfficientEstimation}, GloVe \cite{pennington2014glove} and SENNA \cite{collobert-2011}.
 
-In the case of word2vec, we show that word vector length depends directly on both word-frequency and the level of noise in the co-occurrence distribution of the word in an essentially linear manner.
+In the case of word2vec, we show that word vector length depends directly and linearly on both word-frequency and the level of noise in the co-occurrence distribution of the word.
 In both cases, the coefficient of linearity depends upon the word.
-If the co-occurrence distribution is fixed, then word vector length decreases with increasing word frequency.
+If the co-occurrence distribution is fixed, then word vector length increases with word frequency.
 On the other hand, if word frequency is held constant, then word vector length decreases as the level of noise in the co-occurrence distribution of the word is increased.
@@
 
-This short note is structured as follows.
-Section \ref{related-work} draws connections to related work, while Section \ref{corpus-and-model} describes the corpus and model used to illustrate the experiments.
-Section \ref{WFVE} describes controlled experiment for varying word frequency while holding the co-occurrence distribution fixed.
+This paper is structured as follows.
+Section \ref{related-work} draws connections to related work, while section \ref{corpus-and-model} describes the corpus and model used to illustrate the experiments.
+Section \ref{WFVE} describes a controlled experiment for varying word frequency while holding the co-occurrence distribution fixed.
 Section \ref{CNVE}, in a complementary fashion, describes a controlled experiment for varying the level of noise in the co-occurrence distribution of a word, while holding the word frequency fixed.
@@
 In the word2vec forum\footnote{\googlegroup}, it has been observed that normalisation improves performance on these tasks since word vector length is related to word frequency.
-Figure \ref{fig:frequency-norm-graph} illustrates the global relationship between vector length and word frequency.
 It remained unclear, however, whether word frequency was \textit{directly} related to word vector length or rather via correlated factors such as word significance.
@@
 
-On the other hand, recent theoretical work \cite{Arora2015} has approached the problem of explaining the so-called ``compositionality'' property exhibited by some word embeddings and that has been the subject of much recent attention.
+On the other hand, recent theoretical work \cite{Arora2015} has approached the problem of explaining the so-called ``compositionality'' property exhibited by some word embeddings.
 In that work, unnormalised vectors are used in their model of the word relationship task.
-It is hoped that experimental approaches such as those described here might yield insights into the role of the word vector length in the word relationship tasks.
+It is hoped that experimental approaches such as those described here might enable theoretical investigations to describe the role of the word vector length in the word relationship tasks.
 \end{section}
@@
 \begin{section}{Corpus and model}\label{corpus-and-model}
-Our training data is built from the Wikipedia XML datadump from October 2013.
+Our training data is built from the Wikipedia datadump from October 2013.
 In order to remove from the training data the bulk of robot-generated pages on Wikipedia, only pages with at least 20 monthly page views are retained\footnote{For further justification of this rationale, and to obtain the dataset, see \blogpost}.
 Stubs and disambiguation pages are also removed, leaving 463 thousand pages with a total of 482 million words.
+Word frequency is described in overview in Table \ref{word-occurrences-table}.
 
@@
 Identical experimental results were obtained using hierarchical softmax, but these are omitted for succinctness.
-The relatively high minimum word cut-off is chosen to ensure that word vectors receive a sufficient number of gradient updates to be meaningful.
-\textcolor{red}{This cut-off results in a vocabulary of $62025$ words (only unigrams were considered).}
+The relatively high minimum word cut-off is chosen to ensure that word vectors, in all but degenerate cases, receive a sufficient number of gradient updates to be meaningful.
+This cut-off results in a vocabulary of $81117$ words (only unigrams were considered).
 
@@
 
-\begin{figure}\label{fig:frequency-histogram}
-	\includegraphics[scale=0.5]{occurrence-histogram}
-	\caption{
-	Number of vocabulary words by frequency band.
-	}
-\end{figure}
+\begin{table}\label{word-occurrences-table}
+	\input{words-occurrences.tex}
+	\caption{ Number of words, by frequency band, as observed in the unmodified corpus.  }
+\end{table}
 
@@
 To this set of words, the word \word{the} is added, in order to include a high-frequency stopword.
+Table \ref{word-frequency-counts} lists the words chosen for this experiment along with their frequencies in the corpus.
+
 The replacement procedure of subsection \ref{replacement-procedure} is then performed for each of these words, using a geometric rate of decay of $\lambda = 1/2$, and maximum value $n=20$.
@@
 
-Figure \ref{fig:word-frequency-counts} gives the number of occurrences of the words chosen for this experiment.
-
-\begin{figure}\label{fig:word-frequency-counts}
+\begin{table}\label{word-frequency-counts}
 	\input{word-frequency-experiment-counts.tex}
 	\caption{Occurrence counts for words chosen for the word frequency experiment. }
-\end{figure}
+\end{table}
 
@@
 	The replacement procedure is then performed for \word{VOID}, using the same values for $\lambda$ and $n$ as above.
-	Figure \ref{fig:word-frequency-experiment-text-void} shows the effect of this modifications on a sample text ($f = 0.05$ is used there for illustrative purposes).
+	figure \ref{fig:word-frequency-experiment-text-void} shows the effect of this modifications on a sample text ($f = 0.05$ is used there for illustrative purposes).
 
@@
 
-\subsection{Experimental results}
-Word frequency has no effect on word vector direction.
+\subsection{Results}
+\textcolor{red}{rework}
+For well-learned representations, word frequency has no effect on word vector direction.
 Figure \ref{word-frequency-experiment-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus.
-The cosine similarity measures the extent to which two vectors have the same direction, taking a maximum value of $1$ and a minimum value of $-1$.
-Notice that the word vectors associated to tokens derived from meaningful tokens have the same direction.
+Recall that the cosine similarity measures the extent to which two vectors have the same direction, taking a maximum value of $1$ and a minimum value of $-1$.
+Notice that the word vectors associated to tokens derived from a word other that \word{the} and \word{VOID} have the same direction.
 The same is true of the word vectors associated to the \word{VOID\_i} for $i$ \textcolor{red}{below $7$ or $8$}, but that the word vectors \word{VOID\_i} gradually diverge directionally for higher values of $i$.
@@
 Because of its noisiness, a higher number of samples (occurrences) is required to learn this co-occurrence distribution.
-The same holds, but to a lesser extent, for the co-occurrence distribution of the stopword \word{the}.
+The same holds, but to a less pronounced extent, for the co-occurrence distribution of the stopword \word{the}.
+\textcolor{red}{The poorly learned representations for \word{the\_i} for $i = 15, \dots $ and for \word{VOID\_i} for $i = 7, \dots$ are therefore disregarded.}
+
+\textcolor{red}{include a table with nearest neighbours in the normal vocab for the to support this argument}
+Table \ref{word-frequency-experiment-cosine-similarity}
 
 It remains, therefore, to consider the effect of frequency variation on word vector length.
-Figure \ref{fig:frequency-norm-graph} illustrates the global relationship between vector length and word frequency.
 Figure \ref{fig:word-frequency-experiment-graph}, on the other hand, shows this relationship for individual words, both for the word vectors, ``syn0'' and for the discarded second layer vectors ``syn1neg'' (included for completeness).
 Each line corresponds to a single word, and the points on each line indicate the frequency and vector length for the tokens derived from the corresponding word.
-\textcolor{red}{For example, the five points on the line corresponding to \word{admiral} are labeled, from right to left, by the tokens \word{ADMIRAL\_1}, \word{ADMIRAL\_2}, \dots, \word{ADMIRAL\_5}.}
+For example, the six points on the line corresponding to \word{protestant} are labeled, from right to left, by the tokens \word{PROTESTANT\_1}, \word{PROTESTANT\_2}, \dots, \word{PROTESTANT\_6}.
 The number of points on the line is determined by the frequency of the original word.
-\textcolor{red}{For example, the frequency of the \word{admiral} can be halved at most $5$ times and remain about the minimum frequency cut-off.}
+For example, the frequency of the \word{protestant} can be halved at most $6$ times and remain about the minimum frequency cut-off.
 
-Because all the points on a line share a co-occurrence distribution (only the frequency varies), this figure demonstrates conclusively that length does indeed depend on frequency directly, and not only via some correlation between word frequency and word significance.
-Moreover, this relationship is essentially linear for any particular word.
+Because all the points on a line share a co-occurrence distribution (only the frequency varies), figure \ref{fig:word-frequency-experiment-graph} demonstrates conclusively that length does indeed depend on frequency directly, and not only via some correlation between word frequency and word significance.
+Moreover, this relationship is linear for any particular word.
 Notice also that the relative positions of the norms of the word vectors associated to the test words is independent of the frequency band.
@@
 
-\begin{figure}
-	\includegraphics[scale=0.3]{frequency-norm-scatterplot}
-	\caption{ The global relationship between word frequency and vector length.  }
-	\label{fig:frequency-norm-graph}
-\end{figure}
+Notice also that the word vector for \word{VOID} is non-zero.
 
@@
 
+\begin{table}\label{word-frequency-experiment-cosine-similarity}
+	\input{cosine-similarity.tex}
+	\caption{
+		The most similar words in the original vocabulary to the tokens
+		introduced into the corpus associated to the word \word{the},
+		and their cosine similarity with the most frequent such token,
+		\word{THE\_1}.  It is apparent from the nearest neigbours that
+		the word vectors of the low frequency tokens have not been
+		adequately trained (there were inadequate samples).
+	}
+\end{table}
+
 \begin{sidewaysfigure*}\label{fig:word-frequency-experiment-graph}
@@
 	\caption{
-	Vector length vs. frequency for the words chosen for the word frequency
+	Vector length vs frequency for the words chosen for the word frequency
 	experiment.  For each word, tokens of varying frequency but with the
@@
 These occurrences (and additionally, with probability, some occurrences from the cats section) are ``noise occurrences''.
+Table \ref{cooccurrence-noise-words} lists the words chosen for the experiment, along with their frequencies in the corpus.
 
@@
 
-\begin{figure}
+\begin{table}\label{cooccurrence-noise-words}
 	\input{noise-cooccurrence-experiment-counts.tex}
@@
 	\caption{Occurrence counts for words chosen for the co-occurrence noise experiment. }
-\end{figure}
+\end{table}
 
@@
 \begin{subsection}{Results}
-Figure \ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus.
+Figure \ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus in this experiment.
 Notice that the word vectors associated to tokens derived from the same word all share a common direction.
@@
 
-On the other hand, Figure \ref{fig:co-occurrence-noise-graph} shows that word vector length varies directly and linearly with the proportion $r$ of noise in the co-occurrence distribution.
+On the other hand, figure \ref{fig:co-occurrence-noise-graph} shows that word vector length varies directly and linearly with the proportion $r$ of noise in the co-occurrence distribution.
 This figure motivates an interpretation of vector length, within a sufficiently narrow frequency band, as a measure of the absence of co-occurrence noise, that is, of the extent to which a word determines a context.
@@
 	\caption{
-	Vector length vs. proportion of co-occurrence noise for words
+	Vector length vs proportion of co-occurrence noise for words
 	chosen for the co-occurrence noise experiment.  For each word,
@@
 
-\begin{section}{Discussion and future directions}\label{future-directions}
+\begin{section}{Discussion}\label{future-directions}
 
 \begin{subsection}{Controlled experiments for word embeddings}
-The principle contribution of this short note has been to demonstrate that controlled experiments can be used to gain insight into any word embedding.
-Because they are achieved via modification of the training corpus, these experiments can be carried out for any language model or word embedding.
-The experiments are moreover easy to carry out, requiring modification of code nor even knowledge of the model itself.
+Our principle contribution has been to demonstrate that controlled experiments can be used to gain insight into a word embedding.
+Because they are achieved via modification of the training corpus, these experiments can be carried out for any word embedding (or indeed language model), requiring no knowledge or modification of the model implementation.
 
-More elaborate experiments could be carried out in the context of word embeddings.
-For instance, by introducing tokens into the corpus that mix, with varying proportions, the co-occurrence distributions of two words, the path between the word vectors in the feature space can be studied.
+It would, of course, be interesting to perform these experiments for other word embeddings other than word2vec CBOW and for different hyperparameters settings.
+
+More elaborate experiments could be carried out.
+For instance, by introducing tokens into the corpus that mix, with varying proportions, the co-occurrence distributions of two words, the path between the word vectors in the feature space could be studied.
 The co-occurrence noise variation experiment described here would be a special case of such an experiment where one of the two words was \word{VOID}.
-
-It would, of course, be interesting to perform these experiments for other word embeddings and for different hyperparameters settings.
 \end{subsection}
@@
 Further afield, it would be interesting to perform analogs of these experiments in other contexts where models are trained from co-occurrence data.
-For instance, in the context of e-Commerce, product purchase frequency and product co-occurrence noise can be varied by modifying the item-item matrix.
+For instance, in the context of e-Commerce, product purchase frequency and product co-occurrence noise can be varied by modifying the product-product co-occurrence matrix.
 \end{subsection}
@@
 \begin{subsection}{Questions pertaining to word2vec}
-Questions pertaining to particularly to word2vec arise naturally from the results of the experiments.
+Questions pertaining to word2vec in particular arise naturally from the results of the experiments.
 Figures \ref{fig:frequency-norm-graph} and \ref{fig:co-occurrence-noise-graph}, for example, demonstrate the word vectors obtained from the first synaptic layer, ``syn0'', have very different properties from those that could be obtained from the second layer, ``syn1neg'', which are typically discarded.
@@
 Figure \ref{fig:frequency-norm-graph} shows that this point is not at the origin of the feature space (i.e. is not the zero vector).
-It is the origin, however, that is implicitly used as the point of reference in the word2vec word similarity task.
+The origin, however, is implicitly the point of reference in the word2vec word similarity task.
 This raises the question of whether improved performance on the similarity task could be achieved by transforming the feature space (or modifying the model) such that the representation of pure noise (i.e.

### 2015-08-20 article_generate_cosine_similarity.py

@@ -1,0 +1,31 @@
+import numpy as np
+import pandas as pd
+import sys
+
+from parameters import *
+from functions import *
+
+
+vectors_syn0_filename = sys.argv[1]
+word = sys.argv[2] # e.g. 'the'
+
+vectors, norms = row_normalise_dataframe(load_word2vec_binary(vectors_syn0_filename))
+
+# e.g. 'THE_1' ...
+tokens = [build_experiment_token(word, i) for i in range(1, max(word_freq_experiment_ratio, word_freq_experiment_power_max) + 1)]
+tokens = [idx for idx in tokens if idx in vectors.index]
+
+non_experiment_idxs = [idx for idx in vectors.index if idx != idx.upper()] # all non uppercase words are non experiment words
+non_experiment_vectors = vectors.loc[non_experiment_idxs]
+
+print r'\begin{tabular}{l | c | l}'
+print r'token & similarity to \word{%s} & most similar words in unmodified corpus\\' % tokens[0].replace('_', '\_')
+print '\hline'
+for token in tokens:
+    cs = vectors.loc[token].dot(vectors.loc[tokens[0]])
+    bydist = by_distance_from(non_experiment_vectors, vectors.loc[token])
+    similiar_words = ', '.join(bydist[0:4].index)
+    print r'\word{%s} & %.4f & \word{%s} \\' % (token.replace('_', '\_'), cs, similiar_words)
+print r'\end{tabular}'
+
+

### 2015-08-20 article_generate_word_counts.py

@@ -1,0 +1,25 @@
+import numpy as np
+import pandas as pd
+import sys
+
+from parameters import *
+from functions import *
+
+
+word_counts_filename = sys.argv[1]
+with file(word_counts_filename) as f:
+    word_counts = read_word_counts(f)
+
+words = word_counts.keys()
+stats = pd.DataFrame({'occurrences': [word_counts[word] for word in words], 'word': words})
+stats['occurrences_band'] = np.floor(np.log2(stats.occurrences)).astype(int)
+df = stats.groupby('occurrences_band').word.aggregate({'number': len, 'words': lambda words: ', '.join(random.sample(words, min(len(words), 4)))}).sort_index()
+
+print r'\begin{tabular}{c | r | l}'
+print r'frequency band & \# words & example words  \\'
+print '\hline'
+for band, row in df.iterrows():
+    print r'$2^{%i} - 2^{%i}$ & %i & \word{%s} \\' % (band, band+1, row['number'], row['words'])
+print r'\end{tabular}'
+
+

### 2015-08-20 build_images.py

@@ -37,50 +37,38 @@ stats['log2_frequency'] = np.log2(stats.occurrences * 1. / total_words)
 stats.L2_norm_syn0.name = 'L2 norm (syn0)'
 stats.L2_norm_syn1neg.name = 'L2 norm (syn1neg)'
 
-band_counts = stats.occurrences_band.dropna().value_counts().sort_index()
-band_counts.index.name = 'log2 (# occurrences)'
 
-_ = band_counts.plot(kind='bar')
-plt.title('Number of words in each occurrence count band', y=1.08)
-plt.tight_layout()
-plt.savefig('outputs/occurrence-histogram.eps')
+# WORD FREQUENCY EXPERIMENT
 
-band_gb = stats.groupby('occurrences_band')
-means = band_gb.L2_norm_syn0.mean()
-errors = band_gb.L2_norm_syn0.std()
-ax = means.plot(yerr=errors)
+def wf_experiment_tokens(word):
+    """
+    return the list of tokens introduced into the corpus in the word frequency
+    experiment for the given word.
+    """
+    idxs = [build_experiment_token(word, i) for i in range(1, max(word_freq_experiment_ratio, word_freq_experiment_power_max) + 1)]
+    return [idx for idx in idxs if idx in vectors_syn0.index]
 
-means = band_gb.L2_norm_syn1neg.mean()
-errors = band_gb.L2_norm_syn1neg.std()
-ax = means.plot(yerr=errors, figsize=(7,7))
+# cosine similarity heatmap
+test_words = random.sample([word for word in word_freq_experiment_words if word != 'the'], 4)
+test_words += ['the', meaningless_token]
+idxs = []
+for word in test_words:
+    idxs += wf_experiment_tokens(word)
+test_vecs = vectors_syn0.loc[idxs]
+cosine_similarity_heatmap(test_vecs, figsize=(12, 10))
+plt.savefig('outputs/word-frequency-experiment-heatmap.eps')
 
-_ = ax.set_xlim(7, 20)
-_ = ax.set_ylim(0, 60)
+# remove word vectors that are not well trained
+# drop those whose cosine similarity with word_1 is < 0.8
+for word in word_freq_experiment_words + [meaningless_token]:
+    idxs = wf_experiment_tokens(word) # ordered: word_1, word_2 ..
+    test_vecs = vectors_syn0.loc[idxs]
[…]

### 2015-08-20 functions.py

@@ -2,6 +2,7 @@ from __future__ import print_function
 import random
 import numpy as np
 import pandas as pd
+from scipy.spatial.distance import pdist, cdist
 import matplotlib.pyplot as plt
 from collections import defaultdict
 from parameters import random_seed, experiment_word_occurrence_min
@@ -146,9 +147,16 @@ def load_word2vec_binary(fname):
             vectors[line_no] = vector
     return pd.DataFrame(vectors, index=vocab)
 
+def cosine_similarity(vecs):
+    """
+    return the cosine similarity of each row with all the others.
+    'vecs' is a dataframe
+    """
+    vecs_normed = vecs.as_matrix() / np.sqrt((vecs ** 2).sum(axis=1))[:,np.newaxis]
+    return vecs_normed.dot(vecs_normed.transpose())    
+
 def cosine_similarity_heatmap(test_vecs, **kwargs):
-    test_vecs_normed = test_vecs.as_matrix() / np.sqrt((test_vecs ** 2).sum(axis=1))[:,np.newaxis]
-    mat = test_vecs_normed.dot(test_vecs_normed.transpose())    
+    mat = cosine_similarity(test_vecs)
     plt.figure(**kwargs)
     plt.title('Cosine similarity of word vectors')
     plt.pcolor(mat, vmin=-1, vmax=1)
@@ -156,3 +164,34 @@ def cosine_similarity_heatmap(test_vecs, **kwargs):
     _ = plt.yticks(np.arange(0.5, len(test_vecs.index), 1), test_vecs.index)
     _ = plt.xticks(np.arange(0.5, len(test_vecs.index), 1), test_vecs.index, rotation=90)
 
+def by_distance_from(table, v, **params):
+    """
+    Return a Series, listing the distance of each row
+    from the vector given.
+    To specify different metrics, see docstring of scipy.spatial.distance.cdist
+    (default if Euclidean).
+    """
+    v = np.array(v).reshape(1, -1)
+    dist = pd.Series(
+        cdist(v, table, **params)[0], index=table.index, copy=True)
+    dist.sort()
+    return dist
+
+def row_normalise_matrix(matrix):
+    """
+    All matrices are represented by 2 dimension np.array instances (NOT np.matrix)
+    return the matrix, along with the norms.
+    """
+    matrix = matrix * 1.
+    norms = np.sqrt((matrix ** 2).sum(axis=1))
+    return matrix / norms[:, np.newaxis], norms
+
+def row_normalise_dataframe(df):
[…]

### 2015-08-25 article/main.tex
fix heatmap whitespace; rework WFVE results section; tables

@@
 \textcolor{red}{what about some work introducing tokens into a corpus for experiments?}
+\textcolor{red}{cite ``contextual distinctiveness: a new lexical property ...'' by McDonald and Shillcock?}
 
@@
 
-\subsection{Results}
-\textcolor{red}{rework}
-For well-learned representations, word frequency has no effect on word vector direction.
+\subsection{Results: word frequency and vector direction}\label{WFVE-direction}
 Figure \ref{word-frequency-experiment-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus.
 Recall that the cosine similarity measures the extent to which two vectors have the same direction, taking a maximum value of $1$ and a minimum value of $-1$.
-Notice that the word vectors associated to tokens derived from a word other that \word{the} and \word{VOID} have the same direction.
-The same is true of the word vectors associated to the \word{VOID\_i} for $i$ \textcolor{red}{below $7$ or $8$}, but that the word vectors \word{VOID\_i} gradually diverge directionally for higher values of $i$.
-This demonstrates the difficulty of learning the co-occurrence distribution of $\word{VOID}$, the global frequency distribution.
-Because of its noisiness, a higher number of samples (occurrences) is required to learn this co-occurrence distribution.
-The same holds, but to a less pronounced extent, for the co-occurrence distribution of the stopword \word{the}.
-\textcolor{red}{The poorly learned representations for \word{the\_i} for $i = 15, \dots $ and for \word{VOID\_i} for $i = 7, \dots$ are therefore disregarded.}
+The number of word vectors associated to any word is the number of times that its frequency cat be halved and remain above the chosen minimum word count threshold of $128$.
 
-\textcolor{red}{include a table with nearest neighbours in the normal vocab for the to support this argument}
-Table \ref{word-frequency-experiment-cosine-similarity}
+Consider first the words other than \word{the} and \word{VOID}, e.g. \word{protestant}.
+These words occur in a limited number of contexts, and so fewer samples are required to capture their co-occurrence distributions of such words.
+Thus their word vectors can be learned with less samples.
+It is therefore to be expected that the word vectors of their associated tokens can be properly learned from the modified corpus.
+The red blocks of figure \ref{word-frequency-experiment-heatmap} clearly show that the direction of these word vectors depends only upon the original word.
 
+The word \word{VOID}, on the other hand, occurs with probability in every context, by construction.
+Thus a much higher number of samples are required to capture its co-occurrence distribution, and thereby to learn its word vector.
+The same is true, but to a lesser extent, of the stopword \word{the}.
+The ``difficulty'' of learning the word vector associated to \word{VOID\_i} or \word{the\_i} increases with $i$.
+This is consistent with figure \ref{word-frequency-experiment-heatmap}, where it is apparent that the word vectors associated to \word{VOID\_i} for $i = 1, \dots, 5$ all share a common direction, while the word vectors for diverge directionallly for higher values of $i$.
+The same trend is apparent, but less extremely so, for \word{the}.
+The word vectors for \word{the\_i} share a common direction for $i = 1, \dots, 14$, before diverging directionally for higher values of $i$.
+We conclude that these directional divergences are artifacts of poor learning due to an insufficient number of occurrences.
+Table \ref{word-frequency-experiment-cosine-similarity}, showing the most similar words to the word vector of each \word{the\_i} provides further support for this claim.
+In further discussion, the tokens \word{VOID\_i} (respectively, \word{the\_i}) whose word vectors have a cosine similarity less than $0.8$ with \word{VOID\_1} (respectively, \word{the\_1}) are therefore disregarded.
+
+In conclusion, word frequency has no effect on word vector direction, given a sufficient number of samples to learn the co-occurrence distribution of the word.
+%For well-learned representations, word frequency has no effect on word vector direction.
+
+\subsection{Results: word frequency and vector length}
 It remains, therefore, to consider the effect of frequency variation on word vector length.
-Figure \ref{fig:word-frequency-experiment-graph}, on the other hand, shows this relationship for individual words, both for the word vectors, ``syn0'' and for the discarded second layer vectors ``syn1neg'' (included for completeness).
+Figure \ref{fig:word-frequency-experiment-graph} shows this relationship for individual words, both for the word vectors, ``syn0'' and for the discarded second-layer vectors ``syn1neg'' (included for completeness).
 Each line corresponds to a single word, and the points on each line indicate the frequency and vector length for the tokens derived from the corresponding word.
 For example, the six points on the line corresponding to \word{protestant} are labeled, from right to left, by the tokens \word{PROTESTANT\_1}, \word{PROTESTANT\_2}, \dots, \word{PROTESTANT\_6}.
-The number of points on the line is determined by the frequency of the original word.
+Again, the number of points on the line is determined by the frequency of the original word.
 For example, the frequency of the \word{protestant} can be halved at most $6$ times and remain about the minimum frequency cut-off.
@@
 Moreover, this relationship is linear for any particular word.
-Notice also that the relative positions of the norms of the word vectors associated to the test words is independent of the frequency band.
-Therefore these lines trace lines along which a frequency-adjusted norm would be constant.
+Notice also that the relative positions of the norms of the word vectors associated to the experiment words is independent of the frequency band, i.e. the plotted lines rarely cross.
+The plotted lines may thus be seen as the approximate contours of a hypothetical measure of word vector length normalised for word frequency effects.
 
-Notice also that the word vector for \word{VOID} is non-zero.
+Notice the inflection of the ``syn1neg'' curve for \word{the} at approximately $10^6$.  We have no explanation of this.
+
+As an aside, notice also that the word vector length for the tokens \word{VOID\_i} is approximately constant.
+Since the direction is also constant (cf. subsection \ref{WFVE-direction}), it is therefore sensible to speak of the word vector of \word{VOID} irrespective of its frequency.
+In particular, \word{VOID\_1} may be taken as an approximation.
 
@@
 \begin{subsection}{Results}
+	%The co-occurrence distributions of e.g.
+the tokens \word{}, for $i=1, \dots, 8$ increasingly approximate the global word frequency distribution, i.e.
+co-occurrence distribution of the token \word{VOID}.
+	%Since the vectorisation of \word{VOID} is independent of its frequency (cf.
+remark \ref{}), the word vectors of \word{} can therefore be considered as a series of intermediate points on a hypothetical path in the feature space beginning at the word vector for \word{} and ending at the word vector for \word{VOID}.
+	\textcolor{red}{all of the following is true (using the heatmap that i say it does) but it ignores the artifact that the path converges to the word vector for \word{VOID}, not to zero -- this is fine to the extent that they are approximately the same.}
+
 Figure \ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus in this experiment.
@@
 This raises the question of whether improved performance on the similarity task could be achieved by transforming the feature space (or modifying the model) such that the representation of pure noise (i.e.
-the vector for \word{VOID}) was at the origin of the transformed feature space.
+the vector for \word{VOID}) is at the origin of the transformed feature space.
+
+\textcolor{red}{can it be argued that the feature space must be fundamentally euclidean, given that the path marked by the vectorisations of intermediate words is a straight line?
+i.e.
+if such paths maybe considered as ``geodesics'' between words, and we show that these geodesics appear as straight lines when an euclidean metric is imposed, then suggests that the euclidean metric is in fact a good choice.
+what would happen if we chose another metric?}
+
 \end{subsection}

### 2015-08-25 build_images.py

@@ -8,6 +8,8 @@ import sys
 from parameters import *
 from functions import *
 
+random.seed(0) # fix seed so that we can refer to the randomly chosen words in the article body
+
 vectors_syn0_filename = sys.argv[1]
 vectors_syn1neg_filename = sys.argv[2]
 word_counts_filename = sys.argv[3]
@@ -15,6 +17,7 @@ word_freq_exp_words_filename = sys.argv[4]
 coocc_noise_exp_words_filename = sys.argv[5]
 
 matplotlib.rcParams.update({'font.size': 14})
+matplotlib.rcParams.update({'figure.autholayout': True})
 
 word_freq_experiment_words = read_words(word_freq_exp_words_filename)
 coocc_noise_experiment_words = read_words(coocc_noise_exp_words_filename)

### 2015-08-25 functions.py

@@ -159,10 +159,13 @@ def cosine_similarity_heatmap(test_vecs, **kwargs):
     mat = cosine_similarity(test_vecs)
     plt.figure(**kwargs)
     plt.title('Cosine similarity of word vectors')
+    _ = plt.gca().set_ylim(0, len(test_vecs.index))
+    _ = plt.gca().set_xlim(0, len(test_vecs.index))
     plt.pcolor(mat, vmin=-1, vmax=1)
     plt.colorbar()
     _ = plt.yticks(np.arange(0.5, len(test_vecs.index), 1), test_vecs.index)
     _ = plt.xticks(np.arange(0.5, len(test_vecs.index), 1), test_vecs.index, rotation=90)
+    plt.tight_layout()
 
 def by_distance_from(table, v, **params):
     """

### 2015-08-25 article/main.tex
don't sub VOID vec in coocc noise exp heatmap; fix figure refs

@@
 \textcolor{red}{what about some work introducing tokens into a corpus for experiments?}
+
 \textcolor{red}{cite ``contextual distinctiveness: a new lexical property ...'' by McDonald and Shillcock?}
@@
 
-\begin{table}\label{word-occurrences-table}
+\begin{table}
 	\input{words-occurrences.tex}
 	\caption{ Number of words, by frequency band, as observed in the unmodified corpus.  }
+	\label{word-occurrences-table}
 \end{table}
@@
 
-\begin{table}\label{word-frequency-counts}
+\begin{table}
 	\input{word-frequency-experiment-counts.tex}
 	\caption{Occurrence counts for words chosen for the word frequency experiment. }
+	\label{word-frequency-counts}
 \end{table}
 
-\begin{figure}\label{fig:word-frequency-experiment-text-cat}
+\begin{figure}
 	\begin{mdframed}
@@
 	word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
+	\label{fig:word-frequency-experiment-text-cat}
 \end{figure}
@@
 
-\begin{figure}\label{fig:word-frequency-experiment-text-void}
+\begin{figure}
 	\begin{mdframed}
@@
 For illustrative purposes, the token \word{VOID} was here interspersed with a global frequency of $0.05$.}
+	\label{fig:word-frequency-experiment-text-void}
 \end{figure}
@@
 
-\begin{figure}\label{word-frequency-experiment-heatmap}
+\begin{figure}
 	\includegraphics[scale=0.5]{word-frequency-experiment-heatmap}
@@
 	}
+	\label{word-frequency-experiment-heatmap}
 \end{figure}
 
-\begin{table}\label{word-frequency-experiment-cosine-similarity}
+\begin{table}
 	\input{cosine-similarity.tex}
@@
 	}
+	\label{word-frequency-experiment-cosine-similarity}
 \end{table}
 
-\begin{sidewaysfigure*}\label{fig:word-frequency-experiment-graph}
+\begin{sidewaysfigure*}
 	\centering{\includegraphics[scale=0.6]{word-frequency-experiment-graph}}
@@
 	}
+	\label{fig:word-frequency-experiment-graph}
 \end{sidewaysfigure*}
@@
 
-\begin{table}\label{cooccurrence-noise-words}
+\begin{table}
 	\input{noise-cooccurrence-experiment-counts.tex}
@@
 	\caption{Occurrence counts for words chosen for the co-occurrence noise experiment. }
+	\label{cooccurrence-noise-words}
 \end{table}
@@
 
-\begin{figure}\label{fig:co-occurrence-noise-heatmap}
+\begin{figure}
 	\includegraphics[scale=0.5]{cooccurrence-noise-heatmap}
@@
 	}
+	\label{fig:co-occurrence-noise-heatmap}
 \end{figure}
 
-\begin{sidewaysfigure*}\label{fig:co-occurrence-noise-graph}
+\begin{sidewaysfigure*}
 	\includegraphics[scale=0.6]{cooccurrence-noise-graph}
@@
 	}
+	\label{fig:co-occurrence-noise-graph}
 \end{sidewaysfigure*}

### 2015-08-25 build_images.py

@@ -158,6 +158,6 @@ for word in random.sample(coocc_noise_experiment_words, 4):
 
 # to check for colinearity, we need to subtract the vector that they all converge to, viz. the meaningless vector.
 meaningless_vec = vectors_syn0.loc[build_experiment_token(meaningless_token, 1)]
-test_vecs = vectors_syn0.loc[idxs].dropna() - meaningless_vec
+test_vecs = vectors_syn0.loc[idxs].dropna()
 cosine_similarity_heatmap(test_vecs, figsize=(12, 10))
 plt.savefig('outputs/cooccurrence-noise-heatmap.eps')

### 2015-09-03 article/main.tex
incorporate the changes by adriaan

--- 2015-08-20 article/main.tex
@@
 \usepackage{mdframed}
-%\usepackage{wrapfig}
 \usepackage{rotating}
@@
 \usepackage{amsmath,amssymb,amsfonts,comment}
-%\definecolor{LinkColor}{rgb}{0,0,0}
 
-\newcommand{\p}{\text{P}}
+\newcommand{\p}{\mathrm{P}}
 \newcommand{\word}[1]{\texttt{#1}}
+\newcommand{\mar}[1]{\marginpar{\footnotesize \bfseries
+    \raggedright #1}}
+
 \urldef{\googlegroup}\url{https://groups.google.com/forum/#!forum/word2vec-toolkit}
@@
 
-\title{Corpus Experiments for Word Embeddings}
+\title{Controlled Experiments for Word Embeddings}
 
@@
 	Adriaan Schakel\\
-	\textcolor{red}{NNLP?}\\
+	NNLP\\
 	\texttt{adriaan.schakel@gmail.com}
@@
 \begin{abstract}
-	We introduce an experimental approach to studying the properties of word embeddings.
-	Controlled experiments, achieved through modifications of the training corpus, permit the demonstration of direct, causal relationships between word properties and word vector direction and length.
-	This approach is demonstrated in the case of the word2vec CBOW model with experiments that independently vary word frequency and word co-occurrence noise.
-	In this case, it is shown that word vector length depends directly and linearly on both word frequency and the level of noise in the co-occurrence distribution of the word.
-	In both cases, the coefficient of linearity depends upon the word.
+	An experimental approach to studying the properties of word
+        embeddings is proposed.  Controlled experiments, achieved
+        through modifications of the training corpus, permit the
+        demonstration of direct relations between word properties and
+        word vector direction and length.  The new approach is
+        demonstrated using the word2vec CBOW model with experiments that
+        independently vary word frequency and word co-occurrence noise.
+        The experiments reveal that word vector length depends more or
+        less linearly on both word frequency and the level of noise in
+        the co-occurrence distribution of the word.  The coefficients of
+        linearity depend upon the word.
 \end{abstract}
@@
 \begin{section}{Introduction}
-Word embeddings are typically trained with samples from the word co-occurrence distributions.
-Recall that the co-occurrence distribution of a word $w$ gives the probability $\p(w'|w)$ that a word $w'$ occurs nearby, given that $w$ occurred.
-Samples of the co-occurrence distributions are obtained by scanning a short window over the text.
-The neighbouring words are treated as samples of the co-occurrence distribution the current word. 
-Viewed in this way, the word vectors of the word embedding are determined by the co-occurrence distributions and frequencies of the words in the corpus.
+The field of natural language processing (NLP) is presently undergoing rapid
+developments, not in the least triggered by the tremendous commercial
+promises it holds.  A timely research topic in NLP are word embeddings,
+where words from a vocabulary are represented by dense, real-valued
+vectors.  Instead of one-hot vectors that merely indicate the location
+of a word in the vocabulary, dense
+vectors of dimension much smaller than the vocabulary size are
+constructed such that they carry syntactic and semantic information.
+Irrespective of the technique chosen, word embeddings are
+typically derived from word co-occurrences.  More specifically, in a
+machine-learning setting, word embeddings are typically trained by
+scanning a short window over all the text in a corpus.  This process can
+be seen as sampling word co-occurrence distributions, where it is
+recalled that the co-occurrence distribution of a target word $w$
+denotes the conditional probability $\p(w'|w)$ that a word $w'$ occurs in
+its context, i.e., given that $w$ occurred.  Most applications of word
+embeddings explore not the word vectors themselves, but relations
+between them to solve, for example, similarity and word relation tasks
+\cite{vecchi-baroni-zamparelli2011}.  For these tasks, it was found that using normalised word
+vectors improves performance.  Word vector length is therefore typically
+ignored.
 
-We perform experiments that illustrate the effect of these two determining pieces of information (the co-occurrence distribution and the word frequency) as they are varied independently of one another.
-This is achieved through the introduction of new tokens into the training corpus with varying frequencies and varying levels of noise in their co-occurrence distributions.
-The frequency and co-occurrence distributions of these introduced tokens are modeled on existing words in the corpus.
+In a previous paper \cite{schakel-wilson}, we proposed the use of word
+vector length as measure of word significance.  Using a domain-specific
+corpus of scientific abstracts, we observed that words that appear only
+in similar contexts tend to have longer vectors than words of the same
+frequency that appear in a wide variety of contexts.  For a given
+frequency band, we found meaningless function words clearly separated
+from proper nouns, each of which typically carries the meaning of a
+distinctive context in this corpus.  In other words, the longer its
+vector, the more significant a word is.  We also observed that word
+significance is not the only factor determining the length of a word
+vector, also the frequency with which a word occurs plays an important
+role.
 
-We illustrate our approach in case of the word2vec CBOW model.
-However, these experiments (and other besides) could equally be carried out for word embedding methods such as word2vec skip-gram \cite{DistRepns,EfficientEstimation}, GloVe \cite{pennington2014glove} and SENNA \cite{collobert-2011}.
+In this paper, we wish to study in detail to what extent these two
+factors determine word vectors.  For a given corpus, both term
+frequency and co-occurrence are, of course, fixed and it is not obvious
+how to unravel these dependencies in an unambiguous, objective manner.
+In particular, it is difficult to establish the distinctiveness of the
+contexts in which a word is used.  To overcome these problems, we
+propose to modify the training corpus in a controlled fashion.  To
+this end, we insert new tokens into the corpus with varying frequencies
+and varying levels of noise in their co-occurrence distributions.  By
+modelling the frequency and co-occurrence distributions of these tokens
+on existing words in the corpus, we are able to study their effect on
+word vectors independently of one another.  We can thus study a family
+of tokens that all appear in the same context, but with different
+frequencies, or study a family of tokens that all have the same
+frequency, but appear in a different number of contexts.  Starting from
+the limited number of contexts in which a word appears in the original
+corpus, we can increase this number by inspersing the word in arbitrary
+contexts at random.  The word thus looses its significance in a
+controlled way.  Although we present our approach using the word2vec
+CBOW model, these and related experiments could equally well be carried
+out for other word embedding methods such as the word2vec skip-gram
+model \cite{DistRepns,EfficientEstimation}, GloVe
+\cite{pennington2014glove}, and SENNA \cite{collobert-2011}.
 
-In the case of word2vec, we show that word vector length depends directly and linearly on both word-frequency and the level of noise in the co-occurrence distribution of the word.
-In both cases, the coefficient of linearity depends upon the word.
-If the co-occurrence distribution is fixed, then word vector length increases with word frequency.
-On the other hand, if word frequency is held constant, then word vector length decreases as the level of noise in the co-occurrence distribution of the word is increased.
+We show that the length of the word vectors generated by the CBOW model
+depends more or less linearly on both word frequency and level of noise
+in the co-occurrence distribution of the word.  In both cases, the
+coefficient of linearity depends upon the word.  If the co-occurrence
+distribution is fixed, then word vector length increases with word
+frequency.  If, on the other hand, word frequency is held constant, then
+word vector length decreases as the level of noise in the co-occurrence
+distribution of the word is increased.  We show furthermore that the
+direction of the word vectors is independent of both word frequency and
+co-occurrence noise.
 
-We show furthermore that word vector direction does not vary with word frequency or with the level of co-occurrence noise.
-
-This paper is structured as follows.
-Section \ref{related-work} draws connections to related work, while section \ref{corpus-and-model} describes the corpus and model used to illustrate the experiments.
-Section \ref{WFVE} describes a controlled experiment for varying word frequency while holding the co-occurrence distribution fixed.
-Section \ref{CNVE}, in a complementary fashion, describes a controlled experiment for varying the level of noise in the co-occurrence distribution of a word, while holding the word frequency fixed.
-Section \ref{future-directions} considers further questions and possible future directions.
+This paper is structured as follows.  Section \ref{related-work} draws
+connections to related work, while Section \ref{corpus-and-model} describes
+the corpus and the CBOW model used in our experiments.
+Section \ref{WFVE} describes a controlled experiment for varying word
+frequency while holding the co-occurrence distribution fixed.
+Section \ref{CNVE}, in a complementary fashion, describes a controlled
+experiment for varying the level of noise in the co-occurrence
+distribution of a word while holding the word frequency fixed.  The
+final section, Section \ref{future-directions}, considers further questions
+and possible future directions.
 \end{section}
@@
 \begin{section}{Related work}\label{related-work}
-Our experimental finding (in the context of word2vec), that word vector length decreases with co-occurrence noise is reminiscent of \cite{vecchi-baroni-zamparelli2011}, where a relationship between vector length and the ``semantic deviance'' of an adjective-noun composition was studied observationally.
-In \cite{schakel-wilson}, the interpretation of word vector length as an indication of word significance is explored in the context of the arXiv high-energy physics corpus.
+Our notion of word significance is reminiscent of the concept of
+``contextual distinctiveness'' introduced earlier by McDonald and
+Shillcock \cite{mcdonald2001contextual}, who also based their work on word co-occurrences.
+They define contextual distinctiveness of a target word $w$ as the
+Kullback-Leibler divergence $\mathrm{D} \big(\p(w') ||\p(w'|w) \big)$
+between the distribution of context words, $\p(w')$, and the
+co-occurrence distributions of that target word $w$.  Specifically,
+\begin{equation} 
+\label{cd}
+\mathrm{D} \big(\p(w') ||\p(w'|w) \big) = \sum_{w'} \p(w'|w) \log_2
+\frac{\p(w'|w)}{\p(w')},
+\end{equation} 
+where the sum is over the set of \textcolor{blue}{content?} words $w'$.  They set the
+window size to five words before and five words after the target word.
+To reduce the computational burden, they recorded co-occurrences of 690
+target words, randomly chosen from equally divided frequency bands, with
+only the $n=500$ most frequent content words.  As pointed out by the
+authors, contextual distinctiveness is a measure of the amount of
+information provided by word $w$ about its contexts of use.  More specifically, if a
+word $w$ tends to appear in a wide variety of contexts, the distribution
+of the words it occurs with, i.e., the posterior distribution $\p(w'|w)$,
+tends to diverge less from the prior distribution $\p(w')$, and its
+contextual distinctiveness is accordingly small.  If, on the other hand,
+$w$ typically appears in only a small number of different contexts, the
+posterior distribution tends to diverge more from the prior, and a
+larger value of contextual distinctiveness results.  As for
+word significance, contextual distinctiveness summarizes a word's
+contextual behavior in a single number. \textcolor{blue}{shorten?}
 
-\textcolor{red}{what about some work introducing tokens into a corpus for experiments?}
+Our experimental finding that word vector length decreases with
+co-occurrence noise is related to earlier work by Vecchi, Baroni, and
+Zamparelli \cite{vecchi-baroni-zamparelli2011}, where a relation between
+vector length and the ``semantic deviance'' of an adjective-noun
+composite was studied empirically.  In that paper, which is also based
+on word co-occurrence, the authors study adjective-noun composites.
+They built a vocabulary from the 8k most frequent nouns and 4k most
+frequent adjectives in a large general language corpus and added 22k
+adjective-noun composites.  For each item in the vocabulary, they
+recorded their co-occurrence within the same sentence with any of the
+top 10k most frequent content words (nouns, adjectives or verbs) in the
+corpus, and contructed word embeddings (of dimension 300) through
+singular value decomposition of the co-occurrence matrix.  The authors considered several models for
+constructing vectors of unattested adjective-noun composites, the two
+simplest being adding and component-wise multipying the adjective and
+noun vectors.  They hypothesized that the length of the vectors thus
+constructed can be used to distinguish acceptable and semantically
+deviant adjective-noun composites.  Using a few hundred adjective-noun
+composites selected by humans for evaluation, they found that deviant
+composites have a shorter vector than acceptable ones, in accordance
+with their expectation.
 
@@
 \begin{section}{Corpus and model}\label{corpus-and-model}
-Our training data is built from the Wikipedia datadump from October 2013.
-In order to remove from the training data the bulk of robot-generated pages on Wikipedia, only pages with at least 20 monthly page views are retained\footnote{For further justification of this rationale, and to obtain the dataset, see \blogpost}.
-Stubs and disambiguation pages are also removed, leaving 463 thousand pages with a total of 482 million words.
-Word frequency is described in overview in Table \ref{word-occurrences-table}.
+Our training data is built from the Wikipedia datadump from October
+2013.  To remove the bulk of robot-generated pages from the training
+data, only pages with at least 20 monthly page views are
+retained.\footnote{For further justification and to obtain the dataset,
+  see \blogpost} Stubs and disambiguation pages are also removed,
+leaving 463 thousand pages with a total of 482 million words.
+Punctuation marks and numbers were removed from the pages and all words
+were lower-cased.  Word- (or ``term-'') frequency $tf$ is described in overview
+in Table~\ref{word-occurrences-table}.  This base corpus is then
+modified as described in Sections \ref{WFVE} and \ref{CNVE}.  For
+recognisability, the tokens inserted into the corpus are upper-cased.
 
-This base corpus is then modified as described in sections \ref{WFVE} and \ref{CNVE}.
-Punctuation and numbers were removed and the corpus lower-cased prior to the modifications.
-For recognisability, the tokens introduced into the corpus during the modification are uppercased.
+\begin{subsection}{Word2vec}\label{word2vec}
+Word2vec, a feedforward neural network with a single hidden
+layer, learns word vectors from word co-occurrences in an unsupervised
+manner.  Word2vec comes in two versions.  In the continuous bag-of-words
+(CBOW) model, the words appearing around a target word serve as input.
+That input is projected linearly onto the hidden layer and the networks
+then attempts to predict the target word on output.  Training is
+achieved through backpropagation.  The word vectors are encoded in the
+weights of the first synaptic layer, ``syn0''.  The weights of the
+second synaptic layer (``syn1neg'', in the case of negative sampling) are typically discarded.  In the other model,
+called skip-gram, target and context words swap places, so that the
+target word now serves as input, while the network attempts to predict
+the context words on output.
 
-For simplicity, only the word2vec CBOW word embedding with a single set of hyperparameters is considered.
-Specifically, a $100$-dimensional CBOW model is trained using negative sampling with 5 negative samples, a window size of $10$, a minimum word occurrence of $128$, and $10$ passes through the corpus.
-Sub-sampling was not used so that the influence of word frequency could be more clearly discerned.
-Identical experimental results were obtained using hierarchical softmax, but these are omitted for succinctness.
-The relatively high minimum word cut-off is chosen to ensure that word vectors, in all but degenerate cases, receive a sufficient number of gradient updates to be meaningful.
-This cut-off results in a vocabulary of $81117$ words (only unigrams were considered).
+For simplicity only the word2vec CBOW word embedding with a single set
+of hyperparameters is considered.  Specifically, a CBOW model with a
+hidden layer of size 100 is trained using negative sampling with 5
+negative samples, a window size of $10$, a minimum word occurrence of
+$128$, and $10$ passes through the corpus.  Sub-sampling was not used so
+that the influence of word frequency could be more clearly discerned.
+Similar experimental results were obtained using hierarchical softmax,
+but these are omitted for succinctness.  The relatively high
+low-frequency cut-off is chosen to ensure that word vectors, in all but
+degenerate cases, receive a sufficient number of gradient updates to be
+meaningful.  This frequency cut-off results in a vocabulary of $81117$
+words (only unigrams were considered).
 
-The most recent revision of word2vec was used\footnote{SVN revision 42, see \url{http://word2vec.googlecode.com/svn/trunk/}}.
-The source code for performing the experiments is made available on GitHub\footnote{\url{https://github.com/benjaminwilson/word2vec-norm-experiments}}.
+The most recent revision of word2vec was used.\footnote{SVN revision 42,
+  see \url{http://word2vec.googlecode.com/svn/trunk/}} The source code
+for performing the experiments is made available on
+GitHub.\footnote{\url{https://github.com/benjaminwilson/word2vec-norm-experiments}}
 
-Word vectors are then derived from the first layer of synaptic weights (``syn0'') and the second layer of weights (``syn1neg'') are discarded.
-
-\begin{table}\label{word-occurrences-table}
+\begin{table}
 	\input{words-occurrences.tex}
-	\caption{ Number of words, by frequency band, as observed in the unmodified corpus.  }
+	\caption{ Number of words, by frequency band, as observed in the
+          unmodified corpus.  }
+	\label{word-occurrences-table}
 \end{table}
+\end{subsection}
 
 \begin{subsection}{Replacement procedure}\label{replacement-procedure}
-In the experiments, tokens are introduced into corpus via a replacement procedure.
-To illustrate this procedure, consider a word, e.g. \word{cat}.
-For each occurrence of this word, a sample $i$, $1 \leqslant i \leqslant n$ is drawn from a truncated geometric distribution, and that occurrence of \word{cat} is replaced with \word{CAT\_i}.
-Thus the word \word{cat} is replaced throughout the corpus by a family of tokens with varying frequencies but approximately the same co-occurrence distribution as \word{cat}.
+In the experiments detailed below, we modify the corpus in a controlled
+manner by introducing tokens into the corpus via a replacement procedure
+as follows.  Consider a word, say \word{cat}.  For each occurrence of
+this word, a sample $i$, $1 \leqslant i \leqslant n$ is drawn from a
+truncated geometric distribution, and that occurrence of the word
+\word{cat} is replaced with the token \word{CAT\_i}.  In this way, the
+word \word{cat} is replaced throughout the corpus by a family of tokens
+with varying frequencies but approximately the same co-occurrence
+distribution as \word{cat}.  That is, all these tokens are used in
+roughly the same contexts as the original word.
 
-The geometric distribution is truncated in order to limit the number of tokens introduced into the corpus.
-For any ratio $0 < \lambda < 1$ and maximum value $n > 0$, the truncated geometric distribution is given by
-$$ P_{\lambda, n} (i) = \frac{\lambda^{i-1} \cdot (1-\lambda)}{(1 - \lambda^n)}, \qquad 1 \leqslant i \leqslant n.$$ 
-This is the distribution over $i = 1, 2, \dots n$ for which the probabilities decay exponentially base $\lambda$.
-Of course, other distributions might equally have been chosen for the experiments.
+The geometric distribution is truncated in order to limit the number of
+tokens inserted into the corpus.  For any choice $0 < p < 1$ and
+maximum value $n > 0$, the truncated geometric distribution is given by
+the probability density function
+\begin{equation} 
+\label{distro}
+P_{p, n} (i) = \frac{p^{i-1}  (1-p)}{1 - p^n}, \qquad 1
+\leqslant i \leqslant n.
+\end{equation} 
+The factor in the denominator, which tends to unity in the limit $n \to
+\infty$, assures proper normalisation.  We have chosen this distribution
+because the probabilities decay exponentially base $p$ as a function of
+$i$.  Of course, other distributions might equally well have been chosen
+for the experiments.
 \end{subsection}
@@
 
-\begin{section}{Word frequency variation experiment}\label{WFVE}
-In this experiment, families of tokens are introduced into the corpus.
-The tokens in each family vary in frequency but share a common co-occurrence distribution.
-After model training, comparison of the resulting word vectors associated to these tokens reveals the independent effect of frequency variation on the word embedding.
+\begin{section}{Varying word frequency}\label{WFVE}
+In this first experiment, we investigate the effect of word frequency on
+the word embedding.  Using the replacement procedure, we introduce a
+small number of families of tokens into the corpus.  The tokens in each
+family vary in frequency but, replacing a single word, all share a
+common co-occurrence distribution.  This allows us to study the role of
+word frequency in isolation, everything else being kept equal.  We
+consider two types of tokens.
 
 \begin{subsection}{Tokens derived from existing words}\label{WFVEexisting}
-A small set of words from the unmodified corpus is chosen uniformly at random from the vocabulary.
-In order that the introduced tokens do not have too low a frequency, only words which occur at least 10 thousand times are chosen.
-To this set of words, the word \word{the} is added, in order to include a high-frequency stopword.
-Table \ref{word-frequency-counts} lists the words chosen for this experiment along with their frequencies in the corpus.
+%
+We choose uniformly at random a small number of words from the
+unmodified vocabulary for our experiment.  In order that the inserted
+tokens do not have too low a frequency, only words which occur at least
+10 thousand times are chosen.  We also include the high-frequency
+stopword \word{the} for comparison.  Table~\ref{word-frequency-counts}
+lists the words chosen for this experiment along with their frequencies.
 
-The replacement procedure of subsection \ref{replacement-procedure} is then performed for each of these words, using a geometric rate of decay of $\lambda = 1/2$, and maximum value $n=20$.
-This value of $\lambda$ is one of a range of values that ensure that, for each word, multiple tokens will be introduced with a number of occurrences sufficient to survive the chosen word2vec minimum occurrence cut-off of 128.
-A maximum value $n=20$ suffices for this choice of $\lambda$, since $2^{20 + \log_2{128}}$ exceeds the maximum occurrence of any token in the corpus. 
-Figure \ref{fig:word-frequency-experiment-text-cat} illustrates the effect of these modifications on a sample text, with a family of tokens \word{CAT\_i}, derived from the word \word{cat}.
-Notice that the word \word{cat} has been replaced throughout with the tokens \word{CAT\_i}.
+The replacement procedure of Section ~\ref{replacement-procedure} is then
+performed for each of these words, using a geometric decay rate of $p =
+\tfrac{1}{2}$, and maximum value $n=20$, so that the 1st token is
+inserted with a probability of about $0.5$, the 2nd with a probability of
+about $0.25$, and so on.  This value of $p$ is one of a range of values
+that ensure that, for each word, multiple tokens will be inserted with a
+number of occurrences sufficient to survive the low-frequency cut-off of
+128.  A maximum value $n=20$ suffices for this choice of $p$, since
+$2^{20 + \log_2{128}}$ exceeds the maximum occurrence count of any token in
+the corpus.  Figure~\ref{fig:word-frequency-experiment-text-cat}
+illustrates the effect of these modifications on a sample text, with a
+family of tokens \word{CAT\_i}, derived from the word \word{cat}.
+Notice that all occurences of the word \word{cat} have been replaced
+with the tokens \word{CAT\_i}.
 
-\begin{table}\label{word-frequency-counts}
+\begin{table}
 	\input{word-frequency-experiment-counts.tex}
-	\caption{Occurrence counts for words chosen for the word frequency experiment. }
+	\caption{Term frequency $tf$ of words chosen for the word
+          frequency experiment. }
+	\label{word-frequency-counts}
 \end{table}
 
-\begin{figure}\label{fig:word-frequency-experiment-text-cat}
+\begin{figure}
 	\begin{mdframed}
@@
 	\end{mdframed}
-	\caption{A text modified for the word frequency experiment as per subsection \ref{WFVEexisting}, where the
-	word \word{cat} was chosen, $\lambda=0.5$ and $n=20$.}
+	\caption{Example sentences modified in the word frequency
+          experiment as per Section ~\ref{WFVEexisting}, where the word
+          \word{cat} is replaced with tokens \word{CAT\_i} using the
+          truncated geometric distribution (\ref{distro}) with
+          $p=\tfrac{1}{2}$ and $n=20$.}
+	\label{fig:word-frequency-experiment-text-cat}
 \end{figure}
@@
 
-\begin{subsection}{Tokens derived from an introduced, meaningless word}\label{WFVEmeaningless}
-	A high-frequency, purely meaningless word is included in the experiment for comparison.
-	We choose to introduce a new, entirely meaningless token \word{VOID} into the corpus, rather than choose an existing word whose meaninglessness is only supposed.
-	To achieve this, the token is interspersed uniformly at random throughout the corpus so that its global frequency is $f = 0.005$.
-	The co-occurrence distribution of \word{VOID} thus coincides with the global frequency distribution.
-	The replacement procedure is then performed for \word{VOID}, using the same values for $\lambda$ and $n$ as above.
-	figure \ref{fig:word-frequency-experiment-text-void} shows the effect of this modifications on a sample text ($f = 0.05$ is used there for illustrative purposes).
+\begin{subsection}{Tokens derived from an artificial, meaningless word}\label{WFVEmeaningless}
+%
+Whereas the tokens introduced above all replace an existing word that
+carries a meaning, we now include for comparison a high-frequency,
+meaningless word.  We choose to introduce an artificial, entirely
+meaningless word \word{VOID} into the corpus, rather than choose an
+existing (stop)word whose meaninglessness is only supposed.  To achieve
+this, we insperse the word uniformly at random throughout the corpus so
+that its global frequency is $0.005$.  The co-occurrence
+distribution of \word{VOID} thus coincides with the global word
+distribution.  The replacement procedure is then performed for this
+word, using the same values for $p$ and $n$ as above.
+Figure~\ref{fig:word-frequency-experiment-text-void} shows the effect of
+these modifications on a sample text, where a higher global frequency of $0.05$ is used instead
+for illustrative purposes.  Although the probability of inserting token
+\word{VOID\_1} is twice that of iserting token \word{VOID\_2} and four
+times that of inserting token \word{VOID\_3}, these three tokens all
+appear once in this small piece of text.
 
-\begin{figure}\label{fig:word-frequency-experiment-text-void}
+\begin{figure}
 	\begin{mdframed}
@@
 	\end{mdframed}
-	\caption{A text modified for the word frequency experiment as per subsection \ref{WFVEmeaningless}, where $\lambda=0.5$ and $n=20$.
-For illustrative purposes, the token \word{VOID} was here interspersed with a global frequency of $0.05$.}
+	\caption{The same example sentences as in
+          Figure ~\ref{fig:word-frequency-experiment-text-cat} where
+          instead of the word \word{cat} now the meaningless word
+          \word{VOID} is replaced with tokens \word{VOID\_i}.  For
+          illustrative purposes, the meaningless word \word{VOID} was
+          here interspersed with a global frequency of $0.05$.}
+	\label{fig:word-frequency-experiment-text-void}
 \end{figure}
@@
 
-\subsection{Results}
-\textcolor{red}{rework}
-For well-learned representations, word frequency has no effect on word vector direction.
-Figure \ref{word-frequency-experiment-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus.
-Recall that the cosine similarity measures the extent to which two vectors have the same direction, taking a maximum value of $1$ and a minimum value of $-1$.
-Notice that the word vectors associated to tokens derived from a word other that \word{the} and \word{VOID} have the same direction.
-The same is true of the word vectors associated to the \word{VOID\_i} for $i$ \textcolor{red}{below $7$ or $8$}, but that the word vectors \word{VOID\_i} gradually diverge directionally for higher values of $i$.
-This demonstrates the difficulty of learning the co-occurrence distribution of $\word{VOID}$, the global frequency distribution.
-Because of its noisiness, a higher number of samples (occurrences) is required to learn this co-occurrence distribution.
-The same holds, but to a less pronounced extent, for the co-occurrence distribution of the stopword \word{the}.
-\textcolor{red}{The poorly learned representations for \word{the\_i} for $i = 15, \dots $ and for \word{VOID\_i} for $i = 7, \dots$ are therefore disregarded.}
+\subsection{Experimental results}
+We next present the results of the word frequency experiment. We
+consider the effect of word frequency on the direction and on the length
+of word vectors separately.
 
-\textcolor{red}{include a table with nearest neighbours in the normal vocab for the to support this argument}
-Table \ref{word-frequency-experiment-cosine-similarity}
+\subsubsection{Word frequency and vector direction}\label{WFVE-direction}
+%
+Figure~\ref{word-frequency-experiment-heatmap} shows the cosine
+similarity of pairs of vectors representing some of the tokens used in
+this experiment.  Recall that the cosine similarity measures the extent
+to which two vectors have the same direction, taking a maximum value of
+$1$ and a minimum value of $-1$.  The number of different tokens
+associated with an experiment word is the number of times that its
+frequency can be halved and remain above the low-frequency cut-off of
+$128$.
 
-It remains, therefore, to consider the effect of frequency variation on word vector length.
-Figure \ref{fig:word-frequency-experiment-graph}, on the other hand, shows this relationship for individual words, both for the word vectors, ``syn0'' and for the discarded second layer vectors ``syn1neg'' (included for completeness).
-Each line corresponds to a single word, and the points on each line indicate the frequency and vector length for the tokens derived from the corresponding word.
-For example, the six points on the line corresponding to \word{protestant} are labeled, from right to left, by the tokens \word{PROTESTANT\_1}, \word{PROTESTANT\_2}, \dots, \word{PROTESTANT\_6}.
-The number of points on the line is determined by the frequency of the original word.
-For example, the frequency of the \word{protestant} can be halved at most $6$ times and remain about the minimum frequency cut-off.
+Consider first the words other than \word{the} and \word{VOID}.  These
+words typically occur in a limited number of distinct contexts, and
+so a limited number of samples is required to capture their
+co-occurrence distributions.  We therefore expect that the vectors
+representing the tokens derived from any of these words can be properly
+learned from the modified corpus, even though their combined number of
+occurences equals the term frequency of that word in the original
+corpus.  The red blocks in Figure ~\ref{word-frequency-experiment-heatmap}
+clearly demonstrate that the direction of these vectors is constant for
+each family of tokens.  In other words, the direction of these vectors
+is only determined by the original word and is independent of how often
+that words appear, provided sufficient samples have been processed for
+the word vector to settle in its equilibrium direction.
 
-Because all the points on a line share a co-occurrence distribution (only the frequency varies), figure \ref{fig:word-frequency-experiment-graph} demonstrates conclusively that length does indeed depend on frequency directly, and not only via some correlation between word frequency and word significance.
-Moreover, this relationship is linear for any particular word.
-Notice also that the relative positions of the norms of the word vectors associated to the test words is independent of the frequency band.
-Therefore these lines trace lines along which a frequency-adjusted norm would be constant.
+The artificial word \word{VOID}, on the other hand, occurs by
+construction in every context.  A much higher number of samples is
+therefore required to capture its co-occurrence distribution, and
+thereby to learn its word vector.  The same is true, but to a lesser
+extent, for the stopword \word{the}.  The difficulty of learning the
+vector representing \word{VOID\_i} or \word{THE\_i} increases with $i$,
+for the tokens occur increasingly less often.  This is consistent with
+Figure ~\ref{word-frequency-experiment-heatmap}, where it is apparent that
+the vectors representing \word{VOID\_i} for small $i$, say $i = 1,
+\dots, 5$, roughly share a common direction, while this is no longer
+true for higher values of $i$.  The same trend is apparent, although
+less extremely so, for the word \word{the}.  The vectors for the tokens
+\word{THE\_i} share a common direction for values of $i$ as high as
+$14$, before they start changing direction.  We conclude that these
+deviations are artifacts of poor learning due to an insufficient number
+of occurrences of these tokens.
+Table~\ref{word-frequency-experiment-cosine-similarity}, showing the
+most similar words to each token \word{THE\_i}, supports this further.  In the
+discussions below, the tokens \word{VOID\_i} and
+\word{THE\_i} whose vectors have a cosine similarity less than $0.8$
+with \word{VOID\_1} and \word{THE\_1}, respectively are therefore
+discarded.
 
-Notice also that the word vector for \word{VOID} is non-zero.
+From these results, we conclude that word frequency has no effect on the
+direction of a word vector, given a sufficient number of samples to
+learn the co-occurrence distribution of the word.
+%For well-learned representations, word frequency has no effect on word
+%vector direction.
 
-\begin{figure}\label{word-frequency-experiment-heatmap}
+\subsubsection{Word frequency and vector length}
+%
+We next consider the effect of frequency on word vector length.
+Figure~\ref{fig:word-frequency-experiment-graph} shows this relation for
+individual words, both for the word vectors, represented by the weights
+of the first synaptic layer, syn0, in the word2vec neural network, and
+for the vectors represented by the weights of the second synaptic layer,
+syn1neg.  We include the latter, which are typically ignored, for
+completeness.  Each line corresponds to a single word, and the points on
+each line indicate the frequency and vector length of the tokens derived
+from that word.  For example, the six points on the line corresponding
+to the word \word{protestant} are labeled, from right to left, by the
+tokens \word{PROTESTANT\_1}, \word{PROTESTANT\_2}, \dots,
+\word{PROTESTANT\_6}.  Again, the number of points on the line is
+determined by the frequency of the original word.  For example, the
+frequency of the word \word{protestant} can be halved at most $6$ times
+so that the number of occurences of the last token is still above the
+low-frequency cut-off.  Because all the points on a line share the same
+co-occurrence distribution, the left panel in
+Figure ~\ref{fig:word-frequency-experiment-graph} demonstrates conclusively
+that length does indeed depend on frequency directly.  Moreover, this
+relation is seen to be approximately linear for each word considered.
+Notice also that the relative positions of the lengths of the word vectors associated
+with the experiment words are roughly independent of the frequency band,
+i.e., the plotted lines rarely cross.
+
+Notice the inflection of the syn1neg curve in
+Figure ~\ref{fig:word-frequency-experiment-graph} for \word{the} at
+approximately $tf=10^6$.  We have no explanation of this.  Also observe
+that the lengths of the vectors representing the meaningless tokens
+\word{VOID\_i} are approximately constant.  Since we already found the
+direction to be also constant, it is sensible to speak of the word
+vector of \word{VOID} irrespective of its frequency.  In particular, the
+token \word{VOID\_1} may be taken as an approximation.
+
+\begin{figure}
 	\includegraphics[scale=0.5]{word-frequency-experiment-heatmap}
-	\caption{
-	Heatmap of the cosine similarity of the vectors associated to the
-	tokens introduced into the corpus in the co-occurrence noise experiment
-	(words other than \word{the} and \word{VOID} chosen randomly).
-	}
+	\caption{ Heatmap of the cosine similarity of the vectors
+          representing some of the tokens used in the word frequency
+          experiment.  The words other than \word{the} and \word{VOID}
+          were chosen randomly.  }
+	\label{word-frequency-experiment-heatmap}
 \end{figure}
 
-\begin{table}\label{word-frequency-experiment-cosine-similarity}
+\begin{table}
 	\input{cosine-similarity.tex}
-	\caption{
-		The most similar words in the original vocabulary to the tokens
-		introduced into the corpus associated to the word \word{the},
-		and their cosine similarity with the most frequent such token,
-		\word{THE\_1}.  It is apparent from the nearest neigbours that
-		the word vectors of the low frequency tokens have not been
-		adequately trained (there were inadequate samples).
-	}
+	\caption{ Words in the original vocabulary most similar to the
+          tokens \word{THE\_i}, and their cosine similarity with the
+          most frequent such token, \word{THE\_1}.  It is apparent from
+          the nearest neigbour list that the vectors of the
+          low-frequency tokens have not been adequately trained.  }
+	\label{word-frequency-experiment-cosine-similarity}
 \end{table}
 
-\begin{sidewaysfigure*}\label{fig:word-frequency-experiment-graph}
+\begin{sidewaysfigure*}
 	\centering{\includegraphics[scale=0.6]{word-frequency-experiment-graph}}
-	\caption{
-	Vector length vs frequency for the words chosen for the word frequency
-	experiment.  For each word, tokens of varying frequency but with the
-	co-occurrence distribution of that word were introduced into the
-	corpus, as described in section \ref{WFVE}.
-	The word vectors are obtained from the first synaptic layer, ``syn0''.
-	The second layer ``syn1neg'' is included for completeness.
-	}
+	\caption{ Vector length vs. frequency for tokens derived from a
+          few words chosen at random.  For each word, tokens of varying
+          frequency but with the co-occurrence distribution of that word
+          were inserted into the corpus, as described in
+          Section ~\ref{WFVE}.  The vectors are obtained from the first
+          synaptic layer, syn0, of the word2vec neural network.  The
+          vectors obtained from the second layer, syn1neg, are
+          included for completeness.  }
+	\label{fig:word-frequency-experiment-graph}
 \end{sidewaysfigure*}
-
 \end{section}
 
-\begin{section}{Co-occurrence noise variation experiment}\label{CNVE}
-Here, and throughout, the noise distribution is taken to be the global word frequency distribution.
-Thus noise can be added to the co-occurrence distribution of a word by interspersing occurrences of that word uniformly at randomly throughout the corpus.
-A small set of words is chosen from the unmodified corpus in the same manner as in section \ref{WFVE}.
+\begin{section}{Varying co-occurrence noise}\label{CNVE}
+This second experiment is complementary to the first.  Wheras in the
+first experiment we studied the effect of word frequency on word vectors
+for fixed co-occurence, we next study the effect of co-occurence for
+fixed frequency.
 
-For each of the chosen words, the replacement procedure of subsection \ref{replacement-procedure} is performed using a geometric rate of decay of $\lambda = 5/6$, and truncating at $n=8$.
-Subsequently, for every replacement token (e.g.
-\word{CAT\_i}) random occurrences of this token are interspersed uniformly at random throughout the corpus, such that the frequency of the replacement token is restored to that of the original token \word{cat}.
-For example, if the original token \word{cat} occurred $1000$ times, then after the replacement procedure, \word{CAT\_2} occurs approximately $694$ times, so a further (approximately) $306$ random occurrences of \word{CAT\_2} are interspersed throughout the corpus.
-Thus token \word{cat} is removed from the corpus, and a family of tokens \word{CAT\_i}, $1 \leqslant i \leqslant 8$ are introduced.
-These tokens all have the same frequency as \word{cat}, but their co-occurrence distributions, while based on that of \word{cat}, have an increasing amount of noise.
+\subsection{Generating noise}
+We do so again in a controlled manner by adding noise to the
+co-occurence distribution.  We take the noise distribution to be the
+global word distribution, so that noise can be simply added to the
+co-occurrence distribution of a word by interspersing occurrences of
+that word uniformly at random throughout the corpus.  A word that is
+consistently used in a distinctive context in the unmodified corpus thus
+appears in the modified corpus also in completely unrelated contexts.
+As in Section \ref{WFVE}, we choose a small number of words from the
+unmodified corpus for this experiment.  Table
+\ref{cooccurrence-noise-words} lists the words chosen, along with their
+frequencies in the corpus.
 
-Figure \ref{fig:co-occurrence-noise-experiment-text} illustrates the effect of this modification, in the case where the only word chosen is \word{cat}.
-The original text in this case concerned both cats and dogs.
-Notice that the word \word{cat} has been replaced entirely in the cats section by occurrences of \word{CAT\_i} and moreover that occurrences of these same tokens appear also in the dogs section.
-These occurrences (and additionally, with probability, some occurrences from the cats section) are ``noise occurrences''.
-Table \ref{cooccurrence-noise-words} lists the words chosen for the experiment, along with their frequencies in the corpus.
+For each of these words, the replacement procedure of Subsection
+\ref{replacement-procedure} is performed using a geometric rate of decay
+of $p = \tfrac{5}{6}$, and truncating the distribution at $n=8$.  For
+every replacement token (e.g. \word{CAT\_i}), occurrences of this
+token are subsequently interspersed uniformly at random throughout the
+corpus, such that the frequency of the replacement token is restored to
+that of the original word \word{cat}.  For example, if the original word
+\word{cat} occurred $1000$ times, then after the replacement procedure,
+\word{CAT\_2} occurs approximately $694$ times, so a further
+(approximately) $306$ random occurrences of \word{CAT\_2} are
+interspersed throughout the corpus.  Thus the word \word{cat} is removed
+from the corpus, and a family of tokens \word{CAT\_i}, $1 \leqslant i
+\leqslant 8$ is introduced.  These tokens all have the same frequency as
+the word \word{cat}, but their co-occurrence distributions, while based
+on that of \word{cat}, have an increasing amount of noise.
 
-The proportion of the occurrences of \word{CAT\_i} that arose by replacing \word{cat} is $P_{5/6, 8}(i)$, and the proportion arising from the random interspersal of this token throughout the corpus is $r_i := (1 - P_{5/6, 8}(i))$.
-The particular value of $\lambda$ (large compared to that of section \ref{WFVE}) was chosen so that the proportion of noise $r_i$ did not increase too quickly with $i$.
-These proportions are the sampling points on our axis of variation, and this value of $\lambda$ (together with the choice of $n$) ensures a reasonable spread.
-Note that other parameter values (or indeed other distributions) could equally have been used.
+Figure \ref{fig:co-occurrence-noise-experiment-text} illustrates the
+effect of this modification in the case where the only word chosen is
+\word{cat}.  The original text in this case concerned both cats and
+dogs.  Notice that the word \word{cat} has been replaced entirely in the
+cats section by \word{CAT\_i} and, moreover, that these same tokens
+appear also in the dogs section. These occurrences (and additionally,
+with probability, some occurrences from the cats section) constitute
+noise.
 
-\begin{table}\label{cooccurrence-noise-words}
+The proportion of occurrences of the token \word{CAT\_i} that arise by
+replacing the word \word{cat} is $P_{5/6, 8}(i)$, and the proportion
+that arise from randomly interspersing this token throughout the corpus
+is $r_i := 1 - P_{5/6, 8}(i)$.  The particular value of $p$ (large
+compared to that of section \ref{WFVE}) was chosen so that the
+proportion of noise $r_i$ did not increase too quickly with $i$.  These
+proportions are the sampling points on our axis of variation, and this
+value of $p$ (together with the choice of $n$) assures a reasonable
+spread.  Other parameter values (or indeed other distributions) could,
+of course, have been used equally well.
+
+\begin{table}
 	\input{noise-cooccurrence-experiment-counts.tex}
 	\label{fig:co-occurrence-noise-counts}
-	\caption{Occurrence counts for words chosen for the co-occurrence noise experiment. }
+	\caption{Term frequency $tf$ of words chosen for the
+          co-occurrence noise experiment. }
+	\label{cooccurrence-noise-words}
 \end{table}
@@
 	\end{mdframed}
-	\caption{A text modified for the co-occurrence noise experiment, where the word \word{cat} was chosen, $\lambda = 5/6$ and $n=3$.}
+	\caption{Example sentences modified for the co-occurrence noise
+          experiment, where the word \word{cat} was chosen for
+          replacement.  The tokens were generated using the distribution
+          (\ref{distro}) with $p = \tfrac{5}{6}$ and $n=3$.}
 	\label{fig:co-occurrence-noise-experiment-text}
@@
 
-\begin{subsection}{Results}
-Figure \ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity of the pairs of word vectors for the tokens introduced into the corpus in this experiment.
-Notice that the word vectors associated to tokens derived from the same word all share a common direction.
-We conclude that co-occurrence noise has no effect on word vector direction.
+\begin{subsection}{Experimental results}
+Figure~\ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity
+of pairs of vectors representing some of the tokens used in this
+experiment.  The figure clearly demonstrates that vectors representing
+tokens derived from the same word all share a common direction.  This
+brings us to the conclusion that co-occurrence noise has no discernable effect on
+the direction of word vectors.
 
-On the other hand, figure \ref{fig:co-occurrence-noise-graph} shows that word vector length varies directly and linearly with the proportion $r$ of noise in the co-occurrence distribution.
-This figure motivates an interpretation of vector length, within a sufficiently narrow frequency band, as a measure of the absence of co-occurrence noise, that is, of the extent to which a word determines a context.
+%While the word2vec neural network
+%samples co-occurrence noise, the direction of the vector representing
+%the target word changes, but these changes cancel out, so that the
+%direction remains fixed on average.
 
-\begin{figure}\label{fig:co-occurrence-noise-heatmap}
+The left panel in Figure ~\ref{fig:co-occurrence-noise-graph}, on the other hand, reveals that vector length varies more or less
+linearly with the proportion $r$ of noise in the co-occurrence
+distribution.  This figure motivates our interpretation of vector
+length, within a sufficiently narrow frequency band, as a measure of the
+absence of co-occurrence noise, or put differently, of the extent to
+which a word carries the meaning of a distinctive context.
+
+When read from right to left, this figure can be reinterpreted and
+related to the corresponding figure in
+Figure ~\ref{fig:word-frequency-experiment-graph} as follows.  The
+co-occurrence noise we generated appears to be white noise whose effects
+cancel out on average.  The horizontal axis can then be reinterpreted as
+denoting term frequency and a similar picture as in
+Figure ~\ref{fig:word-frequency-experiment-graph} emerges, namely that the
+word vector length increases more or less linearly with term frequency.
+This can, however, not be the complete picture, for the vectors
+represented by the second synaptic layer, syn1neg, show a different
+behavior in the two figures.  In particular, the length of these vectors
+is seen from the right panel of Figure ~\ref{fig:co-occurrence-noise-graph}
+to be completely insensitive to noise, i.e., frequency in our
+reinterpretation, whereas the length of those shown in the right panel
+of Figure ~\ref{fig:word-frequency-experiment-graph} does depend on
+frequency.
+
+In this experiment, the introduced tokens are intermediate points on a path in the space of distributions over the vocabulary, beginning at the co-occurrence distribution of the word and ending at the global word frequency distribution, i.e.
+the co-occurrence distribution of \word{VOID}.
+Thus it is clear a priori that the word vectors for the introducted tokens are intermediate steps on a path from the word vector of the original word to that of \word{VOID} (recall from \ref{WFVE} that its vector is approximately independent of its frequency).
+While we observe from Figure \ref{fig:co-occurrence-noise-heatmap} that the word vectors for the tokens introduced for a word all share a common direction, this is only true to the extent that the word vector for \word{VOID} is close to zero.
+
+\begin{figure}
 	\includegraphics[scale=0.5]{cooccurrence-noise-heatmap}
-	\caption{
-	Heatmap of the cosine similarity of some of the vectors associated to the
-	tokens introduced into the corpus in the co-occurrence noise experiment
-	(four such words chosen at random).  The
-	largely red blocks demonstrate that the direction of the word vector
-	does not change when noise is added to the co-occurrence distribution.
-	}
+	\caption{ Heatmap of the cosine similarity of the vectors
+          representing some of the tokens used in the co-occurrence
+          noise experiment (the words were chosen at random).  The
+          largely red blocks demonstrate that the direction of the word
+          vectors does not change when noise is added to the
+          co-occurrence distribution.  }
+	\label{fig:co-occurrence-noise-heatmap}
 \end{figure}
 
-\begin{sidewaysfigure*}\label{fig:co-occurrence-noise-graph}
+\begin{sidewaysfigure*}
 	\includegraphics[scale=0.6]{cooccurrence-noise-graph}
-	\caption{
-	Vector length vs proportion of co-occurrence noise for words
-	chosen for the co-occurrence noise experiment.  For each word,
-	tokens of equal frequency but with increasing proportions of
-	co-occurrence noise were introduced into the corpus, as
-	described in section \ref{CNVE}.
-	The word vectors are obtained from the first synaptic layer, ``syn0''.
-	The second layer ``syn1neg'' is included for completeness.
-	}
+	\caption{ Vector length vs. proportion of co-occurrence noise
+          for words chosen for this experiment.  For each word, tokens
+          of equal frequency but with increasing proportion of
+          co-occurrence noise were inserted into the corpus, as
+          described in section \ref{CNVE}.  The word vectors are
+          obtained from the first synaptic layer, syn0.  The second
+          layer, syn1neg, is included for completeness.  }
+	\label{fig:co-occurrence-noise-graph}
 \end{sidewaysfigure*}
@@
 Questions pertaining to word2vec in particular arise naturally from the results of the experiments.
-Figures \ref{fig:frequency-norm-graph} and \ref{fig:co-occurrence-noise-graph}, for example, demonstrate the word vectors obtained from the first synaptic layer, ``syn0'', have very different properties from those that could be obtained from the second layer, ``syn1neg'', which are typically discarded.
+Figures \ref{fig:word-frequency-experiment-graph} and \ref{fig:co-occurrence-noise-graph}, for example, demonstrate the word vectors obtained from the first synaptic layer, ``syn0'', have very different properties from those that could be obtained from the second layer, ``syn1neg'', which are typically discarded.
 These differences warrant further investigation.
@@
 Thus the word vector of \word{VOID} is a special point in the feature space.
-Figure \ref{fig:frequency-norm-graph} shows that this point is not at the origin of the feature space (i.e. is not the zero vector).
+Figure \ref{fig:word-frequency-experiment-graph} shows that this point is not at the origin of the feature space (i.e. is not the zero vector).
 The origin, however, is implicitly the point of reference in the word2vec word similarity task.
 This raises the question of whether improved performance on the similarity task could be achieved by transforming the feature space (or modifying the model) such that the representation of pure noise (i.e.
-the vector for \word{VOID}) was at the origin of the transformed feature space.
+the vector for \word{VOID}) is at the origin of the transformed feature space.
+
+%\textcolor{red}{can it be argued that the feature space must be fundamentally euclidean, given that the path marked by the vectorisations of intermediate words is a straight line?
+i.e.
+if such paths maybe considered as ``geodesics'' between words, and we show that these geodesics appear as straight lines when an euclidean metric is imposed, then suggests that the euclidean metric is in fact a good choice.
+what would happen if we chose another metric?}
+
 \end{subsection}

### 2015-09-03 article_markup_wordcounts.py

@@ -2,7 +2,7 @@ import sys
 
 print r"""\begin{center}
 \begin{tabular}{l | r}
-word & \# occurrences \\
+word & tf \\
 \hline"""
 wcs = []
 for line in sys.stdin:

### 2015-09-03 article/main.tex
moderate adriaan changes; remove tf; varied plot markers

@@
 denotes the conditional probability $\p(w'|w)$ that a word $w'$ occurs in
-its context, i.e., given that $w$ occurred.  Most applications of word
+its context, i.e. given that $w$ occurred.  Most applications of word
 embeddings explore not the word vectors themselves, but relations
@@
 and varying levels of noise in their co-occurrence distributions.  By
-modelling the frequency and co-occurrence distributions of these tokens
+modeling the frequency and co-occurrence distributions of these tokens
 on existing words in the corpus, we are able to study their effect on
@@
 the limited number of contexts in which a word appears in the original
-corpus, we can increase this number by inspersing the word in arbitrary
+corpus, we can increase this number by interspersing the word in arbitrary
 contexts at random.  The word thus looses its significance in a
@@
 \end{equation} 
-where the sum is over the set of \textcolor{blue}{content?} words $w'$.  They set the
-window size to five words before and five words after the target word.
-To reduce the computational burden, they recorded co-occurrences of 690
-target words, randomly chosen from equally divided frequency bands, with
-only the $n=500$ most frequent content words.  As pointed out by the
+where the sum is over the set of \textcolor{blue}{content?} words $w'$.
+As pointed out by the
 authors, contextual distinctiveness is a measure of the amount of
@@
 word $w$ tends to appear in a wide variety of contexts, the distribution
-of the words it occurs with, i.e., the posterior distribution $\p(w'|w)$,
+of the words it occurs with, i.e. the posterior distribution $\p(w'|w)$,
 tends to diverge less from the prior distribution $\p(w')$, and its
@@
 posterior distribution tends to diverge more from the prior, and a
-larger value of contextual distinctiveness results.  As for
-word significance, contextual distinctiveness summarizes a word's
-contextual behavior in a single number. \textcolor{blue}{shorten?}
+larger value of contextual distinctiveness results.
 
@@
 on word co-occurrence, the authors study adjective-noun composites.
-They built a vocabulary from the 8k most frequent nouns and 4k most
-frequent adjectives in a large general language corpus and added 22k
-adjective-noun composites.  For each item in the vocabulary, they
-recorded their co-occurrence within the same sentence with any of the
-top 10k most frequent content words (nouns, adjectives or verbs) in the
-corpus, and contructed word embeddings (of dimension 300) through
+For each item in the vocabulary, they
+recorded their co-occurrence within the same sentence with 
+ frequent content words (nouns, adjectives or verbs) in the
+corpus, and constructed word embeddings via 
 singular value decomposition of the co-occurrence matrix.  The authors considered several models for
 constructing vectors of unattested adjective-noun composites, the two
-simplest being adding and component-wise multipying the adjective and
+simplest being adding and component-wise multiplying the adjective and
 noun vectors.  They hypothesized that the length of the vectors thus
@@
 It is shown moreover that the normalisation of the word vectors also prevents word co-occurrence noise from affecting the performance of the word similarity and word relationship tasks.
-
 On the other hand, recent theoretical work \cite{Arora2015} has approached the problem of explaining the so-called ``compositionality'' property exhibited by some word embeddings.
@@
 \begin{section}{Corpus and model}\label{corpus-and-model}
-Our training data is built from the Wikipedia datadump from October
+Our training data is built from the Wikipedia data dump from October
 2013.  To remove the bulk of robot-generated pages from the training
@@
 Punctuation marks and numbers were removed from the pages and all words
-were lower-cased.  Word- (or ``term-'') frequency $tf$ is described in overview
-in Table~\ref{word-occurrences-table}.  This base corpus is then
+were lower-cased.  Word occurrence counts are summarised 
+in Table \ref{word-occurrences-table}.  This base corpus is then
 modified as described in Sections \ref{WFVE} and \ref{CNVE}.  For
@@
 \begin{subsection}{Word2vec}\label{word2vec}
-Word2vec, a feedforward neural network with a single hidden
+Word2vec, a feed-forward neural network with a single hidden
 layer, learns word vectors from word co-occurrences in an unsupervised
@@
 then attempts to predict the target word on output.  Training is
-achieved through backpropagation.  The word vectors are encoded in the
+achieved through back-propagation.  The word vectors are encoded in the
 weights of the first synaptic layer, ``syn0''.  The weights of the
@@
 \begin{subsection}{Tokens derived from existing words}\label{WFVEexisting}
-%
 We choose uniformly at random a small number of words from the
@@
 10 thousand times are chosen.  We also include the high-frequency
-stopword \word{the} for comparison.  Table~\ref{word-frequency-counts}
+stopword \word{the} for comparison.  Table \ref{word-frequency-counts}
 lists the words chosen for this experiment along with their frequencies.
 
-The replacement procedure of Section ~\ref{replacement-procedure} is then
+The replacement procedure of Section \ref{replacement-procedure} is then
 performed for each of these words, using a geometric decay rate of $p =
@@
 $2^{20 + \log_2{128}}$ exceeds the maximum occurrence count of any token in
-the corpus.  Figure~\ref{fig:word-frequency-experiment-text-cat}
+the corpus.  Figure \ref{fig:word-frequency-experiment-text-cat}
 illustrates the effect of these modifications on a sample text, with a
 family of tokens \word{CAT\_i}, derived from the word \word{cat}.
-Notice that all occurences of the word \word{cat} have been replaced
+Notice that all occurrences of the word \word{cat} have been replaced
 with the tokens \word{CAT\_i}.
@@
 	\input{word-frequency-experiment-counts.tex}
-	\caption{Term frequency $tf$ of words chosen for the word
-          frequency experiment. }
+	\caption{Words chosen for the word frequency variation experiment, along with their occurrence counts in the unmodified corpus. }
 	\label{word-frequency-counts}
@@
 	\caption{Example sentences modified in the word frequency
-          experiment as per Section ~\ref{WFVEexisting}, where the word
+          experiment as per Section \ref{WFVEexisting}, where the word
           \word{cat} is replaced with tokens \word{CAT\_i} using the
@@
 \begin{subsection}{Tokens derived from an artificial, meaningless word}\label{WFVEmeaningless}
-%
 Whereas the tokens introduced above all replace an existing word that
@@
 existing (stop)word whose meaninglessness is only supposed.  To achieve
-this, we insperse the word uniformly at random throughout the corpus so
+this, we intersperse the word uniformly at random throughout the corpus so
 that its global frequency is $0.005$.  The co-occurrence
@@
 word, using the same values for $p$ and $n$ as above.
-Figure~\ref{fig:word-frequency-experiment-text-void} shows the effect of
+Figure \ref{fig:word-frequency-experiment-text-void} shows the effect of
 these modifications on a sample text, where a higher global frequency of $0.05$ is used instead
 for illustrative purposes.  Although the probability of inserting token
-\word{VOID\_1} is twice that of iserting token \word{VOID\_2} and four
+\word{VOID\_1} is twice that of inserting token \word{VOID\_2} and four
 times that of inserting token \word{VOID\_3}, these three tokens all
@@
 	\caption{The same example sentences as in
-          Figure ~\ref{fig:word-frequency-experiment-text-cat} where
+          Figure \ref{fig:word-frequency-experiment-text-cat} where
           instead of the word \word{cat} now the meaningless word
@@
 
-\subsection{Experimental results}
+\subsection{Experimental results}\label{WFVE-results}
 We next present the results of the word frequency experiment. We
@@
 \subsubsection{Word frequency and vector direction}\label{WFVE-direction}
-%
-Figure~\ref{word-frequency-experiment-heatmap} shows the cosine
+Figure \ref{word-frequency-experiment-heatmap} shows the cosine
 similarity of pairs of vectors representing some of the tokens used in
@@
 learned from the modified corpus, even though their combined number of
-occurences equals the term frequency of that word in the original
-corpus.  The red blocks in Figure ~\ref{word-frequency-experiment-heatmap}
+occurrences equals the term frequency of that word in the original
+corpus.  The red blocks in Figure \ref{word-frequency-experiment-heatmap}
 clearly demonstrate that the direction of these vectors is constant for
@@
 for the tokens occur increasingly less often.  This is consistent with
-Figure ~\ref{word-frequency-experiment-heatmap}, where it is apparent that
+Figure \ref{word-frequency-experiment-heatmap}, where it is apparent that
 the vectors representing \word{VOID\_i} for small $i$, say $i = 1,
@@
 of occurrences of these tokens.
-Table~\ref{word-frequency-experiment-cosine-similarity}, showing the
+Table \ref{word-frequency-experiment-cosine-similarity}, showing the
 most similar words to each token \word{THE\_i}, supports this further.  In the
@@
 learn the co-occurrence distribution of the word.
-%For well-learned representations, word frequency has no effect on word
-%vector direction.
 
 \subsubsection{Word frequency and vector length}
-%
 We next consider the effect of frequency on word vector length.
-Figure~\ref{fig:word-frequency-experiment-graph} shows this relation for
+Figure \ref{fig:word-frequency-experiment-graph} shows this relation for
 individual words, both for the word vectors, represented by the weights
@@
 frequency of the word \word{protestant} can be halved at most $6$ times
-so that the number of occurences of the last token is still above the
+so that the number of occurrences of the last token is still above the
 low-frequency cut-off.  Because all the points on a line share the same
 co-occurrence distribution, the left panel in
-Figure ~\ref{fig:word-frequency-experiment-graph} demonstrates conclusively
+Figure \ref{fig:word-frequency-experiment-graph} demonstrates conclusively
 that length does indeed depend on frequency directly.  Moreover, this
@@
 with the experiment words are roughly independent of the frequency band,
-i.e., the plotted lines rarely cross.
+i.e. the plotted lines rarely cross.
 
 Notice the inflection of the syn1neg curve in
-Figure ~\ref{fig:word-frequency-experiment-graph} for \word{the} at
-approximately $tf=10^6$.  We have no explanation of this.  Also observe
+Figure \ref{fig:word-frequency-experiment-graph} for \word{the} at
+approximately $10^6$.  We have no explanation of this.  Also observe
 that the lengths of the vectors representing the meaningless tokens
@@
           most frequent such token, \word{THE\_1}.  It is apparent from
-          the nearest neigbour list that the vectors of the
+          the nearest neighbour list that the vectors of the
           low-frequency tokens have not been adequately trained.  }
@@
           were inserted into the corpus, as described in
-          Section ~\ref{WFVE}.  The vectors are obtained from the first
+          Section \ref{WFVE}.  The vectors are obtained from the first
           synaptic layer, syn0, of the word2vec neural network.  The
@@
 \begin{section}{Varying co-occurrence noise}\label{CNVE}
-This second experiment is complementary to the first.  Wheras in the
+This second experiment is complementary to the first.  Whereas in the
 first experiment we studied the effect of word frequency on word vectors
-for fixed co-occurence, we next study the effect of co-occurence for
+for fixed co-occurrence, we next study the effect of co-occurrence for
 fixed frequency.
@@
 We do so again in a controlled manner by adding noise to the
-co-occurence distribution.  We take the noise distribution to be the
+co-occurrence distribution.  We take the noise distribution to be the
 global word distribution, so that noise can be simply added to the
@@
 is $r_i := 1 - P_{5/6, 8}(i)$.  The particular value of $p$ (large
-compared to that of section \ref{WFVE}) was chosen so that the
+compared to that of Section \ref{WFVE}) was chosen so that the
 proportion of noise $r_i$ did not increase too quickly with $i$.  These
@@
 	\label{fig:co-occurrence-noise-counts}
-	\caption{Term frequency $tf$ of words chosen for the
-          co-occurrence noise experiment. }
+	\caption{Words chosen for the co-occurrence noise experiment, along with the occurrence counts in the unmodified corpus. }
 	\label{cooccurrence-noise-words}
@@
 \begin{subsection}{Experimental results}
-Figure~\ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity
+Figure \ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity
 of pairs of vectors representing some of the tokens used in this
@@
 tokens derived from the same word all share a common direction.  This
-brings us to the conclusion that co-occurrence noise has no discernable effect on
+brings us to the conclusion that co-occurrence noise has no discernible effect on
 the direction of word vectors.
@@
 
-The left panel in Figure ~\ref{fig:co-occurrence-noise-graph}, on the other hand, reveals that vector length varies more or less
+The left panel in Figure \ref{fig:co-occurrence-noise-graph}, on the other hand, reveals that vector length varies more or less
 linearly with the proportion $r$ of noise in the co-occurrence
@@
 
-When read from right to left, this figure can be reinterpreted and
-related to the corresponding figure in
-Figure ~\ref{fig:word-frequency-experiment-graph} as follows.  The
-co-occurrence noise we generated appears to be white noise whose effects
-cancel out on average.  The horizontal axis can then be reinterpreted as
-denoting term frequency and a similar picture as in
-Figure ~\ref{fig:word-frequency-experiment-graph} emerges, namely that the
+%When read from right to left, this figure can be reinterpreted and
+%related to the corresponding figure in
+%Figure \ref{fig:word-frequency-experiment-graph} as follows.
+If the average effect of co-occurrence noise were neglible, then the horizontal axis could be reinterpreted as
+denoting term frequency and (reading the graph from right to left).
+Thus a similar picture to Figure \ref{fig:word-frequency-experiment-graph} would emerge, namely that the
 word vector length increases more or less linearly with term frequency.
-This can, however, not be the complete picture, for the vectors
+This can not, however, be the complete picture, for the vectors
 represented by the second synaptic layer, syn1neg, show a different
 behavior in the two figures.  In particular, the length of these vectors
-is seen from the right panel of Figure ~\ref{fig:co-occurrence-noise-graph}
-to be completely insensitive to noise, i.e., frequency in our
+is seen from the right panel of Figure \ref{fig:co-occurrence-noise-graph}
+to be completely insensitive to noise, i.e. frequency in our
 reinterpretation, whereas the length of those shown in the right panel
-of Figure ~\ref{fig:word-frequency-experiment-graph} does depend on
+of Figure \ref{fig:word-frequency-experiment-graph} does depend on
 frequency.
@@
 the co-occurrence distribution of \word{VOID}.
-Thus it is clear a priori that the word vectors for the introducted tokens are intermediate steps on a path from the word vector of the original word to that of \word{VOID} (recall from \ref{WFVE} that its vector is approximately independent of its frequency).
-While we observe from Figure \ref{fig:co-occurrence-noise-heatmap} that the word vectors for the tokens introduced for a word all share a common direction, this is only true to the extent that the word vector for \word{VOID} is close to zero.
+Thus it is clear a priori that the word vectors for the introduced tokens are intermediate steps on a path from the word vector of the original word to that of \word{VOID} (recall from Subsection \ref{WFVE-results} that word vector of \word{VOID} is approximately independent of its frequency).
+Thus, while we do indeed observe from Figure \ref{fig:co-occurrence-noise-heatmap} that the word vectors for the tokens introduced for a word all share a common direction, this is only true because the word vector for \word{VOID} is close to the origin.
 
@@
 \begin{section}{Discussion}\label{future-directions}
-
-\begin{subsection}{Controlled experiments for word embeddings}
 Our principle contribution has been to demonstrate that controlled experiments can be used to gain insight into a word embedding.
-Because they are achieved via modification of the training corpus, these experiments can be carried out for any word embedding (or indeed language model), requiring no knowledge or modification of the model implementation.
-
-It would, of course, be interesting to perform these experiments for other word embeddings other than word2vec CBOW and for different hyperparameters settings.
+These experiments can be carried out for any word embedding (or indeed language model), because they are achieved via modification of the training corpus.
+Indeed, the experiments require no knowledge or modification of the model implementation.
+It would naturally be of interest to perform these experiments for other word embeddings other than word2vec CBOW, such as skipgrams and GloVe, as well as for different hyperparameters settings.
 
@@
 The co-occurrence noise variation experiment described here would be a special case of such an experiment where one of the two words was \word{VOID}.
-\end{subsection}
 
-\begin{subsection}{Applications in other contexts}
 Further afield, it would be interesting to perform analogs of these experiments in other contexts where models are trained from co-occurrence data.
 For instance, in the context of e-Commerce, product purchase frequency and product co-occurrence noise can be varied by modifying the product-product co-occurrence matrix.
-\end{subsection}
 
-\begin{subsection}{Questions pertaining to word2vec}
 Questions pertaining to word2vec in particular arise naturally from the results of the experiments.
-Figures \ref{fig:word-frequency-experiment-graph} and \ref{fig:co-occurrence-noise-graph}, for example, demonstrate the word vectors obtained from the first synaptic layer, ``syn0'', have very different properties from those that could be obtained from the second layer, ``syn1neg'', which are typically discarded.
+Figures \ref{fig:word-frequency-experiment-graph} and \ref{fig:co-occurrence-noise-graph}, for example, demonstrate the word vectors obtained from the first synaptic layer, syn0, have very different properties from those that could be obtained from the second layer, syn1neg, which are typically discarded.
 These differences warrant further investigation.
@@
 
-%\textcolor{red}{can it be argued that the feature space must be fundamentally euclidean, given that the path marked by the vectorisations of intermediate words is a straight line?
+%\textcolor{red}{can it be argued that the feature space must be fundamentally Euclidean, given that the path marked by the vectorisations of intermediate words is a straight line?
 i.e.
-if such paths maybe considered as ``geodesics'' between words, and we show that these geodesics appear as straight lines when an euclidean metric is imposed, then suggests that the euclidean metric is in fact a good choice.
+if such paths maybe considered as ``geodesics'' between words, and we show that these geodesics appear as straight lines when an Euclidean metric is imposed, then suggests that the Euclidean metric is in fact a good choice.
 what would happen if we chose another metric?}
-
-\end{subsection}
 

### 2015-09-03 article/word-frequency-experiment-text-void.tex
moderate adriaan changes; remove tf; varied plot markers

--- 2015-08-18 article/word-frequency-experiment-text-void.tex
@@
-\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}
+\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned \textcolor{red}{VOID\_3} by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in \textcolor{red}{VOID\_1} having very large eye sockets\newline \textcolor{red}{VOID\_2} another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others}

### 2015-09-03 article_markup_wordcounts.py

@@ -2,7 +2,7 @@ import sys
 
 print r"""\begin{center}
 \begin{tabular}{l | r}
-word & tf \\
+word & \# occurrences \\
 \hline"""
 wcs = []
 for line in sys.stdin:

### 2015-09-03 build_images.py

@@ -75,8 +75,7 @@ for word in word_freq_experiment_words + [meaningless_token]:
 
 words = word_freq_experiment_words + [meaningless_token]
 fig = plt.figure(figsize=(16, 6))
-colormap = plt.cm.gist_ncar
-colorcycle = [colormap(i) for i in np.linspace(0, 0.9, len(words))]
+colorcycle = plt.cm.gist_rainbow(np.linspace(0, 1, len(words)))
 
 ax_syn0 = plt.subplot(131)
 ax_syn0.set_xlabel('# occurrences')
@@ -85,7 +84,7 @@ ax_syn0.set_xscale('log')
 ax_syn0.set_color_cycle(colorcycle)
 ax_syn0.set_title('syn0', y=1.04)
 
-ax_syn1neg = plt.subplot(132, sharex=ax_syn0, sharey=ax_syn0)
+ax_syn1neg = plt.subplot(132, sharex=ax_syn0)
 ax_syn1neg.set_xlabel('# occurrences')
 ax_syn1neg.set_color_cycle(colorcycle)
 ax_syn1neg.set_title('syn1neg', y=1.04)
@@ -94,7 +93,8 @@ def plot_for_word(ax, word, series, **kwargs):
     idxs = [build_experiment_token(word, i) for i in range(1, max(word_freq_experiment_ratio, word_freq_experiment_power_max) + 1)]
     x = stats.loc[idxs].occurrences
     y = series.loc[idxs]
-    return ax.plot(x, y, marker='o', **kwargs)[0]
+    marker = ['o', 's', 'D'][ord(word[0]) % 3]
+    return ax.plot(x, y, marker=marker, **kwargs)[0]
 
 lines = []
 for word in words:
@@ -104,6 +104,7 @@ for word in words:
     plot_for_word(ax_syn1neg, word, stats.L2_norm_syn1neg)
 
 ax_syn0.set_ylim(0, 45)
+ax_syn1neg.set_ylim(0, 25)
 
 _ = fig.legend(lines, words, bbox_to_anchor=(0.76, 0.56), loc='center', fontsize=14, frameon=False)
 plt.tight_layout()
@@ -115,8 +116,7 @@ plt.savefig('outputs/word-frequency-experiment-graph.eps')
 
 words = coocc_noise_experiment_words
 fig = plt.figure(figsize=(16, 6))
-colormap = plt.cm.gist_ncar
[…]

### 2015-09-03 functions.py
documentation fixmes

@@ -76,14 +76,14 @@ def truncated_geometric_sampling(word, ratio, max_value):
 
 def intersperse_words(interspersal_rates, f_in, f_out):
     """
-    Insperse words uniformly at random throughout the text in 'in_filename',
-    writing the result to 'out_filename'.  'interspersal_rates' is a dict
+    Insperse words uniformly at random throughout the text in file-like object 'f_in',
+    writing the result to file-like object 'f_out'.  'interspersal_rates' is a dict
     mapping words to the rate at which they should be interspersed, e.g.
         interspersal_rates = {'CAT_3': 0.004, 'MEANINGLESS': 0.0001}
 
     Our use of this function in the experiments assumes that total number of
     words remains essentially unchanged.
-    """ #FIXME update doc
+    """
     insertion_proba = sum(interspersal_rates.values())
     insertion_words = interspersal_rates.keys()
     relative_probas = np.array([interspersal_rates[word] for word in insertion_words]) / insertion_proba
@@ -103,11 +103,11 @@ def intersperse_words(interspersal_rates, f_in, f_out):
 
 def replace_words(word_sampler_dict, f_in, f_out):
     """
-    Performs a replacement procedure on the text in 'in_filename', writing the
-    results to 'out_filename'.  'word_sampler_dict' is a dict mapping words to
+    Performs a replacement procedure on the text in read from file-like object 'f_in', writing the
+    results to the file-like object 'f_out'.  'word_sampler_dict' is a dict mapping words to
     be replaced to functions (without arguments) that return their replacement.
     e.g. word_sampler_dict = {'cat': truncated_geometric_proba('cat', 0.5, 20)} 
-    """ #FIXME update doc
+    """
     for line in f_in:
         words_out = []
         for word in line.strip().split(' '):

### 2015-09-04 article/main.tex
moderate adriaan's modifications

@@
 \begin{section}{Introduction}
-The field of natural language processing (NLP) is presently undergoing rapid
-developments, not in the least triggered by the tremendous commercial
-promises it holds.  A timely research topic in NLP are word embeddings,
-where words from a vocabulary are represented by dense, real-valued
+Word embeddings represent words from the vocabulary as dense, real-valued
 vectors.  Instead of one-hot vectors that merely indicate the location
@@
 Shillcock \cite{mcdonald2001contextual}, who also based their work on word co-occurrences.
-They define contextual distinctiveness of a target word $w$ as the
-Kullback-Leibler divergence $\mathrm{D} \big(\p(w') ||\p(w'|w) \big)$
-between the distribution of context words, $\p(w')$, and the
-co-occurrence distributions of that target word $w$.  Specifically,
-\begin{equation} 
-\label{cd}
-\mathrm{D} \big(\p(w') ||\p(w'|w) \big) = \sum_{w'} \p(w'|w) \log_2
-\frac{\p(w'|w)}{\p(w')},
-\end{equation} 
-where the sum is over the set of \textcolor{blue}{content?} words $w'$.
-As pointed out by the
-authors, contextual distinctiveness is a measure of the amount of
+They define the contextual distinctiveness of a target word $w$ as the
+Kullback-Leibler divergence
+between the global word distribution $\p(w')$, and the
+co-occurrence distribution $\p(w' | w)$ of the target word $w$.
+So defined, contextual distinctiveness is a measure of the amount of
 information provided by word $w$ about its contexts of use.  More specifically, if a
 word $w$ tends to appear in a wide variety of contexts, the distribution
-of the words it occurs with, i.e. the posterior distribution $\p(w'|w)$,
+of the words it occurs with, i.e. the posterior distribution $\p(w' | w)$,
 tends to diverge less from the prior distribution $\p(w')$, and its
@@
 this, we intersperse the word uniformly at random throughout the corpus so
-that its global frequency is $0.005$.  The co-occurrence
+that its global occurrence rate is $0.005$.  The co-occurrence
 distribution of \word{VOID} thus coincides with the global word
@@
 Figure \ref{fig:word-frequency-experiment-text-void} shows the effect of
-these modifications on a sample text, where a higher global frequency of $0.05$ is used instead
+these modifications on a sample text, where a higher word occurrence rate of $0.05$ is used instead
 for illustrative purposes.  Although the probability of inserting token
@@
           illustrative purposes, the meaningless word \word{VOID} was
-          here interspersed with a global frequency of $0.05$.}
+          here interspersed with a occurrence rate of $0.05$.}
 	\label{fig:word-frequency-experiment-text-void}
@@
 %Figure \ref{fig:word-frequency-experiment-graph} as follows.
-If the average effect of co-occurrence noise were neglible, then the horizontal axis could be reinterpreted as
-denoting term frequency and (reading the graph from right to left).
+If the average effect of co-occurrence noise were neglible, then the horizontal axis of Figure \ref{fig:co-occurrence-noise-graph} could be reinterpreted as
+denoting word frequency by reading the graph from right to left.
 Thus a similar picture to Figure \ref{fig:word-frequency-experiment-graph} would emerge, namely that the
-word vector length increases more or less linearly with term frequency.
+word vector length increases more or less linearly with word frequency.
 This can not, however, be the complete picture, for the vectors

### 2015-09-07 article/main.tex
hopefully final round of changes

@@
 \begin{section}{Introduction}
+Word embeddings, or distributed representations of words, have been the subject
+of much recent research in the natural language processing and machine learning
+communities, demonstrating state-of-the-art performance on word similarity and
+word analogy tasks, amongst others.
 Word embeddings represent words from the vocabulary as dense, real-valued
@@
 ``contextual distinctiveness'' introduced earlier by McDonald and
-Shillcock \cite{mcdonald2001contextual}, who also based their work on word co-occurrences.
+Shillcock \cite{mcdonald2001contextual}.
 They define the contextual distinctiveness of a target word $w$ as the
 Kullback-Leibler divergence
-between the global word distribution $\p(w')$, and the
+between the unconditional word distribution $\p(w')$, and the
 co-occurrence distribution $\p(w' | w)$ of the target word $w$.
 So defined, contextual distinctiveness is a measure of the amount of
-information provided by word $w$ about its contexts of use.  More specifically, if a
-word $w$ tends to appear in a wide variety of contexts, the distribution
-of the words it occurs with, i.e. the posterior distribution $\p(w' | w)$,
-tends to diverge less from the prior distribution $\p(w')$, and its
+information provided by word $w$ about its contexts of use.
+More specifically, if a word $w$ tends to appear in a wide variety of contexts, then its co-occurrence distribution tends to diverge less from unconditional word distribution, and its
 contextual distinctiveness is accordingly small.  If, on the other hand,
-$w$ typically appears in only a small number of different contexts, the
-posterior distribution tends to diverge more from the prior, and a
-larger value of contextual distinctiveness results.
+$w$ typically appears in only a small number of different contexts, its co-occurrence distribution tends to diverge more from the unconditional word distribution, and a larger value of contextual distinctiveness results.
+In order to reduce the computational burden, the authors considered only $690$ target words, randomly chosen from equally divided frequency bands.
+Furthermore, the authors considered only co-occurrence with the $500$ most frequent content words.
+This significant restriction, necessary in their case in order to reduce the variance of the estimates of their measure of contextual distinctiveness, does not apply in our study.
 
@@
 vector length and the ``semantic deviance'' of an adjective-noun
-composite was studied empirically.  In that paper, which is also based
+composite was studied empirically.  
+In that paper, which is also based
 on word co-occurrence, the authors study adjective-noun composites.
-For each item in the vocabulary, they
-recorded their co-occurrence within the same sentence with 
- frequent content words (nouns, adjectives or verbs) in the
-corpus, and constructed word embeddings via 
-singular value decomposition of the co-occurrence matrix.  The authors considered several models for
-constructing vectors of unattested adjective-noun composites, the two
-simplest being adding and component-wise multiplying the adjective and
-noun vectors.  They hypothesized that the length of the vectors thus
-constructed can be used to distinguish acceptable and semantically
-deviant adjective-noun composites.  Using a few hundred adjective-noun
-composites selected by humans for evaluation, they found that deviant
-composites have a shorter vector than acceptable ones, in accordance
-with their expectation.
+They built a vocabulary from the 8k most frequent nouns and 4k most frequent adjectives in a large general language corpus and added 22k adjective-noun composites.
+For each item in the vocabulary, they recorded the co-occurrences with the top 10k most frequent content words (nouns, adjectives or verbs), and contructed word embeddings via singular value decomposition of the co-occurrence matrix \cite{landauer-dumais1997}.
+The authors considered several models for constructing vectors of unattested adjective-noun composites, the two simplest being adding and component-wise multiplying the adjective and noun vectors.
+They hypothesized that the length of the vectors thus constructed can be used to distinguish acceptable and semantically deviant adjective-noun composites.
+Using a few hundred adjective-noun composites selected by humans for evaluation, they found that deviant composites have a shorter vector than acceptable ones, in accordance with their expectation.
+In contrast to their work, our approach does not require extensive human evaluation.
 
-In the word2vec similarity and word relationship tasks, normalised word vectors are used.
-Word vector length is thus disregarded.
-In the word2vec forum\footnote{\googlegroup}, it has been observed that normalisation improves performance on these tasks since word vector length is related to word frequency.
-It remained unclear, however, whether word frequency was \textit{directly} related to word vector length or rather via correlated factors such as word significance.
-Our experimental approach demonstrates that a direct relationship indeed exists.
-It is shown moreover that the normalisation of the word vectors also prevents word co-occurrence noise from affecting the performance of the word similarity and word relationship tasks.
-On the other hand, recent theoretical work \cite{Arora2015} has approached the problem of explaining the so-called ``compositionality'' property exhibited by some word embeddings.
+Recent theoretical work \cite{Arora2015} has approached the problem of explaining the so-called ``compositionality'' property exhibited by some word embeddings.
 In that work, unnormalised vectors are used in their model of the word relationship task.
@@
 retained.\footnote{For further justification and to obtain the dataset,
-  see \blogpost} Stubs and disambiguation pages are also removed,
+  see\\ \blogpost} Stubs and disambiguation pages are also removed,
 leaving 463 thousand pages with a total of 482 million words.
 Punctuation marks and numbers were removed from the pages and all words
-were lower-cased.  Word occurrence counts are summarised 
+were lower-cased.  Word frequencies are summarised 
 in Table \ref{word-occurrences-table}.  This base corpus is then
@@
 hidden layer of size 100 is trained using negative sampling with 5
-negative samples, a window size of $10$, a minimum word occurrence of
+negative samples, a window size of $10$, a minimum frequency of
 $128$, and $10$ passes through the corpus.  Sub-sampling was not used so
@@
 that ensure that, for each word, multiple tokens will be inserted with a
-number of occurrences sufficient to survive the low-frequency cut-off of
+frequency sufficient to survive the low-frequency cut-off of
 128.  A maximum value $n=20$ suffices for this choice of $p$, since
-$2^{20 + \log_2{128}}$ exceeds the maximum occurrence count of any token in
+$2^{20 + \log_2{128}}$ exceeds the maximum frequency of any token in
 the corpus.  Figure \ref{fig:word-frequency-experiment-text-cat}
@@
 	\input{word-frequency-experiment-counts.tex}
-	\caption{Words chosen for the word frequency variation experiment, along with their occurrence counts in the unmodified corpus. }
+	\caption{Words chosen for the word frequency variation experiment, along with their frequency in the unmodified corpus. }
 	\label{word-frequency-counts}
@@
 this, we intersperse the word uniformly at random throughout the corpus so
-that its global occurrence rate is $0.005$.  The co-occurrence
-distribution of \word{VOID} thus coincides with the global word
+that its relative frequency is $0.005$.  The co-occurrence
+distribution of \word{VOID} thus coincides with the unconditional word
 distribution.  The replacement procedure is then performed for this
@@
 Figure \ref{fig:word-frequency-experiment-text-void} shows the effect of
-these modifications on a sample text, where a higher word occurrence rate of $0.05$ is used instead
+these modifications on a sample text, where a higher relative frequency of $0.05$ is used instead
 for illustrative purposes.  Although the probability of inserting token
@@
           illustrative purposes, the meaningless word \word{VOID} was
-          here interspersed with a occurrence rate of $0.05$.}
+          here interspersed with a relative frequency of $0.05$.}
 	\label{fig:word-frequency-experiment-text-void}
@@
 representing the tokens derived from any of these words can be properly
-learned from the modified corpus, even though their combined number of
-occurrences equals the term frequency of that word in the original
+learned from the modified corpus, even though their combined frequency equals the frequency of that word in the original
 corpus.  The red blocks in Figure \ref{word-frequency-experiment-heatmap}
@@
 $14$, before they start changing direction.  We conclude that these
-deviations are artifacts of poor learning due to an insufficient number
-of occurrences of these tokens.
+deviations are artifacts of poor learning due to an insufficient number of occurrences of these tokens.
 Table \ref{word-frequency-experiment-cosine-similarity}, showing the
@@
 frequency of the word \word{protestant} can be halved at most $6$ times
-so that the number of occurrences of the last token is still above the
+so that the frequency of the last token is still above the
 low-frequency cut-off.  Because all the points on a line share the same
@@
           vectors obtained from the second layer, syn1neg, are
-          included for completeness.  }
+          included for completeness.
+	  Legend entries are ordered by the L2-norm at the lowest-frequency
+  (i.e. left-most) sample point for the corresponding syn0 plot, descending.}
 	\label{fig:word-frequency-experiment-graph}
@@
 first experiment we studied the effect of word frequency on word vectors
-for fixed co-occurrence, we next study the effect of co-occurrence for
-fixed frequency.
+for fixed co-occurrence, we next study the effect of co-occurrence noise when the frequency is fixed.
+As before, we do so in a controlled manner.
 
 \subsection{Generating noise}
-We do so again in a controlled manner by adding noise to the
-co-occurrence distribution.  We take the noise distribution to be the
-global word distribution, so that noise can be simply added to the
-co-occurrence distribution of a word by interspersing occurrences of
+We take the noise distribution to be the (observed) unconditional word distribution.
+Thus noise can be added to the co-occurrence distribution of a word by simply interspersing occurrences of
 that word uniformly at random throughout the corpus.  A word that is
@@
 experiment.  The figure clearly demonstrates that vectors representing
-tokens derived from the same word all share a common direction.  This
-brings us to the conclusion that co-occurrence noise has no discernible effect on
-the direction of word vectors.
+tokens derived from the same word all share a common direction.  Thus
+co-occurrence noise has no discernible effect on the direction of word vectors.
 
-%While the word2vec neural network
-%samples co-occurrence noise, the direction of the vector representing
-%the target word changes, but these changes cancel out, so that the
-%direction remains fixed on average.
+On the other hand, for any experiment word, the introduced tokens mark
+intermediate points on a path in the space of distributions over the
+vocabulary, beginning at the co-occurrence distribution of the word and ending
+at the unconditional word frequency distribution, i.e. the co-occurrence
+distribution of \word{VOID}.
+Thus, a priori, the word vectors for the introduced tokens are intermediate
+steps on a path from the word vector of the original word to that of
+\word{VOID} (recall from Subsection \ref{WFVE-results} that word vector of
+\word{VOID} is approximately independent of its frequency).
+Since the word vector for \word{VOID} is close to the origin (see Subsection \ref{WFVE-results}), it may therefore be concluded that this path is approximately linear.
 
-The left panel in Figure \ref{fig:co-occurrence-noise-graph}, on the other hand, reveals that vector length varies more or less
+The left panel in Figure \ref{fig:co-occurrence-noise-graph} reveals that vector length varies more or less
 linearly with the proportion $r$ of noise in the co-occurrence
-distribution.  This figure motivates our interpretation of vector
+distribution.  This figure motivates an interpretation of vector
 length, within a sufficiently narrow frequency band, as a measure of the
@@
 
-%When read from right to left, this figure can be reinterpreted and
-%related to the corresponding figure in
-%Figure \ref{fig:word-frequency-experiment-graph} as follows.
-If the average effect of co-occurrence noise were neglible, then the horizontal axis of Figure \ref{fig:co-occurrence-noise-graph} could be reinterpreted as
-denoting word frequency by reading the graph from right to left.
-Thus a similar picture to Figure \ref{fig:word-frequency-experiment-graph} would emerge, namely that the
+It was observed that the average effect of noise occurrences on word vector direction was negligible.
+If the average effect of noise occurrences on word vector \textit{length} were also negligible, then the phenomenon depicted in Figure \ref{fig:co-occurrence-noise-graph} could be accounted for by the word frequency variation effects described in Section \ref{WFVE}.
+For then the horizontal axis of Figure \ref{fig:co-occurrence-noise-graph} could be reinterpreted as
+denoting word frequency by reading the graph from right to left,
+and a similar picture to Figure \ref{fig:word-frequency-experiment-graph} would emerge, namely that the
 word vector length increases more or less linearly with word frequency.
@@
 frequency.
-
-In this experiment, the introduced tokens are intermediate points on a path in the space of distributions over the vocabulary, beginning at the co-occurrence distribution of the word and ending at the global word frequency distribution, i.e.
-the co-occurrence distribution of \word{VOID}.
-Thus it is clear a priori that the word vectors for the introduced tokens are intermediate steps on a path from the word vector of the original word to that of \word{VOID} (recall from Subsection \ref{WFVE-results} that word vector of \word{VOID} is approximately independent of its frequency).
-Thus, while we do indeed observe from Figure \ref{fig:co-occurrence-noise-heatmap} that the word vectors for the tokens introduced for a word all share a common direction, this is only true because the word vector for \word{VOID} is close to the origin.
+We conclude that co-occurrence noise does indeed effect word vector length.
+In particular, the results of the two experiments illustrate distinct phenomena.
 
@@
           obtained from the first synaptic layer, syn0.  The second
-          layer, syn1neg, is included for completeness.  }
+          layer, syn1neg, is included for completeness.  
+	  Legend entries are ordered by the left-most value of the corresponding syn0 plot, descending.}
 	\label{fig:co-occurrence-noise-graph}
@@
 These experiments can be carried out for any word embedding (or indeed language model), because they are achieved via modification of the training corpus.
-Indeed, the experiments require no knowledge or modification of the model implementation.
+Indeed, the experiments even do not require knowledge of the model implementation.
 It would naturally be of interest to perform these experiments for other word embeddings other than word2vec CBOW, such as skipgrams and GloVe, as well as for different hyperparameters settings.
@@
 
-Further afield, it would be interesting to perform analogs of these experiments in other contexts where models are trained from co-occurrence data.
-For instance, in the context of e-Commerce, product purchase frequency and product co-occurrence noise can be varied by modifying the product-product co-occurrence matrix.
-
 Questions pertaining to word2vec in particular arise naturally from the results of the experiments.
@@
 
-The co-occurrence distribution of \word{VOID} is the global frequency distribution, and in this sense pure background noise.
+The co-occurrence distribution of \word{VOID} is the unconditional frequency distribution, and in this sense pure background noise.
 Thus the word vector of \word{VOID} is a special point in the feature space.

### 2015-09-07 build_images.py

@@ -55,10 +55,14 @@ def wf_experiment_tokens(word):
 test_words = random.sample([word for word in word_freq_experiment_words if word != 'the'], 4)
 test_words += ['the', meaningless_token]
 idxs = []
+ticks = []
 for word in test_words:
-    idxs += wf_experiment_tokens(word)
+    tokens = wf_experiment_tokens(word) 
+    idxs += tokens
+    ticks += tokens[:2] + ['.  '] * len(tokens[2:-1]) + tokens[-1:]
+
 test_vecs = vectors_syn0.loc[idxs]
-cosine_similarity_heatmap(test_vecs, figsize=(12, 10))
+cosine_similarity_heatmap(test_vecs, ticks, figsize=(12, 10))
 plt.savefig('outputs/word-frequency-experiment-heatmap.eps')
 
 # remove word vectors that are not well trained
@@ -74,34 +78,44 @@ for word in word_freq_experiment_words + [meaningless_token]:
 
 
 words = word_freq_experiment_words + [meaningless_token]
+
+# reorder by the left-most value of the corresponding syn0 plot
+def get_leftmost_value(word):
+    for i in range(1, word_freq_experiment_power_max + 1):
+        idx = build_experiment_token(word, i)
+        if idx in vectors_syn0.index:
+            last_idx = idx
+    return stats.L2_norm_syn0.loc[last_idx]
+words = sorted(words, key=get_leftmost_value, reverse=True)
+markers = [['o', 's', 'D'][i % 3] for i in range(len(words))]
+
 fig = plt.figure(figsize=(16, 6))
-colorcycle = plt.cm.gist_rainbow(np.linspace(0, 1, len(words)))
+colorcycle = plt.cm.gist_rainbow(np.linspace(0, 1, 5))
 
 ax_syn0 = plt.subplot(131)
-ax_syn0.set_xlabel('# occurrences')
+ax_syn0.set_xlabel('frequency')
 ax_syn0.set_ylabel('L2 norm')
 ax_syn0.set_xscale('log')
 ax_syn0.set_color_cycle(colorcycle)
 ax_syn0.set_title('syn0', y=1.04)
 
 ax_syn1neg = plt.subplot(132, sharex=ax_syn0)
-ax_syn1neg.set_xlabel('# occurrences')
+ax_syn1neg.set_xlabel('frequency')
 ax_syn1neg.set_color_cycle(colorcycle)
[…]

### 2015-09-07 functions.py

@@ -155,7 +155,7 @@ def cosine_similarity(vecs):
     vecs_normed = vecs.as_matrix() / np.sqrt((vecs ** 2).sum(axis=1))[:,np.newaxis]
     return vecs_normed.dot(vecs_normed.transpose())    
 
-def cosine_similarity_heatmap(test_vecs, **kwargs):
+def cosine_similarity_heatmap(test_vecs, ticks, **kwargs):
     mat = cosine_similarity(test_vecs)
     plt.figure(**kwargs)
     plt.title('Cosine similarity of word vectors')
@@ -163,8 +163,8 @@ def cosine_similarity_heatmap(test_vecs, **kwargs):
     _ = plt.gca().set_xlim(0, len(test_vecs.index))
     plt.pcolor(mat, vmin=-1, vmax=1)
     plt.colorbar()
-    _ = plt.yticks(np.arange(0.5, len(test_vecs.index), 1), test_vecs.index)
-    _ = plt.xticks(np.arange(0.5, len(test_vecs.index), 1), test_vecs.index, rotation=90)
+    _ = plt.yticks(np.arange(0.5, len(test_vecs.index), 1), ticks, fontsize=11)
+    _ = plt.xticks(np.arange(0.5, len(test_vecs.index), 1), ticks, rotation=90, fontsize=11)
     plt.tight_layout()
 
 def by_distance_from(table, v, **params):

### 2015-09-09 article/main.tex
final draft

@@
 	\And
-	Adriaan Schakel\\
+	Adriaan M.\,J.\, Schakel\\
 	NNLP\\
@@
 \begin{section}{Introduction}
-Word embeddings, or distributed representations of words, have been the subject
-of much recent research in the natural language processing and machine learning
-communities, demonstrating state-of-the-art performance on word similarity and
-word analogy tasks, amongst others.
-Word embeddings represent words from the vocabulary as dense, real-valued
+Word embeddings, or distributed representations of words, have been the
+subject of much recent research in the natural language processing and
+machine learning communities, demonstrating state-of-the-art performance
+on word similarity and word analogy tasks, amongst others.  Word
+embeddings represent words from the vocabulary as dense, real-valued
 vectors.  Instead of one-hot vectors that merely indicate the location
-of a word in the vocabulary, dense
-vectors of dimension much smaller than the vocabulary size are
-constructed such that they carry syntactic and semantic information.
-Irrespective of the technique chosen, word embeddings are
-typically derived from word co-occurrences.  More specifically, in a
-machine-learning setting, word embeddings are typically trained by
-scanning a short window over all the text in a corpus.  This process can
-be seen as sampling word co-occurrence distributions, where it is
-recalled that the co-occurrence distribution of a target word $w$
-denotes the conditional probability $\p(w'|w)$ that a word $w'$ occurs in
-its context, i.e. given that $w$ occurred.  Most applications of word
-embeddings explore not the word vectors themselves, but relations
-between them to solve, for example, similarity and word relation tasks
-\cite{vecchi-baroni-zamparelli2011}.  For these tasks, it was found that using normalised word
-vectors improves performance.  Word vector length is therefore typically
-ignored.
+of a word in the vocabulary, dense vectors of dimension much smaller
+than the vocabulary size are constructed such that they carry syntactic
+and semantic information.  Irrespective of the technique chosen, word
+embeddings are typically derived from word co-occurrences.  More
+specifically, in a machine-learning setting, word embeddings are
+typically trained by scanning a short window over all the text in a
+corpus.  This process can be seen as sampling word co-occurrence
+distributions, where it is recalled that the co-occurrence distribution
+of a target word $w$ denotes the conditional probability $\p(w'|w)$ that
+a word $w'$ occurs in its context, i.e., given that $w$ occurred.  Most
+applications of word embeddings explore not the word vectors themselves,
+but relations between them to solve, for example, similarity and word
+relation tasks \cite{dontcountpredict}.  For these tasks, it
+was found that using normalised word vectors improves performance.  Word
+vector length is therefore typically ignored.
 
@@
 
-This paper is structured as follows.  Section \ref{related-work} draws
-connections to related work, while Section \ref{corpus-and-model} describes
+This paper is structured as follows.  Section~\ref{related-work} draws
+connections to related work, while Section~\ref{corpus-and-model} describes
 the corpus and the CBOW model used in our experiments.
-Section \ref{WFVE} describes a controlled experiment for varying word
+Section~\ref{WFVE} describes a controlled experiment for varying word
 frequency while holding the co-occurrence distribution fixed.
-Section \ref{CNVE}, in a complementary fashion, describes a controlled
+Section~\ref{CNVE}, in a complementary fashion, describes a controlled
 experiment for varying the level of noise in the co-occurrence
 distribution of a word while holding the word frequency fixed.  The
-final section, Section \ref{future-directions}, considers further questions
+final section, Section~\ref{future-directions}, considers further questions
 and possible future directions.
@@
 ``contextual distinctiveness'' introduced earlier by McDonald and
-Shillcock \cite{mcdonald2001contextual}.
-They define the contextual distinctiveness of a target word $w$ as the
-Kullback-Leibler divergence
+Shillcock \cite{mcdonald2001contextual}.  They define the contextual
+distinctiveness of a target word $w$ as the Kullback-Leibler divergence
 between the unconditional word distribution $\p(w')$, and the
-co-occurrence distribution $\p(w' | w)$ of the target word $w$.
-So defined, contextual distinctiveness is a measure of the amount of
-information provided by word $w$ about its contexts of use.
-More specifically, if a word $w$ tends to appear in a wide variety of contexts, then its co-occurrence distribution tends to diverge less from unconditional word distribution, and its
-contextual distinctiveness is accordingly small.  If, on the other hand,
-$w$ typically appears in only a small number of different contexts, its co-occurrence distribution tends to diverge more from the unconditional word distribution, and a larger value of contextual distinctiveness results.
-In order to reduce the computational burden, the authors considered only $690$ target words, randomly chosen from equally divided frequency bands.
-Furthermore, the authors considered only co-occurrence with the $500$ most frequent content words.
-This significant restriction, necessary in their case in order to reduce the variance of the estimates of their measure of contextual distinctiveness, does not apply in our study.
+co-occurrence distribution $\p(w' | w)$ of the target word $w$.  So
+defined, contextual distinctiveness is a measure of the amount of
+information provided by word $w$ about its contexts of use.  More
+specifically, if a word $w$ tends to appear in a wide variety of
+contexts, then its co-occurrence distribution tends to diverge less from
+unconditional word distribution, and its contextual distinctiveness is
+accordingly small.  If, on the other hand, $w$ typically appears in only
+a small number of different contexts, its co-occurrence distribution
+tends to diverge more from the unconditional word distribution, and a
+larger value of contextual distinctiveness results.  To reduce the
+computational burden, the authors considered only $690$ target words,
+randomly chosen from equally divided frequency bands.  Furthermore, the
+authors considered only co-occurrence with the $500$ most frequent
+content words.  This significant restriction, necessary in their case in
+order to reduce the variance of the estimates of their measure of
+contextual distinctiveness, does not apply in our study.
 
@@
 vector length and the ``semantic deviance'' of an adjective-noun
-composite was studied empirically.  
-In that paper, which is also based
-on word co-occurrence, the authors study adjective-noun composites.
-They built a vocabulary from the 8k most frequent nouns and 4k most frequent adjectives in a large general language corpus and added 22k adjective-noun composites.
-For each item in the vocabulary, they recorded the co-occurrences with the top 10k most frequent content words (nouns, adjectives or verbs), and contructed word embeddings via singular value decomposition of the co-occurrence matrix \cite{landauer-dumais1997}.
-The authors considered several models for constructing vectors of unattested adjective-noun composites, the two simplest being adding and component-wise multiplying the adjective and noun vectors.
-They hypothesized that the length of the vectors thus constructed can be used to distinguish acceptable and semantically deviant adjective-noun composites.
-Using a few hundred adjective-noun composites selected by humans for evaluation, they found that deviant composites have a shorter vector than acceptable ones, in accordance with their expectation.
-In contrast to their work, our approach does not require extensive human evaluation.
+composite was studied empirically.  In that paper, which is also based
+on word co-occurrence statistics, the authors study adjective-noun
+composites.  They built a vocabulary from the 8k most frequent nouns and
+4k most frequent adjectives in a large general language corpus and added
+22k adjective-noun composites.  For each item in the vocabulary, they
+recorded the co-occurrences with the top 10k most frequent content words
+(nouns, adjectives or verbs), and constructed word embeddings via
+singular value decomposition of the co-occurrence matrix
+\cite{landauer-dumais1997}.  The authors considered several models for
+constructing vectors of unattested adjective-noun composites, the two
+simplest being adding and component-wise multiplying the adjective and
+noun vectors.  They hypothesized that the length of the vectors thus
+constructed can be used to distinguish acceptable and semantically
+deviant adjective-noun composites.  Using a few hundred adjective-noun
+composites selected by humans for evaluation, they found that deviant
+composites have a shorter vector than acceptable ones, in accordance
+with their expectation.  In contrast to their work, our approach does
+not require extensive human evaluation.
 
-Recent theoretical work \cite{Arora2015} has approached the problem of explaining the so-called ``compositionality'' property exhibited by some word embeddings.
-In that work, unnormalised vectors are used in their model of the word relationship task.
-It is hoped that experimental approaches such as those described here might enable theoretical investigations to describe the role of the word vector length in the word relationship tasks.
+Recent theoretical work \cite{Arora2015} has approached the problem of
+explaining the so-called ``compositionality'' property exhibited by some
+word embeddings.  In that work, unnormalised vectors are used in their
+model of the word relation task.  It is hoped that experimental
+approaches such as those described here might enable theoretical
+investigations to describe the role of the word vector length in the
+word relation tasks.
 \end{section}
@@
 Punctuation marks and numbers were removed from the pages and all words
-were lower-cased.  Word frequencies are summarised 
-in Table \ref{word-occurrences-table}.  This base corpus is then
-modified as described in Sections \ref{WFVE} and \ref{CNVE}.  For
+were lower-cased.  Word frequencies are summarised in
+Table~\ref{word-occurrences-table}.  This base corpus is then modified
+as described in Sections~\ref{WFVE} and \ref{CNVE}.  For
 recognisability, the tokens inserted into the corpus are upper-cased.
@@
 \begin{subsection}{Word2vec}\label{word2vec}
-Word2vec, a feed-forward neural network with a single hidden
-layer, learns word vectors from word co-occurrences in an unsupervised
-manner.  Word2vec comes in two versions.  In the continuous bag-of-words
-(CBOW) model, the words appearing around a target word serve as input.
-That input is projected linearly onto the hidden layer and the networks
-then attempts to predict the target word on output.  Training is
-achieved through back-propagation.  The word vectors are encoded in the
-weights of the first synaptic layer, ``syn0''.  The weights of the
-second synaptic layer (``syn1neg'', in the case of negative sampling) are typically discarded.  In the other model,
-called skip-gram, target and context words swap places, so that the
-target word now serves as input, while the network attempts to predict
-the context words on output.
+Word2vec, a feed-forward neural network with a single hidden layer,
+learns word vectors from word co-occurrences in an unsupervised manner.
+Word2vec comes in two versions.  In the continuous bag-of-words (CBOW)
+model, the words appearing around a target word serve as input.  That
+input is projected linearly onto the hidden layer and the network then
+attempts to predict the target word on output.  Training is achieved
+through back-propagation.  The word vectors are encoded in the weights
+of the first synaptic layer, ``syn0''.  The weights of the second
+synaptic layer (``syn1neg'', in the case of negative sampling) are
+typically discarded.  In the other model, called skip-gram, target and
+context words swap places, so that the target word now serves as input,
+while the network attempts to predict the context words on output.
 
@@
 
-The geometric distribution is truncated in order to limit the number of
-tokens inserted into the corpus.  For any choice $0 < p < 1$ and
-maximum value $n > 0$, the truncated geometric distribution is given by
-the probability density function
+The geometric distribution is truncated to limit the number of tokens
+inserted into the corpus.  For any choice $0 < p < 1$ and maximum value
+$n > 0$, the truncated geometric distribution is given by the
+probability density function
 \begin{equation} 
@@
 10 thousand times are chosen.  We also include the high-frequency
-stopword \word{the} for comparison.  Table \ref{word-frequency-counts}
+stopword \word{the} for comparison.  Table~\ref{word-frequency-counts}
 lists the words chosen for this experiment along with their frequencies.
 
-The replacement procedure of Section \ref{replacement-procedure} is then
+The replacement procedure of Section~\ref{replacement-procedure} is then
 performed for each of these words, using a geometric decay rate of $p =
@@
 $2^{20 + \log_2{128}}$ exceeds the maximum frequency of any token in
-the corpus.  Figure \ref{fig:word-frequency-experiment-text-cat}
+the corpus.  Figure~\ref{fig:word-frequency-experiment-text-cat}
 illustrates the effect of these modifications on a sample text, with a
@@
 	\input{word-frequency-experiment-counts.tex}
-	\caption{Words chosen for the word frequency variation experiment, along with their frequency in the unmodified corpus. }
+	\caption{Words chosen for the word frequency experiment, along with their frequency in the unmodified corpus. }
 	\label{word-frequency-counts}
@@
 	\caption{Example sentences modified in the word frequency
-          experiment as per Section \ref{WFVEexisting}, where the word
+          experiment as per Section~\ref{WFVEexisting}, where the word
           \word{cat} is replaced with tokens \word{CAT\_i} using the
@@
 word, using the same values for $p$ and $n$ as above.
-Figure \ref{fig:word-frequency-experiment-text-void} shows the effect of
+Figure~\ref{fig:word-frequency-experiment-text-void} shows the effect of
 these modifications on a sample text, where a higher relative frequency of $0.05$ is used instead
@@
 	\caption{The same example sentences as in
-          Figure \ref{fig:word-frequency-experiment-text-cat} where
+          Figure~\ref{fig:word-frequency-experiment-text-cat} where
           instead of the word \word{cat} now the meaningless word
@@
 \subsubsection{Word frequency and vector direction}\label{WFVE-direction}
-Figure \ref{word-frequency-experiment-heatmap} shows the cosine
+Figure~\ref{word-frequency-experiment-heatmap} shows the cosine
 similarity of pairs of vectors representing some of the tokens used in
@@
 associated with an experiment word is the number of times that its
-frequency can be halved and remain above the low-frequency cut-off of
+frequency can be halved and remains above the low-frequency cut-off of
 $128$.
@@
 Consider first the words other than \word{the} and \word{VOID}.  These
-words typically occur in a limited number of distinct contexts, and
-so a limited number of samples is required to capture their
-co-occurrence distributions.  We therefore expect that the vectors
-representing the tokens derived from any of these words can be properly
-learned from the modified corpus, even though their combined frequency equals the frequency of that word in the original
-corpus.  The red blocks in Figure \ref{word-frequency-experiment-heatmap}
-clearly demonstrate that the direction of these vectors is constant for
-each family of tokens.  In other words, the direction of these vectors
-is only determined by the original word and is independent of how often
-that words appear, provided sufficient samples have been processed for
-the word vector to settle in its equilibrium direction.
+words typically occur in a limited number of distinct contexts, and so a
+limited number of samples is required to capture their co-occurrence
+distributions.  We therefore expect that the vectors representing the
+tokens derived from any of these words can be properly learned from the
+modified corpus, even though their combined frequency equals the
+frequency of that word in the original corpus.  The red blocks in
+Figure~\ref{word-frequency-experiment-heatmap} clearly demonstrate that
+the direction of these vectors is constant for each family of tokens.
+In other words, the direction of these vectors is only determined by the
+original word and is independent of how often that word appears,
+provided sufficient samples have been processed for the word vector to
+settle in its equilibrium direction.
 
@@
 for the tokens occur increasingly less often.  This is consistent with
-Figure \ref{word-frequency-experiment-heatmap}, where it is apparent that
-the vectors representing \word{VOID\_i} for small $i$, say $i = 1,
+Figure~\ref{word-frequency-experiment-heatmap}, where it is apparent
+that the vectors representing \word{VOID\_i} for small $i$, say $i = 1,
 \dots, 5$, roughly share a common direction, while this is no longer
@@
 $14$, before they start changing direction.  We conclude that these
-deviations are artifacts of poor learning due to an insufficient number of occurrences of these tokens.
-Table \ref{word-frequency-experiment-cosine-similarity}, showing the
-most similar words to each token \word{THE\_i}, supports this further.  In the
-discussions below, the tokens \word{VOID\_i} and
-\word{THE\_i} whose vectors have a cosine similarity less than $0.8$
-with \word{VOID\_1} and \word{THE\_1}, respectively are therefore
-discarded.
+deviations are artifacts of poor learning due to an insufficient number
+of occurrences of these tokens.
+Table~\ref{word-frequency-experiment-cosine-similarity}, showing the
+most similar words to each token \word{THE\_i}, supports this further.
+In the discussions below, the tokens \word{VOID\_i} and \word{THE\_i}
+whose vectors have a cosine similarity less than $0.8$ with
+\word{VOID\_1} and \word{THE\_1}, respectively, are therefore discarded.
 
@@
 We next consider the effect of frequency on word vector length.
-Figure \ref{fig:word-frequency-experiment-graph} shows this relation for
+Throughout, we measure vector length using the Euclidean norm.
+Figure~\ref{fig:word-frequency-experiment-graph} shows this relation for
 individual words, both for the word vectors, represented by the weights
@@
 co-occurrence distribution, the left panel in
-Figure \ref{fig:word-frequency-experiment-graph} demonstrates conclusively
+Figure~\ref{fig:word-frequency-experiment-graph} demonstrates conclusively
 that length does indeed depend on frequency directly.  Moreover, this
@@
 with the experiment words are roughly independent of the frequency band,
-i.e. the plotted lines rarely cross.
+i.e., the plotted lines rarely cross.
 
 Notice the inflection of the syn1neg curve in
-Figure \ref{fig:word-frequency-experiment-graph} for \word{the} at
+Figure~\ref{fig:word-frequency-experiment-graph} for \word{the} at
 approximately $10^6$.  We have no explanation of this.  Also observe
 that the lengths of the vectors representing the meaningless tokens
-\word{VOID\_i} are approximately constant.  Since we already found the
+\word{VOID\_i} are approximately constant (about $2.5$).  Since we already found the
 direction to be also constant, it is sensible to speak of the word
@@
 	\centering{\includegraphics[scale=0.6]{word-frequency-experiment-graph}}
-	\caption{ Vector length vs. frequency for tokens derived from a
+	\caption{ Vector length vs.\ frequency for tokens derived from a
           few words chosen at random.  For each word, tokens of varying
@@
           were inserted into the corpus, as described in
-          Section \ref{WFVE}.  The vectors are obtained from the first
+          Section~\ref{WFVE}.  The vectors are obtained from the first
           synaptic layer, syn0, of the word2vec neural network.  The
-          vectors obtained from the second layer, syn1neg, are
-          included for completeness.
-	  Legend entries are ordered by the L2-norm at the lowest-frequency
-  (i.e. left-most) sample point for the corresponding syn0 plot, descending.}
+          vectors obtained from the second layer, syn1neg, are included
+          for completeness.  Legend entries are ordered by vector length
+          of the left-most data point in the syn0 plot, descending.}
 	\label{fig:word-frequency-experiment-graph}
@@
 first experiment we studied the effect of word frequency on word vectors
-for fixed co-occurrence, we next study the effect of co-occurrence noise when the frequency is fixed.
-As before, we do so in a controlled manner.
+for fixed co-occurrence, we here study the effect of co-occurrence noise
+when the frequency is fixed.  As before, we do so in a controlled
+manner.
 
 \subsection{Generating noise}
-We take the noise distribution to be the (observed) unconditional word distribution.
-Thus noise can be added to the co-occurrence distribution of a word by simply interspersing occurrences of
-that word uniformly at random throughout the corpus.  A word that is
-consistently used in a distinctive context in the unmodified corpus thus
-appears in the modified corpus also in completely unrelated contexts.
-As in Section \ref{WFVE}, we choose a small number of words from the
-unmodified corpus for this experiment.  Table
-\ref{cooccurrence-noise-words} lists the words chosen, along with their
-frequencies in the corpus.
+We take the noise distribution to be the (observed) unconditional word
+distribution.  Thus noise can be added to the co-occurrence distribution
+of a word by simply interspersing occurrences of that word uniformly at
+random throughout the corpus.  A word that is consistently used in a
+distinctive context in the unmodified corpus thus appears in the
+modified corpus also in completely unrelated contexts.  As in
+Section~\ref{WFVE}, we choose a small number of words from the
+unmodified corpus for this experiment.
+Table~\ref{cooccurrence-noise-words} lists the words chosen, along with
+their frequencies in the corpus.
 
-For each of these words, the replacement procedure of Subsection
-\ref{replacement-procedure} is performed using a geometric rate of decay
-of $p = \tfrac{5}{6}$, and truncating the distribution at $n=8$.  For
-every replacement token (e.g. \word{CAT\_i}), occurrences of this
-token are subsequently interspersed uniformly at random throughout the
-corpus, such that the frequency of the replacement token is restored to
-that of the original word \word{cat}.  For example, if the original word
-\word{cat} occurred $1000$ times, then after the replacement procedure,
-\word{CAT\_2} occurs approximately $694$ times, so a further
+For each of these words, the replacement procedure of
+Section~\ref{replacement-procedure} is performed using a geometric rate
+of decay of $p = \tfrac{5}{6}$, and truncating the distribution at
+$n=8$.  For every replacement token (e.g. \word{CAT\_i}), occurrences of
+this token are subsequently interspersed uniformly at random throughout
+the corpus, such that the frequency of the replacement token is restored
+to that of the original word \word{cat}.  For example, if the original
+word \word{cat} occurred $1000$ times, then after the replacement
+procedure, \word{CAT\_2} occurs approximately $694$ times, so a further
 (approximately) $306$ random occurrences of \word{CAT\_2} are
@@
 
-Figure \ref{fig:co-occurrence-noise-experiment-text} illustrates the
+Figure~\ref{fig:co-occurrence-noise-experiment-text} illustrates the
 effect of this modification in the case where the only word chosen is
@@
 is $r_i := 1 - P_{5/6, 8}(i)$.  The particular value of $p$ (large
-compared to that of Section \ref{WFVE}) was chosen so that the
+compared to that of Section~\ref{WFVE}) was chosen so that the
 proportion of noise $r_i$ did not increase too quickly with $i$.  These
@@
 \begin{subsection}{Experimental results}
-Figure \ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity
+Figure~\ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity
 of pairs of vectors representing some of the tokens used in this
@@
 tokens derived from the same word all share a common direction.  Thus
-co-occurrence noise has no discernible effect on the direction of word vectors.
+co-occurrence noise has no discernible effect on the direction of word
+vectors.
 
-On the other hand, for any experiment word, the introduced tokens mark
-intermediate points on a path in the space of distributions over the
-vocabulary, beginning at the co-occurrence distribution of the word and ending
-at the unconditional word frequency distribution, i.e. the co-occurrence
-distribution of \word{VOID}.
-Thus, a priori, the word vectors for the introduced tokens are intermediate
-steps on a path from the word vector of the original word to that of
-\word{VOID} (recall from Subsection \ref{WFVE-results} that word vector of
-\word{VOID} is approximately independent of its frequency).
-Since the word vector for \word{VOID} is close to the origin (see Subsection \ref{WFVE-results}), it may therefore be concluded that this path is approximately linear.
+The left panel in Figure~\ref{fig:co-occurrence-noise-graph} reveals
+that word vector length varies more or less linearly with the proportion
+of noise in the co-occurrence distribution of the word.  This figure motivates an
+interpretation of vector length, within a sufficiently narrow frequency
+band, as a measure of the absence of co-occurrence noise, or put
+differently, of the extent to which a word carries the meaning of a
+distinctive context.
+The right panel of Figure~\ref{fig:co-occurrence-noise-graph} shows that the
+syn1neg vector length does not vary with the proportion of co-occurrence noise.
 
-The left panel in Figure \ref{fig:co-occurrence-noise-graph} reveals that vector length varies more or less
-linearly with the proportion $r$ of noise in the co-occurrence
-distribution.  This figure motivates an interpretation of vector
-length, within a sufficiently narrow frequency band, as a measure of the
-absence of co-occurrence noise, or put differently, of the extent to
-which a word carries the meaning of a distinctive context.
-
-It was observed that the average effect of noise occurrences on word vector direction was negligible.
-If the average effect of noise occurrences on word vector \textit{length} were also negligible, then the phenomenon depicted in Figure \ref{fig:co-occurrence-noise-graph} could be accounted for by the word frequency variation effects described in Section \ref{WFVE}.
-For then the horizontal axis of Figure \ref{fig:co-occurrence-noise-graph} could be reinterpreted as
-denoting word frequency by reading the graph from right to left,
-and a similar picture to Figure \ref{fig:word-frequency-experiment-graph} would emerge, namely that the
-word vector length increases more or less linearly with word frequency.
-This can not, however, be the complete picture, for the vectors
-represented by the second synaptic layer, syn1neg, show a different
-behavior in the two figures.  In particular, the length of these vectors
-is seen from the right panel of Figure \ref{fig:co-occurrence-noise-graph}
-to be completely insensitive to noise, i.e. frequency in our
-reinterpretation, whereas the length of those shown in the right panel
-of Figure \ref{fig:word-frequency-experiment-graph} does depend on
-frequency.
-We conclude that co-occurrence noise does indeed effect word vector length.
-In particular, the results of the two experiments illustrate distinct phenomena.
+Recall from Section~\ref{WFVE-results} that the word vector for \word{VOID} is independent of the frequency of \word{VOID}, and is non-zero.
+Since the noise distribution coincides with the co-occurrence distribution of \word{VOID}, the word vectors for the experiment words must tend to the word vector for \word{VOID} as the proportion of noise in their co-occurrence distributions approaches $1$.
+This is not discernible in Figure~\ref{fig:co-occurrence-noise-heatmap} since the word vector for \word{VOID} is relatively close to the origin.
 
@@
 	\includegraphics[scale=0.6]{cooccurrence-noise-graph}
-	\caption{ Vector length vs. proportion of co-occurrence noise
+	\caption{ Vector length vs.\ proportion of occurrences from the noise distribution
           for words chosen for this experiment.  For each word, tokens
@@
           co-occurrence noise were inserted into the corpus, as
-          described in section \ref{CNVE}.  The word vectors are
+          described in Section~\ref{CNVE}.  The word vectors are
           obtained from the first synaptic layer, syn0.  The second
-          layer, syn1neg, is included for completeness.  
-	  Legend entries are ordered by the left-most value of the corresponding syn0 plot, descending.}
+          layer, syn1neg, is included for completeness.  Legend entries
+          are ordered by vector length of the left-most data point in
+          the syn0 plot, descending.}
 	\label{fig:co-occurrence-noise-graph}
@@
 \begin{section}{Discussion}\label{future-directions}
-Our principle contribution has been to demonstrate that controlled experiments can be used to gain insight into a word embedding.
-These experiments can be carried out for any word embedding (or indeed language model), because they are achieved via modification of the training corpus.
-Indeed, the experiments even do not require knowledge of the model implementation.
-It would naturally be of interest to perform these experiments for other word embeddings other than word2vec CBOW, such as skipgrams and GloVe, as well as for different hyperparameters settings.
+Our principle contribution has been to demonstrate that controlled
+experiments can be used to gain insight into a word embedding.  These
+experiments can be carried out for any word embedding (or indeed
+language model), for they are achieved via modification of the training
+corpus only.  They do not require knowledge of the model implementation.
+It would naturally be of interest to perform these experiments for other
+word embeddings other than word2vec CBOW, such as skipgrams and GloVe,
+as well as for different hyperparameters settings.
 
-More elaborate experiments could be carried out.
-For instance, by introducing tokens into the corpus that mix, with varying proportions, the co-occurrence distributions of two words, the path between the word vectors in the feature space could be studied.
-The co-occurrence noise variation experiment described here would be a special case of such an experiment where one of the two words was \word{VOID}.
+More elaborate experiments could be carried out.  For instance, by
+introducing tokens into the corpus that mix, with varying proportions,
+the co-occurrence distributions of two words, the path between the word
+vectors in the feature space could be studied.  The co-occurrence noise
+experiment described here would be a special case of such an
+experiment where one of the two words was \word{VOID}.
 
-Questions pertaining to word2vec in particular arise naturally from the results of the experiments.
-Figures \ref{fig:word-frequency-experiment-graph} and \ref{fig:co-occurrence-noise-graph}, for example, demonstrate the word vectors obtained from the first synaptic layer, syn0, have very different properties from those that could be obtained from the second layer, syn1neg, which are typically discarded.
-These differences warrant further investigation.
+Questions pertaining to word2vec in particular arise naturally from the
+results of the experiments.
+Figures~\ref{fig:word-frequency-experiment-graph} and
+\ref{fig:co-occurrence-noise-graph}, for example, demonstrate that the
+word vectors obtained from the first synaptic layer, syn0, have very
+different properties from those that could be obtained from the second
+layer, syn1neg.  These differences warrant further investigation.
 
-The co-occurrence distribution of \word{VOID} is the unconditional frequency distribution, and in this sense pure background noise.
-Thus the word vector of \word{VOID} is a special point in the feature space.
-Figure \ref{fig:word-frequency-experiment-graph} shows that this point is not at the origin of the feature space (i.e. is not the zero vector).
-The origin, however, is implicitly the point of reference in the word2vec word similarity task.
-This raises the question of whether improved performance on the similarity task could be achieved by transforming the feature space (or modifying the model) such that the representation of pure noise (i.e.
-the vector for \word{VOID}) is at the origin of the transformed feature space.
+The co-occurrence distribution of \word{VOID} is the unconditional
+frequency distribution, and in this sense pure background noise.  Thus
+the word vector of \word{VOID} is a special point in the feature space.
+Figure~\ref{fig:word-frequency-experiment-graph} shows that this point
+is not at the origin of the feature space, i.e., is not the zero vector.
+The origin, however, is implicitly the point of reference in word2vec
+word similarity tasks.  This raises the question of whether improved
+performance on similarity tasks could be achieved by transforming the
+feature space or modifying the model such that the representation of
+pure noise, i.e., the vector for \word{VOID}, is at the origin of the
+transformed feature space.
 
 %\textcolor{red}{can it be argued that the feature space must be fundamentally Euclidean, given that the path marked by the vectorisations of intermediate words is a straight line?
-i.e.
-if such paths maybe considered as ``geodesics'' between words, and we show that these geodesics appear as straight lines when an Euclidean metric is imposed, then suggests that the Euclidean metric is in fact a good choice.
+i.e., if such paths maybe considered as ``geodesics'' between words, and we show that these geodesics appear as straight lines when an Euclidean metric is imposed, then suggests that the Euclidean metric is in fact a good choice.
 what would happen if we chose another metric?}

### 2015-09-09 article_markup_wordcounts.py

@@ -2,7 +2,7 @@ import sys
 
 print r"""\begin{center}
 \begin{tabular}{l | r}
-word & \# occurrences \\
+word & frequency \\
 \hline"""
 wcs = []
 for line in sys.stdin:

### 2015-09-09 build_images.py

@@ -37,8 +37,8 @@ total_words = sum(new_counts.values())
 stats = pd.DataFrame({'occurrences': new_counts, 'L2_norm_syn0': norms_syn0, 'L2_norm_syn1neg': norms_syn1neg}).dropna()
 stats['occurrences_band'] = np.floor(np.log2(stats.occurrences)).astype(int)
 stats['log2_frequency'] = np.log2(stats.occurrences * 1. / total_words)
-stats.L2_norm_syn0.name = 'L2 norm (syn0)'
-stats.L2_norm_syn1neg.name = 'L2 norm (syn1neg)'
+stats.L2_norm_syn0.name = 'vector length (syn0)'
+stats.L2_norm_syn1neg.name = 'vector length (syn1neg)'
 
 
 # WORD FREQUENCY EXPERIMENT
@@ -94,7 +94,7 @@ colorcycle = plt.cm.gist_rainbow(np.linspace(0, 1, 5))
 
 ax_syn0 = plt.subplot(131)
 ax_syn0.set_xlabel('frequency')
-ax_syn0.set_ylabel('L2 norm')
+ax_syn0.set_ylabel('vector length')
 ax_syn0.set_xscale('log')
 ax_syn0.set_color_cycle(colorcycle)
 ax_syn0.set_title('syn0', y=1.04)
@@ -138,13 +138,13 @@ fig = plt.figure(figsize=(16, 6))
 colorcycle = plt.cm.gist_rainbow(np.linspace(0, 1, 5))
 
 ax_syn0 = plt.subplot(131)
-ax_syn0.set_xlabel('Proportion of noise occurrences')
-ax_syn0.set_ylabel('L2 norm')
+ax_syn0.set_xlabel('Proportion of occurrences from noise distribution')
+ax_syn0.set_ylabel('vector length')
 ax_syn0.set_color_cycle(colorcycle)
 ax_syn0.set_title('syn0', y=1.04)
 
 ax_syn1neg = plt.subplot(132, sharex=ax_syn0, sharey=ax_syn0)
-ax_syn1neg.set_xlabel('Proportion of noise occurrences')
+ax_syn1neg.set_xlabel('Proportion of occurrences from noise distribution')
 ax_syn1neg.set_color_cycle(colorcycle)
 ax_syn1neg.set_title('syn1neg', y=1.04)
     

### 2015-09-10 article/cooccurrence-noise-experiment.tex
first attempt at redefining the coocc noise experiment

--- 2015-08-18 article/cooccurrence-noise-experiment.tex
@@
-\texttt{\textcolor{red}{CAT\_3} the domestic \textcolor{red}{CAT\_3} was first classified as felis \textcolor{red}{CAT\_2} catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_3} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_2} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_2} is one \textcolor{red}{CAT\_2} whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_2} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated \textcolor{red}{CAT\_2} canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time \textcolor{red}{CAT\_3} at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}
+\texttt{the domestic \textcolor{red}{CAT\_3} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_2} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_3} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_6} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_4} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_4} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

### 2015-09-10 article/word-frequency-experiment-text-cat.tex
first attempt at redefining the coocc noise experiment

@@
-\texttt{the domestic \textcolor{red}{CAT\_3} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_3} a mostly outdoor \textcolor{red}{CAT\_7} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_1} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_1} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_3} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_1} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_3} is usually dominant over the others}
+\texttt{the domestic \textcolor{red}{CAT\_1} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_1} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_1} is one whose ancestry is recorded by a \textcolor{red}{CAT\_2} fancier organization\newline a purebred \textcolor{red}{CAT\_1} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_3} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_3} is usually dominant over the others}

### 2015-09-10 article/word-frequency-experiment-text-void.tex
first attempt at redefining the coocc noise experiment

@@
-\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned \textcolor{red}{VOID\_3} by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in \textcolor{red}{VOID\_1} having very large eye sockets\newline \textcolor{red}{VOID\_2} another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others}
+\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline \textcolor{red}{VOID\_1} within groups one cat is usually dominant over the others}

### 2015-09-10 article_generate_sample_texts.py

@@ -21,7 +21,8 @@ def write_markup(raw_output, f_out):
 # PERFORM THE REPLACEMENT PROCEDURE FOR THE WORD FREQUENCY EXPERIMENT FOR THE WORD "CAT"
 
 word = 'cat'
-word_samplers = {word: truncated_geometric_sampling(word, word_freq_experiment_ratio, word_freq_experiment_power_max)}
+distn = lambda i: truncated_geometric_proba(word_freq_experiment_ratio, i, word_freq_experiment_power_max)
+word_samplers = {word: distribution_to_sampling_function(word, distn, word_freq_experiment_power_max)}
 
 tmp_file = StringIO()
 with file('article/word-frequency-experiment-text-cat.input') as f_in:
@@ -39,16 +40,17 @@ with file('article/word-frequency-experiment-text-void.input') as f_in:
 tmp_file1.seek(0)
 
 tmp_file2 = StringIO()
-word_samplers = {meaningless_token: truncated_geometric_sampling(meaningless_token, word_freq_experiment_ratio, word_freq_experiment_power_max)}
+distn = lambda i: truncated_geometric_proba(word_freq_experiment_ratio, i, word_freq_experiment_power_max)
+word_samplers = {meaningless_token: distribution_to_sampling_function(meaningless_token, distn, word_freq_experiment_power_max)}
 replace_words(word_samplers, tmp_file1, tmp_file2)
 
 with file('article/word-frequency-experiment-text-void.tex', 'w') as f_out:
     write_markup(tmp_file2.getvalue(), f_out)
 
 # COOCCURRENCE NOISE EXPERIMENT
-coocc_noise_experiment_power_max = 3 # better for the example, be sure to note in caption
 word = 'cat'
-word_samplers = {word: truncated_geometric_sampling(word, coocc_noise_experiment_ratio, coocc_noise_experiment_power_max)}
+distn = lambda i: evenly_spaced_proba(i, coocc_noise_experiment_max_value)
+word_samplers = {word: distribution_to_sampling_function(word, distn, coocc_noise_experiment_max_value)}
 
 with file('article/cooccurrence-noise-experiment.input') as f_in:
     counts = count_words(f_in)
@@ -59,11 +61,11 @@ with file('article/cooccurrence-noise-experiment.input') as f_in:
     replace_words(word_samplers, f_in, tmp_file1)
 tmp_file1.seek(0)
 
-# add noise to the cooccurrence distributions of experiment 2 words
[…]

### 2015-09-10 build_images.py

@@ -149,9 +149,9 @@ ax_syn1neg.set_color_cycle(colorcycle)
 ax_syn1neg.set_title('syn1neg', y=1.04)
     
 def plot_for_word(ax, word, series, **kwargs):
-    exponents = filter(lambda i: build_experiment_token(word, i) in stats.index, range(1, coocc_noise_experiment_power_max + 1))
-    idxs = [build_experiment_token(word, i) for i in exponents]
-    x = [1 - coocc_noise_experiment_ratio ** exponent for exponent in exponents]
+    outcomes = range(1, coocc_noise_experiment_max_value + 1)
+    idxs = [build_experiment_token(word, i) for i in outcomes]
+    x = [noise_proportion(i, coocc_noise_experiment_max_value) for i in outcomes]
     y = series.loc[idxs]
     marker = ['o', 's', 'D'][ord(word[0]) % 3]
     return ax.plot(x, y, marker=marker, **kwargs)[0]
@@ -175,7 +175,7 @@ plt.savefig('outputs/cooccurrence-noise-graph.eps')
 idxs = []
 ticks = []
 for word in random.sample(coocc_noise_experiment_words, 4):
-    candidates = [build_experiment_token(word, i) for i in range(1, coocc_noise_experiment_power_max + 1)]
+    candidates = [build_experiment_token(word, i) for i in range(1, coocc_noise_experiment_max_value + 1)]
     tokens = [token for token in candidates if token in vectors_syn0.index]
     idxs += tokens
     ticks += tokens[:2] + ['.  '] * len(tokens[2:-1]) + tokens[-1:]

### 2015-09-10 functions.py

@@ -65,14 +65,32 @@ def truncated_geometric_proba(ratio, i, n):
     """
     return (ratio ** (i - 1)) * (1 - ratio) / (1 - ratio ** n)
 
-def truncated_geometric_sampling(word, ratio, max_value):
+def distribution_to_sampling_function(word, dist_fn, max_value):
     """
-    returns a function that samples from the truncated geometric distribution, truncated
-    at max_value, returning WORD_1, .. WORD_<max_value>
+    returns a function that samples from the given distribution on [1 ..
+    max_value], returning WORD_1, .. WORD_<max_value>
     """
     outcomes = [build_experiment_token(word, value) for value in range(1, max_value + 1)]
-    probs = np.array([truncated_geometric_proba(ratio, i, max_value) for i in range(1, max_value + 1)])
+    probs = np.array([dist_fn(i) for i in range(1, max_value + 1)])
     return lambda: np.random.choice(outcomes, p=probs)
+
+def evenly_spaced_proba(i, M): 
+    """
+    A probability distribution on [1 .. M] with the property that the
+    probability densities are evenly spaced, i.e. p(i) - p(i+1) = c for all i.
+    The sequence is decreasing, so c > 0.
+    """
+    return 2. * (M - i + 1) / (M * (M+1))
+
+def noise_proportion(i, M):
+    """
+    Assuming that the evenly_spaced_proba(i,M) distribution was used and that
+    the total number of occurrences (original + noise) is given by
+        2 * #original / (M+1),
+    irrespective of i, return the expected proportion of
+    noise occurrences.
+    """
+    return 1 - 0.5 * (M + 1) * evenly_spaced_proba(i, M)
 
 def intersperse_words(interspersal_rates, f_in, f_out):
     """

### 2015-09-10 modify_corpus_coocc_noise_experiment.py

@@ -15,8 +15,9 @@ total_words = sum(counts.values())
 
 # perform the replacement procedure
 word_samplers = {}
+distn = lambda i: evenly_spaced_proba(i, coocc_noise_experiment_max_value)
 for word in cn_experiment_words:
-    word_samplers[word] = truncated_geometric_sampling(word, coocc_noise_experiment_ratio, coocc_noise_experiment_power_max)
+    word_samplers[word] = distribution_to_sampling_function(word, distn, coocc_noise_experiment_max_value)
 
 intermediate_file = 'delete.me.word_freq_experiment'
 with open(intermediate_file, 'w') as f_out:
@@ -25,9 +26,9 @@ with open(intermediate_file, 'w') as f_out:
 # add noise to the cooccurrence distributions
 token_freq_dict = dict()
 for word in cn_experiment_words:
-    target_freq = counts[word] * 1. / total_words
-    for i in range(1, coocc_noise_experiment_power_max + 1):
-        current_freq = target_freq * truncated_geometric_proba(coocc_noise_experiment_ratio, i, coocc_noise_experiment_power_max)
+    target_freq = counts[word] * 1. * coocc_noise_experiment_freq_reduction / total_words
+    for i in range(1, coocc_noise_experiment_max_value + 1):
+        current_freq = counts[word] * evenly_spaced_proba(i, coocc_noise_experiment_max_value)
         token_freq_dict[build_experiment_token(word, i)] = target_freq - current_freq
 
 with open(intermediate_file) as f_in:

### 2015-09-10 modify_corpus_word_freq_experiment.py

@@ -21,8 +21,9 @@ wf_experiment_words.append(meaningless_token)
 
 # perform the replacement procedure
 word_samplers = {}
+distn = lambda i: truncated_geometric_proba(word_freq_experiment_ratio, i, word_freq_experiment_power_max)
 for word in wf_experiment_words:
-    word_samplers[word] = truncated_geometric_sampling(word, word_freq_experiment_ratio, word_freq_experiment_power_max)
+    word_samplers[word] = distribution_to_sampling_function(word, distn, word_freq_experiment_power_max)
 
 with open(intermediate_file) as f_in:
     replace_words(word_samplers, f_in, sys.stdout)

### 2015-09-10 parameters.py

@@ -9,5 +9,5 @@ experiment_word_occurrence_min = 10000
 word_freq_experiment_ratio = 1./2
 word_freq_experiment_power_max = 20
 
-coocc_noise_experiment_ratio = 5./6
-coocc_noise_experiment_power_max = 8
+coocc_noise_experiment_max_value = 7
+coocc_noise_experiment_freq_reduction = 2. / (coocc_noise_experiment_max_value + 1)

### 2015-09-10 build_images.py
fix coocc noise experiment setup

@@ -163,7 +163,7 @@ for word in words:
 for word in words:
     plot_for_word(ax_syn1neg, word, stats.L2_norm_syn1neg)
 
-ax_syn0.set_ylim(2.5, 8.5)
+#ax_syn0.set_ylim(2.5, 8.5)
 ax_syn0.set_xlim(0, 1)
 
 _ = fig.legend(lines, words, bbox_to_anchor=(0.76, 0.56), loc='center', fontsize=14, frameon=False)

### 2015-09-10 functions.py

@@ -5,7 +5,7 @@ import pandas as pd
 from scipy.spatial.distance import pdist, cdist
 import matplotlib.pyplot as plt
 from collections import defaultdict
-from parameters import random_seed, experiment_word_occurrence_min
+from parameters import random_seed, experiment_word_occurrence_min, coocc_noise_experiment_freq_reduction
 
 def count_words(f):
     """
@@ -79,18 +79,19 @@ def evenly_spaced_proba(i, M):
     A probability distribution on [1 .. M] with the property that the
     probability densities are evenly spaced, i.e. p(i) - p(i+1) = c for all i.
     The sequence is decreasing, so c > 0.
+    p(M) = 0
     """
-    return 2. * (M - i + 1) / (M * (M+1))
+    return 2. * (M - i) / (M * (M-1))
 
 def noise_proportion(i, M):
     """
     Assuming that the evenly_spaced_proba(i,M) distribution was used and that
     the total number of occurrences (original + noise) is given by
-        2 * #original / (M+1),
+        #original * coocc_noise_experiment_freq_reduction
     irrespective of i, return the expected proportion of
     noise occurrences.
     """
-    return 1 - 0.5 * (M + 1) * evenly_spaced_proba(i, M)
+    return 1 - evenly_spaced_proba(i, M) / coocc_noise_experiment_freq_reduction
 
 def intersperse_words(interspersal_rates, f_in, f_out):
     """

### 2015-09-10 modify_corpus_coocc_noise_experiment.py

@@ -28,7 +28,7 @@ token_freq_dict = dict()
 for word in cn_experiment_words:
     target_freq = counts[word] * 1. * coocc_noise_experiment_freq_reduction / total_words
     for i in range(1, coocc_noise_experiment_max_value + 1):
-        current_freq = counts[word] * evenly_spaced_proba(i, coocc_noise_experiment_max_value)
+        current_freq = counts[word] * evenly_spaced_proba(i, coocc_noise_experiment_max_value) / total_words
         token_freq_dict[build_experiment_token(word, i)] = target_freq - current_freq
 
 with open(intermediate_file) as f_in:

### 2015-09-10 parameters.py
fix parameters to give full span of coocc noise proportions

@@ -10,4 +10,4 @@ word_freq_experiment_ratio = 1./2
 word_freq_experiment_power_max = 20
 
 coocc_noise_experiment_max_value = 7
-coocc_noise_experiment_freq_reduction = 2. / (coocc_noise_experiment_max_value + 1)
+coocc_noise_experiment_freq_reduction = 2. / coocc_noise_experiment_max_value

### 2015-09-10 build_images.py
images resulting from a run on the entire corpus

@@ -8,7 +8,7 @@ import sys
 from parameters import *
 from functions import *
 
-random.seed(0) # fix seed so that we can refer to the randomly chosen words in the article body
+random.seed(1) # fix seed so that we can refer to the randomly chosen words in the article body
 
 vectors_syn0_filename = sys.argv[1]
 vectors_syn1neg_filename = sys.argv[2]
@@ -143,7 +143,7 @@ ax_syn0.set_ylabel('vector length')
 ax_syn0.set_color_cycle(colorcycle)
 ax_syn0.set_title('syn0', y=1.04)
 
-ax_syn1neg = plt.subplot(132, sharex=ax_syn0, sharey=ax_syn0)
+ax_syn1neg = plt.subplot(132, sharex=ax_syn0)
 ax_syn1neg.set_xlabel('Proportion of occurrences from noise distribution')
 ax_syn1neg.set_color_cycle(colorcycle)
 ax_syn1neg.set_title('syn1neg', y=1.04)
@@ -163,7 +163,8 @@ for word in words:
 for word in words:
     plot_for_word(ax_syn1neg, word, stats.L2_norm_syn1neg)
 
-#ax_syn0.set_ylim(2.5, 8.5)
+ax_syn0.set_ylim(0, 30)
+ax_syn1neg.set_ylim(0, 15)
 ax_syn0.set_xlim(0, 1)
 
 _ = fig.legend(lines, words, bbox_to_anchor=(0.76, 0.56), loc='center', fontsize=14, frameon=False)

### 2015-10-01 article/cooccurrence-noise-experiment.tex
integration of the corrected noise experiment into article text

@@
-\texttt{the domestic \textcolor{red}{CAT\_3} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_2} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_3} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline a purebred \textcolor{red}{CAT\_6} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_4} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_4} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}
+\texttt{the domestic \textcolor{red}{CAT\_2} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_3} a mostly outdoor \textcolor{red}{CAT\_4} is not \textcolor{red}{CAT\_2} owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_4} is one whose ancestry is recorded by a \textcolor{red}{CAT\_1} fancier organization\newline \textcolor{red}{CAT\_6} a purebred \textcolor{red}{CAT\_3} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very \textcolor{red}{CAT\_4} large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_4} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_2} is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively \textcolor{red}{CAT\_5} bred\newline dogs perform many roles for people such as hunting herding and pulling loads\newline \textcolor{red}{CAT\_7} in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is \textcolor{red}{CAT\_6} the time at \textcolor{red}{CAT\_3} which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}

### 2015-10-01 article/main.tex
integration of the corrected noise experiment into article text

@@
 \usepackage{rotating}
+
 \usepackage[outdir=./]{epstopdf}
@@
         demonstration of direct relations between word properties and
-        word vector direction and length.  The new approach is
-        demonstrated using the word2vec CBOW model with experiments that
+        word vector direction and length.  The approach is demonstrated
+        using the word2vec CBOW model with experiments that
         independently vary word frequency and word co-occurrence noise.
@@
         the co-occurrence distribution of the word.  The coefficients of
-        linearity depend upon the word.
+        linearity depend upon the word.  The special vector in feature
+        space, defined by the (artificial) word with pure noise in its
+        co-occurrence distribution, is found to be small but non-zero.
 \end{abstract}
@@
 distribution of the word is increased.  We show furthermore that the
-direction of the word vectors is independent of both word frequency and
-co-occurrence noise.
+direction of the word vectors is independent of word frequency and only
+moderately depends upon co-occurrence noise.  When noise is added to the
+co-occurrence distribution of a word, the corresponding vector smoothly
+interpolates between the original word vector and a small vector
+perpendicular to it that represents a word with pure noise in its
+co-occurrence distribution.  Surprisingly, this special vector in the feature
+space, obtained by interspersing a token uniformly at random throughout
+the corpus with a frequency sufficiently large to sample all contexts, is
+non-zero.
 
@@
 \begin{section}{Related work}\label{related-work}
+\mar{move to other paper?}
 Our notion of word significance is reminiscent of the concept of
@@
 with their expectation.  In contrast to their work, our approach does
-not require extensive human evaluation.
+not require human annotation.
 
@@
 In the experiments detailed below, we modify the corpus in a controlled
-manner by introducing tokens into the corpus via a replacement procedure
-as follows.  Consider a word, say \word{cat}.  For each occurrence of
-this word, a sample $i$, $1 \leqslant i \leqslant n$ is drawn from a
-truncated geometric distribution, and that occurrence of the word
-\word{cat} is replaced with the token \word{CAT\_i}.  In this way, the
-word \word{cat} is replaced throughout the corpus by a family of tokens
-with varying frequencies but approximately the same co-occurrence
-distribution as \word{cat}.  That is, all these tokens are used in
-roughly the same contexts as the original word.
+manner by introducing tokens into the corpus via a replacement
+procedure.  For the frequency experiment, the procedure is as follows.
+Consider a word, say \word{cat}.  For each occurrence of this word, a
+sample $i$, $1 \leqslant i \leqslant n$ is drawn from a truncated
+geometric distribution, and that occurrence of the word \word{cat} is
+replaced with the token \word{CAT\_i}.  In this way, the word \word{cat}
+is replaced throughout the corpus by a family of tokens with varying
+frequencies but approximately the same co-occurrence distribution as
+\word{cat}.  That is, all these tokens are used in roughly the same
+contexts as the original word.
 
@@
 for the experiments.
+
+For the noise experiment, we take instead of a geometric distribution,
+the distribution
+\begin{equation}
+\label{distro2} 
+P_n (i) = \frac{2 (n-i)}{n(n-1)} , \qquad 1 \leqslant i \leqslant n.
+\end{equation} 
+We have chosen this distribution for the noise experiment, because it
+leads to evenly spaced proportions of co-occurrence noise that cover the entire interval
+$[0,1]$.
 \end{subsection}
@@
 
-\begin{subsection}{Tokens derived from an artificial, meaningless word}\label{WFVEmeaningless}
+\begin{subsection}{Tokens derived from an artificial, meaningless word}
+\label{WFVEmeaningless}
 Whereas the tokens introduced above all replace an existing word that
@@
 existing (stop)word whose meaninglessness is only supposed.  To achieve
-this, we intersperse the word uniformly at random throughout the corpus so
-that its relative frequency is $0.005$.  The co-occurrence
+this, we intersperse the word uniformly at random throughout the corpus
+so that its relative frequency is $0.005$.  The co-occurrence
 distribution of \word{VOID} thus coincides with the unconditional word
@@
 Figure~\ref{fig:word-frequency-experiment-text-void} shows the effect of
-these modifications on a sample text, where a higher relative frequency of $0.05$ is used instead
-for illustrative purposes.  Although the probability of inserting token
+these modifications on a sample text, where a higher relative frequency
+of $0.05$ is used instead for illustrative purposes.  Although the probability of inserting token
 \word{VOID\_1} is twice that of inserting token \word{VOID\_2} and four
@@
           illustrative purposes, the meaningless word \word{VOID} was
-          here interspersed with a relative frequency of $0.05$.}
+          here interspersed with a relative frequency of
+          $0.05$. }
 	\label{fig:word-frequency-experiment-text-void}
@@
 frequency of the word \word{protestant} can be halved at most $6$ times
-so that the frequency of the last token is still above the
-low-frequency cut-off.  Because all the points on a line share the same
-co-occurrence distribution, the left panel in
-Figure~\ref{fig:word-frequency-experiment-graph} demonstrates conclusively
-that length does indeed depend on frequency directly.  Moreover, this
-relation is seen to be approximately linear for each word considered.
-Notice also that the relative positions of the lengths of the word vectors associated
-with the experiment words are roughly independent of the frequency band,
-i.e., the plotted lines rarely cross.
+so that the frequency of the last token is still above the low-frequency
+cut-off.  Because all the points on a line share the same co-occurrence
+distribution, the left panel in
+Figure~\ref{fig:word-frequency-experiment-graph} demonstrates
+conclusively that length does indeed depend on frequency directly.
+Moreover, this relation is seen to be approximately linear for each word
+considered.  Notice also that the relative positions of the lengths of
+the word vectors associated with the experiment words are roughly
+independent of the frequency band, i.e., the plotted lines rarely cross.
 
@@
 that the lengths of the vectors representing the meaningless tokens
-\word{VOID\_i} are approximately constant (about $2.5$).  Since we already found the
-direction to be also constant, it is sensible to speak of the word
-vector of \word{VOID} irrespective of its frequency.  In particular, the
-token \word{VOID\_1} may be taken as an approximation.
+\word{VOID\_i} are approximately constant (about $2.5$).  Since we
+already found the direction to be also constant, it is sensible to speak
+of the word vector of \word{VOID} irrespective of its frequency.  In
+particular, the token \word{VOID\_1} may be taken as an approximation.
 
@@
 We take the noise distribution to be the (observed) unconditional word
-distribution.  Thus noise can be added to the co-occurrence distribution
+distribution.  Noise can then be added to the co-occurrence distribution
 of a word by simply interspersing occurrences of that word uniformly at
@@
 For each of these words, the replacement procedure of
-Section~\ref{replacement-procedure} is performed using a geometric rate
-of decay of $p = \tfrac{5}{6}$, and truncating the distribution at
-$n=8$.  For every replacement token (e.g. \word{CAT\_i}), occurrences of
-this token are subsequently interspersed uniformly at random throughout
-the corpus, such that the frequency of the replacement token is restored
-to that of the original word \word{cat}.  For example, if the original
-word \word{cat} occurred $1000$ times, then after the replacement
-procedure, \word{CAT\_2} occurs approximately $694$ times, so a further
-(approximately) $306$ random occurrences of \word{CAT\_2} are
-interspersed throughout the corpus.  Thus the word \word{cat} is removed
-from the corpus, and a family of tokens \word{CAT\_i}, $1 \leqslant i
-\leqslant 8$ is introduced.  These tokens all have the same frequency as
-the word \word{cat}, but their co-occurrence distributions, while based
-on that of \word{cat}, have an increasing amount of noise.
+Section~\ref{replacement-procedure} is performed using the
+distribution~(\ref{distro2}) with $n=7$.  For every replacement token
+(e.g. \word{CAT\_i}), additional occurrences of this token are
+interspersed uniformly at random throughout the corpus, such that the
+final frequency of the replacement token is $2/n$ times that of the
+original word \word{cat}.  For example, if the original word \word{cat}
+occurred $1000$ times, then after the replacement procedure,
+\word{CAT\_2} occurs approximately $238$ times, so a further
+(approximately) $2/7 \times 1000 - 238 \approx 48$ random occurrences of
+\word{CAT\_2} are interspersed throughout the corpus.  In this way, the
+word \word{cat} is removed from the corpus and replaced with a family of
+tokens \word{CAT\_i}, $1 \leqslant i \leqslant 7$.  These tokens all
+have the same frequency, but their co-occurrence distributions, while
+based on that of \word{cat}, have an increasing amount of noise.
+Specifically, the proportion of noise for the $i$th token is
+%
+$$ 1 - \frac{n}{2} P_n(i) = \frac{i-1}{n-1} , \;\; \textrm{ or } \;\; 0,
+\frac{1}{n-1}, \frac{2}{n-1}, \ldots, 1 \;\; \textrm{ for } \;\; i =
+1,2, \ldots, n , 
+$$ 
+%
+which is evenly distributed.  The first token contains no noise at all,
+while the last token stands for pure noise.  The particular choice of
+$n$ assures a reasonable coverage of the interval $[0,1]$.  Other
+parameter values (or indeed other distributions) could, of course, have
+been used equally well.
 
@@
 
-The proportion of occurrences of the token \word{CAT\_i} that arise by
-replacing the word \word{cat} is $P_{5/6, 8}(i)$, and the proportion
-that arise from randomly interspersing this token throughout the corpus
-is $r_i := 1 - P_{5/6, 8}(i)$.  The particular value of $p$ (large
-compared to that of Section~\ref{WFVE}) was chosen so that the
-proportion of noise $r_i$ did not increase too quickly with $i$.  These
-proportions are the sampling points on our axis of variation, and this
-value of $p$ (together with the choice of $n$) assures a reasonable
-spread.  Other parameter values (or indeed other distributions) could,
-of course, have been used equally well.
-
 \begin{table}
@@
 	\label{fig:co-occurrence-noise-counts}
-	\caption{Words chosen for the co-occurrence noise experiment, along with the occurrence counts in the unmodified corpus. }
+	\caption{Words chosen for the co-occurrence noise experiment,
+          along with the word frequencies in the unmodified corpus. }
 	\label{cooccurrence-noise-words}
@@
           replacement.  The tokens were generated using the distribution
-          (\ref{distro}) with $p = \tfrac{5}{6}$ and $n=3$.}
+          (\ref{distro2}) with $n=7$.}
 	\label{fig:co-occurrence-noise-experiment-text}
@@
 \begin{subsection}{Experimental results}
+%
 Figure~\ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity
 of pairs of vectors representing some of the tokens used in this
-experiment.  The figure clearly demonstrates that vectors representing
-tokens derived from the same word all share a common direction.  Thus
-co-occurrence noise has no discernible effect on the direction of word
-vectors.
+experiment.  Remember that the first token ($i=1$) in a family is
+without noise in its co-occurrence distribution, while the last one
+($i=n$, with $n=7$) stands for pure noise and has therefore no relation
+anymore with the word it derives from.  The figure demonstrates that the
+vectors within a family only moderately deviate from the original
+direction defined by the first token ($i=1$) when noise is added to the
+co-occurrence distribution.  For $1<i<7$, the deviation typically
+increases with the proportion of noise.  The vector of the last token
+($i=n$), associated with pure noise, is seen within each of the families
+to point in a completely different direction, more or less perpendicular
+to the original one.  To understand this interpolating behavior, recall
+from Section~\ref{WFVE-results} that the vector for the entirely
+meaningless word \word{VOID} is small but non-zero.  Since the noise
+distribution coincides with the co-occurrence distribution of
+\word{VOID}, the vectors for the experiment words must tend to the word
+vector for \word{VOID} as the proportion of noise in their co-occurrence
+distributions approaches $1$.
+This convergence to a common point is only indistinctly apparent in
+Figure~\ref{fig:co-occurrence-noise-heatmap} as the frequency of the experiment
+tokens is insufficient to sample the full variety of the contexts of VOID, i.e.
+all contexts (see Section \ref{WFVE-direction}).
 
 The left panel in Figure~\ref{fig:co-occurrence-noise-graph} reveals
-that word vector length varies more or less linearly with the proportion
-of noise in the co-occurrence distribution of the word.  This figure motivates an
-interpretation of vector length, within a sufficiently narrow frequency
-band, as a measure of the absence of co-occurrence noise, or put
-differently, of the extent to which a word carries the meaning of a
-distinctive context.
-The right panel of Figure~\ref{fig:co-occurrence-noise-graph} shows that the
-syn1neg vector length does not vary with the proportion of co-occurrence noise.
-
-Recall from Section~\ref{WFVE-results} that the word vector for \word{VOID} is independent of the frequency of \word{VOID}, and is non-zero.
-Since the noise distribution coincides with the co-occurrence distribution of \word{VOID}, the word vectors for the experiment words must tend to the word vector for \word{VOID} as the proportion of noise in their co-occurrence distributions approaches $1$.
-This is not discernible in Figure~\ref{fig:co-occurrence-noise-heatmap} since the word vector for \word{VOID} is relatively close to the origin.
+that vector length varies more or less linearly with the proportion of
+noise in the co-occurrence distribution of the word.  This figure
+motivates an interpretation of vector length, within a sufficiently
+narrow frequency band, as a measure of the absence of co-occurrence
+noise, or put differently, of the extent to which a word carries the
+meaning of a distinctive context.
 
@@
           noise experiment (the words were chosen at random).  The
-          largely red blocks demonstrate that the direction of the word
-          vectors does not change when noise is added to the
-          co-occurrence distribution.  }
+          largely red blocks demonstrate that for $i<7$ the direction of
+          the vectors only moderately changes when noise is added to the
+          co-occurrence distribution.  The vector of the tokens
+          associated with pure noise ($i=7$) is seen to be almost
+          perpendicular to the word vectors they derive from. }
 	\label{fig:co-occurrence-noise-heatmap}

### 2015-10-01 article/word-frequency-experiment-text-cat.tex
integration of the corrected noise experiment into article text

@@
-\texttt{the domestic \textcolor{red}{CAT\_1} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_1} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_1} is one whose ancestry is recorded by a \textcolor{red}{CAT\_2} fancier organization\newline a purebred \textcolor{red}{CAT\_1} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_3} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_3} is usually dominant over the others}
+\texttt{the domestic \textcolor{red}{CAT\_2} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_1} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_1} is one whose ancestry is recorded by a \textcolor{red}{CAT\_2} fancier organization\newline a purebred \textcolor{red}{CAT\_2} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_4} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_1} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} is usually dominant over the others}

### 2015-10-01 article/word-frequency-experiment-text-void.tex
integration of the corrected noise experiment into article text

@@
-\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline \textcolor{red}{VOID\_1} within groups one cat is usually dominant over the others}
+\texttt{\textcolor{red}{VOID\_1} the domestic cat was first classified as felis catus\newline the semiferal cat \textcolor{red}{VOID\_3} a mostly outdoor cat is not \textcolor{red}{VOID\_2} owned by \textcolor{red}{VOID\_1} any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among \textcolor{red}{VOID\_1} mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others}

### 2015-10-01 article_generate_sample_texts.py

@@ -65,9 +65,8 @@ tmp_file1.seek(0)
 token_freq_dict = dict()
 target_freq = counts[word] * 1. * coocc_noise_experiment_freq_reduction / total_words
 for i in range(1, coocc_noise_experiment_max_value + 1):
-    current_freq = counts[word] * evenly_spaced_proba(i, coocc_noise_experiment_max_value)
+    current_freq = counts[word] * evenly_spaced_proba(i, coocc_noise_experiment_max_value) / total_words
     token_freq_dict[build_experiment_token(word, i)] = target_freq - current_freq
-
 tmp_file2 = StringIO()
 intersperse_words(token_freq_dict, tmp_file1, tmp_file2)
 

### 2015-10-06 audit_modifications.py
simple script for checking the word counts

@@ -1,0 +1,34 @@
+"""
+A simple script for outputing the frequencies of the experiment tokens.
+"""
+from parameters import *
+from functions import *
+
+with file('outputs/word_counts_unmodified_corpus.csv') as f:
+    wc_unmodified = read_word_counts(f)
+            
+with file('outputs/word_counts_modified_corpus.csv') as f:
+    wc_modified = read_word_counts(f)
+
+wfve_words = read_words('outputs/word_freq_experiment_words')
+cnve_words = read_words('outputs/coocc_noise_experiment_words')
+
+print 'word freq variation experiment token frequencies:'
+for word in wfve_words:
+    print 'token "%s" with original frequency %i' % (word, wc_unmodified[word])
+    for index in range(1, word_freq_experiment_power_max + 1):
+        token = build_experiment_token(word, index)
+        freq = wc_modified[token] if token in wc_modified else 0
+        print token, freq
+    print '-' * 80
+print '=' * 80
+
+print 'coocc noise variation experiment token frequencies:'
+for word in cnve_words:
+    print 'token "%s" with original frequency %i' % (word, wc_unmodified[word])
+    for index in range(1, coocc_noise_experiment_max_value + 1):
+        token = build_experiment_token(word, index)
+        freq = wc_modified[token] if token in wc_modified else 0
+        print token, freq
+    print '-' * 80
+print '=' * 80

### 2015-10-08 build_images.py
split the build images script

@@ -1,186 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib; matplotlib.use('Agg') # must be set before importing pyplot
-import matplotlib.pyplot as plt
-import random
-import sys
-
-from parameters import *
-from functions import *
-
-random.seed(1) # fix seed so that we can refer to the randomly chosen words in the article body
-
-vectors_syn0_filename = sys.argv[1]
-vectors_syn1neg_filename = sys.argv[2]
-word_counts_filename = sys.argv[3]
-word_freq_exp_words_filename = sys.argv[4]
-coocc_noise_exp_words_filename = sys.argv[5]
-
-matplotlib.rcParams.update({'font.size': 14})
-matplotlib.rcParams.update({'figure.autholayout': True})
-
-word_freq_experiment_words = read_words(word_freq_exp_words_filename)
-coocc_noise_experiment_words = read_words(coocc_noise_exp_words_filename)
-
-vectors_syn0 = load_word2vec_binary(vectors_syn0_filename)
-norms_syn0 = np.sqrt((vectors_syn0 ** 2).sum(axis=1))
-
-vectors_syn1neg = load_word2vec_binary(vectors_syn1neg_filename)
-norms_syn1neg = np.sqrt((vectors_syn1neg ** 2).sum(axis=1))
-
-vocab = list(vectors_syn0.index)
-
-with file(word_counts_filename) as f:
-    new_counts = read_word_counts(f)
-total_words = sum(new_counts.values())
-
-stats = pd.DataFrame({'occurrences': new_counts, 'L2_norm_syn0': norms_syn0, 'L2_norm_syn1neg': norms_syn1neg}).dropna()
-stats['occurrences_band'] = np.floor(np.log2(stats.occurrences)).astype(int)
-stats['log2_frequency'] = np.log2(stats.occurrences * 1. / total_words)
-stats.L2_norm_syn0.name = 'vector length (syn0)'
-stats.L2_norm_syn1neg.name = 'vector length (syn1neg)'
-
-
-# WORD FREQUENCY EXPERIMENT
-
-def wf_experiment_tokens(word):
-    """
-    return the list of tokens introduced into the corpus in the word frequency
-    experiment for the given word.
-    """
[…]

### 2015-10-08 build_images_coocc_noise.py

@@ -1,0 +1,104 @@
+"""
+Given the vectors and word counts resulting from the experiments, build
+graphics for the cooccurrence noise variation experiment section of the
+article.
+"""
+import numpy as np
+import pandas as pd
+import matplotlib; matplotlib.use('Agg') # must be set before importing pyplot
+import matplotlib.pyplot as plt
+import random
+import sys
+
+from parameters import *
+from functions import *
+
+random.seed(1) # fix seed so that we can refer to the randomly chosen words in the article body
+
+vectors_syn0_filename = sys.argv[1]
+vectors_syn1neg_filename = sys.argv[2]
+word_counts_filename = sys.argv[3]
+coocc_noise_exp_words_filename = sys.argv[4]
+
+matplotlib.rcParams.update({'font.size': 14})
+matplotlib.rcParams.update({'figure.autholayout': True})
+
+coocc_noise_experiment_words = read_words(coocc_noise_exp_words_filename)
+
+def calculate_norms(vecs):
+    return np.sqrt((vecs ** 2).sum(axis=1))
+
+vectors_syn0 = load_word2vec_binary(vectors_syn0_filename)
+norms_syn0 = calculate_norms(vectors_syn0)
+
+vectors_syn1neg = load_word2vec_binary(vectors_syn1neg_filename)
+norms_syn1neg = calculate_norms(vectors_syn1neg)
+
+vocab = list(vectors_syn0.index)
+
+with file(word_counts_filename) as f:
+    counts_modified_corpus = read_word_counts(f)
+
+stats = pd.DataFrame({'occurrences': counts_modified_corpus,
+                      'L2_norm_syn0': norms_syn0,
+                      'L2_norm_syn1neg': norms_syn1neg}).dropna()
+
+# WORD VECTOR LENGTH AS A FUNCTION OF NOISE PROPORTION
+
+# reorder by the left-most value of the corresponding syn0 plot
+# i.e. by word vector length of the noiseless token
+words = sorted(coocc_noise_experiment_words,
+               key=lambda word: stats.L2_norm_syn0.loc[build_experiment_token(word, 1)],
+               reverse=True)
+
+markers = [['o', 's', 'D'][i % 3] for i in range(len(words))]
[…]

### 2015-10-08 build_images_word_frequency.py

@@ -1,0 +1,125 @@
+"""
+Given the vectors and word counts resulting from the experiments, build
+graphics for the word frequency variation experiment section of the article.
+"""
+import numpy as np
+import pandas as pd
+import matplotlib; matplotlib.use('Agg') # must be set before importing pyplot
+import matplotlib.pyplot as plt
+import random
+import sys
+
+from parameters import *
+from functions import *
+
+random.seed(1) # fix seed so that we can refer to the randomly chosen words in the article body
+
+vectors_syn0_filename = sys.argv[1]
+vectors_syn1neg_filename = sys.argv[2]
+word_counts_filename = sys.argv[3]
+word_freq_exp_words_filename = sys.argv[4]
+
+matplotlib.rcParams.update({'font.size': 14})
+matplotlib.rcParams.update({'figure.autholayout': True})
+
+word_freq_experiment_words = read_words(word_freq_exp_words_filename)
+
+def calculate_norms(vecs):
+    return np.sqrt((vecs ** 2).sum(axis=1))
+
+vectors_syn0 = load_word2vec_binary(vectors_syn0_filename)
+norms_syn0 = calculate_norms(vectors_syn0)
+
+vectors_syn1neg = load_word2vec_binary(vectors_syn1neg_filename)
+norms_syn1neg = calculate_norms(vectors_syn1neg)
+
+vocab = list(vectors_syn0.index)
+
+with file(word_counts_filename) as f:
+    counts_modified_corpus = read_word_counts(f)
+
+stats = pd.DataFrame({'occurrences': counts_modified_corpus,
+                      'L2_norm_syn0': norms_syn0,
+                      'L2_norm_syn1neg': norms_syn1neg}).dropna()
+
+def wf_experiment_tokens(word):
+    """
+    return the list of tokens introduced into the corpus in the word frequency
+    experiment for the given word for which we have word vectors.
+    """
+    idxs = [build_experiment_token(word, i) for i in range(1, word_freq_experiment_power_max + 1)]
+    return [idx for idx in idxs if idx in vectors_syn0.index]
+
+# COSINE SIMILARITY HEATMAP
+example_words = random.sample([word for word in word_freq_experiment_words if word != 'the'], 4)
[…]

### 2015-10-08 functions.py
tidy up functions.py

@@ -2,10 +2,9 @@ from __future__ import print_function
 import random
 import numpy as np
 import pandas as pd
-from scipy.spatial.distance import pdist, cdist
 import matplotlib.pyplot as plt
 from collections import defaultdict
-from parameters import random_seed, experiment_word_occurrence_min, coocc_noise_experiment_freq_reduction
+from parameters import coocc_noise_experiment_freq_reduction
 
 def count_words(f):
     """
@@ -18,28 +17,14 @@ def count_words(f):
                 counts[word] += 1
     return counts
 
-def count_document_frequency(f, vocab):
-    """
-    Return a dictionary mapping each word to the number of documents it occurs
-    in, in the file 'f' specified.  Documents are separated from one another by
-    linefeeds.
-    Words are separated by ' '.
-    """
-    counts = {word: 0 for word in vocab}
-    for line in f:
-        for word in set(line.strip().split(' ')):
-                counts[word] += 1
-    return counts
-
-def read_words(filename):
+def read_words(f):
     """
     given a CSV of word counts, returns the list of words.
     """
     words = []
-    with file(filename) as f:
-        for line in f:
-            word = line.split(',')[0]
-            words.append(word)
+    for line in f:
+        word = line.split(',')[0]
+        words.append(word)
     return words
 
 def read_word_counts(f):
@@ -60,19 +45,21 @@ def build_experiment_token(word, sample):
 
 def truncated_geometric_proba(ratio, i, n):
     """
-    return the probability of i being sampled from [1 .. n] from the truncated geometric distribution with the given ratio,
-    i.e. the unique distn such that the probabilities decrease by ratio each time, and all are non-zero.
+    return the probability of i being sampled from [1 .. n] from the truncated
+    geometric distribution with the given the ratio, i.e. the unique distn such
+    that the probabilities decrease by ratio each time, and all are non-zero.
     """
     return (ratio ** (i - 1)) * (1 - ratio) / (1 - ratio ** n)
 
 def distribution_to_sampling_function(word, dist_fn, max_value):
     """
-    returns a function that samples from the given distribution on [1 ..
-    max_value], returning WORD_1, .. WORD_<max_value>
+    returns a function that samples from the given distribution 'dist_fn' on
[…]

### 2015-10-08 modify_corpus_coocc_noise_experiment.py

@@ -1,7 +1,9 @@
 """
-Modifies an input text for the experiments according to the parameters defined in parameters.py
+Modifies an input text for the cooccurrence noise variation experiment
+according to the parameters defined in parameters.py
 Reads from stdin, writes to stdout.
-Requires sufficient diskspace to write out the modified text at intermediate steps.
+Requires sufficient diskspace to write out the modified text at intermediate
+steps.
 """
 import os
 import sys
@@ -9,8 +11,7 @@ from parameters import *
 from functions import *
 
 cn_experiment_words = read_words(sys.argv[1])
-with file(sys.argv[2]) as f:
-    counts = read_word_counts(f)
+counts = read_word_counts(sys.argv[2])
 total_words = sum(counts.values())
 
 # perform the replacement procedure
@@ -19,16 +20,17 @@ distn = lambda i: evenly_spaced_proba(i, coocc_noise_experiment_max_value)
 for word in cn_experiment_words:
     word_samplers[word] = distribution_to_sampling_function(word, distn, coocc_noise_experiment_max_value)
 
-intermediate_file = 'delete.me.word_freq_experiment'
+intermediate_file = 'delete.me.coocc_noise_experiment'
 with open(intermediate_file, 'w') as f_out:
     replace_words(word_samplers, sys.stdin, f_out)
 
 # add noise to the cooccurrence distributions
 token_freq_dict = dict()
 for word in cn_experiment_words:
-    target_freq = counts[word] * 1. * coocc_noise_experiment_freq_reduction / total_words
+    original_freq = counts[word] * 1. / total_words
+    target_freq = original_freq * coocc_noise_experiment_freq_reduction
     for i in range(1, coocc_noise_experiment_max_value + 1):
-        current_freq = counts[word] * evenly_spaced_proba(i, coocc_noise_experiment_max_value) / total_words
+        current_freq = original_freq * distn(i)
         token_freq_dict[build_experiment_token(word, i)] = target_freq - current_freq
 
 with open(intermediate_file) as f_in:

### 2015-10-08 modify_corpus_word_freq_experiment.py

@@ -1,7 +1,9 @@
 """
-Modifies an input text for the word frequency experiment.
+Modifies an input text for the word frequency experiment according to the
+parameters defined in parameters.py
 Reads from stdin, writes to stdout.
-Requires sufficient diskspace to write out the modified text at intermediate steps.
+Requires sufficient diskspace to write out the modified text at intermediate
+steps.
 """
 import os
 import sys
@@ -9,12 +11,11 @@ from parameters import *
 from functions import *
 
 wf_experiment_words = read_words(sys.argv[1])
-with file(sys.argv[2]) as f:
-    counts = read_word_counts(f)
+counts = read_word_counts(sys.argv[2])
 total_words = sum(counts.values())
 
 # intersperse the meaningless token throughout the corpus
-intermediate_file = 'delete.me.coocc_experiment'
+intermediate_file = 'delete.me.word_freq_experiment'
 with open(intermediate_file, 'w') as f_out:
     intersperse_words({meaningless_token: meaningless_token_frequency}, sys.stdin, f_out)
 wf_experiment_words.append(meaningless_token)

### 2015-10-08 article_generate_sample_texts.py
tidy modification example generation and fix makefile

@@ -18,7 +18,7 @@ def write_markup(raw_output, f_out):
     f_out.write(output)
     f_out.write(r"}")
 
-# PERFORM THE REPLACEMENT PROCEDURE FOR THE WORD FREQUENCY EXPERIMENT FOR THE WORD "CAT"
+# PERFORM THE REPLACEMENT PROCEDURE FOR THE WORD FREQUENCY VARIATION EXPERIMENT FOR THE WORD "CAT"
 
 word = 'cat'
 distn = lambda i: truncated_geometric_proba(word_freq_experiment_ratio, i, word_freq_experiment_power_max)
@@ -47,7 +47,7 @@ replace_words(word_samplers, tmp_file1, tmp_file2)
 with file('article/word-frequency-experiment-text-void.tex', 'w') as f_out:
     write_markup(tmp_file2.getvalue(), f_out)
 
-# COOCCURRENCE NOISE EXPERIMENT
+# PERFORM THE PROCEDURE FOR THE COOCCURRENCE NOISE VARIATION EXPERIMENT FOR THE WORD "CAT"
 word = 'cat'
 distn = lambda i: evenly_spaced_proba(i, coocc_noise_experiment_max_value)
 word_samplers = {word: distribution_to_sampling_function(word, distn, coocc_noise_experiment_max_value)}
@@ -63,9 +63,10 @@ tmp_file1.seek(0)
 
 # add noise to the cooccurrence distribution
 token_freq_dict = dict()
-target_freq = counts[word] * 1. * coocc_noise_experiment_freq_reduction / total_words
+original_freq = counts[word] * 1. / total_words
+target_freq = original_freq * coocc_noise_experiment_freq_reduction
 for i in range(1, coocc_noise_experiment_max_value + 1):
-    current_freq = counts[word] * evenly_spaced_proba(i, coocc_noise_experiment_max_value) / total_words
+    current_freq = original_freq * distn(i) 
     token_freq_dict[build_experiment_token(word, i)] = target_freq - current_freq
 tmp_file2 = StringIO()
 intersperse_words(token_freq_dict, tmp_file1, tmp_file2)

### 2015-10-08 article/main.tex
related work to that of schnabel et al 2015

@@
 propose to modify the training corpus in a controlled fashion.  To
-this end, we insert new tokens into the corpus with varying frequencies
+this end, we insert new pseudowords into the corpus with varying frequencies
 and varying levels of noise in their co-occurrence distributions.  By
-modeling the frequency and co-occurrence distributions of these tokens
+modeling the frequency and co-occurrence distributions of these pseudowords
 on existing words in the corpus, we are able to study their effect on
 word vectors independently of one another.  We can thus study a family
-of tokens that all appear in the same context, but with different
-frequencies, or study a family of tokens that all have the same
+of pseudowords that all appear in the same context, but with different
+frequencies, or study a family of pseudowords that all have the same
 frequency, but appear in a different number of contexts.  Starting from
@@
 co-occurrence distribution.  Surprisingly, this special vector in the feature
-space, obtained by interspersing a token uniformly at random throughout
+space, obtained by interspersing a pseudoword uniformly at random throughout
 the corpus with a frequency sufficiently large to sample all contexts, is
@@
 \begin{section}{Related work}\label{related-work}
-\mar{move to other paper?}
-Our notion of word significance is reminiscent of the concept of
-``contextual distinctiveness'' introduced earlier by McDonald and
-Shillcock \cite{mcdonald2001contextual}.  They define the contextual
-distinctiveness of a target word $w$ as the Kullback-Leibler divergence
-between the unconditional word distribution $\p(w')$, and the
-co-occurrence distribution $\p(w' | w)$ of the target word $w$.  So
-defined, contextual distinctiveness is a measure of the amount of
-information provided by word $w$ about its contexts of use.  More
-specifically, if a word $w$ tends to appear in a wide variety of
-contexts, then its co-occurrence distribution tends to diverge less from
-unconditional word distribution, and its contextual distinctiveness is
-accordingly small.  If, on the other hand, $w$ typically appears in only
-a small number of different contexts, its co-occurrence distribution
-tends to diverge more from the unconditional word distribution, and a
-larger value of contextual distinctiveness results.  To reduce the
-computational burden, the authors considered only $690$ target words,
-randomly chosen from equally divided frequency bands.  Furthermore, the
-authors considered only co-occurrence with the $500$ most frequent
-content words.  This significant restriction, necessary in their case in
-order to reduce the variance of the estimates of their measure of
-contextual distinctiveness, does not apply in our study.
+In recent work on the evaluation of word embeddings, Schnabel, Labutov, Mimno
+and Joachims \cite{schnabelemnlp2015} trained logistic regression models to
+predict whether a word was rare or frequent given only the direction of its
+word vector.  The prediction accuracy was then measured as a function of word
+embedding method and of the threshold for word rarity.  It was found in the
+case of word2vec CBOW that word vector direction could be used to distinguish
+very rare words from all other words. The authors claimed further that word
+vector direction must encode frequency directly, and not indirectly via
+semantic information.  Our experimental finding that word vector direction is
+independent of word frequency contradicts this claim in the case of word2vec
+CBOW.
 
-Our experimental finding that word vector length decreases with
-co-occurrence noise is related to earlier work by Vecchi, Baroni, and
-Zamparelli \cite{vecchi-baroni-zamparelli2011}, where a relation between
-vector length and the ``semantic deviance'' of an adjective-noun
-composite was studied empirically.  In that paper, which is also based
-on word co-occurrence statistics, the authors study adjective-noun
-composites.  They built a vocabulary from the 8k most frequent nouns and
-4k most frequent adjectives in a large general language corpus and added
-22k adjective-noun composites.  For each item in the vocabulary, they
-recorded the co-occurrences with the top 10k most frequent content words
-(nouns, adjectives or verbs), and constructed word embeddings via
-singular value decomposition of the co-occurrence matrix
+Our experimental finding that word vector length decreases with co-occurrence
+noise is related to earlier work by Vecchi, Baroni, and Zamparelli
+\cite{vecchi-baroni-zamparelli2011}, where a relation between vector length and
+the ``semantic deviance'' of an adjective-noun composite was studied
+empirically.  In that paper, which is also based on word co-occurrence
+statistics, the authors study adjective-noun composites.  They built a
+vocabulary from the 8k most frequent nouns and 4k most frequent adjectives in a
+large general language corpus and added 22k adjective-noun composites.  For
+each item in the vocabulary, they recorded the co-occurrences with the top 10k
+most frequent content words (nouns, adjectives or verbs), and constructed word
+embeddings via singular value decomposition of the co-occurrence matrix
 \cite{landauer-dumais1997}.  The authors considered several models for
-constructing vectors of unattested adjective-noun composites, the two
-simplest being adding and component-wise multiplying the adjective and
-noun vectors.  They hypothesized that the length of the vectors thus
-constructed can be used to distinguish acceptable and semantically
-deviant adjective-noun composites.  Using a few hundred adjective-noun
-composites selected by humans for evaluation, they found that deviant
-composites have a shorter vector than acceptable ones, in accordance
-with their expectation.  In contrast to their work, our approach does
-not require human annotation.
+constructing vectors of unattested adjective-noun composites, the two simplest
+being adding and component-wise multiplying the adjective and noun vectors.
+They hypothesized that the length of the vectors thus constructed can be used
+to distinguish acceptable and semantically deviant adjective-noun composites.
+Using a few hundred adjective-noun composites selected by humans for
+evaluation, they found that deviant composites have a shorter vector than
+acceptable ones, in accordance with their expectation.  In contrast to their
+work, our approach does not require human annotation.
 
 Recent theoretical work \cite{Arora2015} has approached the problem of
-explaining the so-called ``compositionality'' property exhibited by some
-word embeddings.  In that work, unnormalised vectors are used in their
-model of the word relation task.  It is hoped that experimental
-approaches such as those described here might enable theoretical
-investigations to describe the role of the word vector length in the
-word relation tasks.
-\end{section}
+explaining the so-called ``compositionality'' property exhibited by some word
+embeddings.  In that work, unnormalised vectors are used in their model of the
+word relation task.  It is hoped that experimental approaches such as those
+described here might enable theoretical investigations to describe the role of
+the word vector length in the word relation tasks.  \end{section}
 
-\begin{section}{Corpus and model}\label{corpus-and-model}
-Our training data is built from the Wikipedia data dump from October
-2013.  To remove the bulk of robot-generated pages from the training
-data, only pages with at least 20 monthly page views are
-retained.\footnote{For further justification and to obtain the dataset,
-  see\\ \blogpost} Stubs and disambiguation pages are also removed,
-leaving 463 thousand pages with a total of 482 million words.
-Punctuation marks and numbers were removed from the pages and all words
-were lower-cased.  Word frequencies are summarised in
-Table~\ref{word-occurrences-table}.  This base corpus is then modified
-as described in Sections~\ref{WFVE} and \ref{CNVE}.  For
-recognisability, the tokens inserted into the corpus are upper-cased.
+\begin{section}{Corpus and model}\label{corpus-and-model} Our training data is
+	built from the Wikipedia data dump from October 2013.  To remove the
+	bulk of robot-generated pages from the training data, only pages with
+	at least 20 monthly page views are retained.\footnote{For further
+	justification and to obtain the dataset, see\\ \blogpost} Stubs and
+	disambiguation pages are also removed, leaving 463 thousand pages with
+	a total of 482 million words.  Punctuation marks and numbers were
+	removed from the pages and all words were lower-cased.  Word
+	frequencies are summarised in Table~\ref{word-occurrences-table}.  This
+	base corpus is then modified as described in Sections~\ref{WFVE} and
+	\ref{CNVE}.  For recognisability, the pseudowords inserted into the
+	corpus are upper-cased.
 
-\begin{subsection}{Word2vec}\label{word2vec}
-Word2vec, a feed-forward neural network with a single hidden layer,
-learns word vectors from word co-occurrences in an unsupervised manner.
-Word2vec comes in two versions.  In the continuous bag-of-words (CBOW)
-model, the words appearing around a target word serve as input.  That
-input is projected linearly onto the hidden layer and the network then
-attempts to predict the target word on output.  Training is achieved
-through back-propagation.  The word vectors are encoded in the weights
-of the first synaptic layer, ``syn0''.  The weights of the second
-synaptic layer (``syn1neg'', in the case of negative sampling) are
-typically discarded.  In the other model, called skip-gram, target and
-context words swap places, so that the target word now serves as input,
-while the network attempts to predict the context words on output.
+\begin{subsection}{Word2vec}\label{word2vec} Word2vec, a feed-forward neural
+	network with a single hidden layer, learns word vectors from word
+	co-occurrences in an unsupervised manner.  Word2vec comes in two
+	versions.  In the continuous bag-of-words (CBOW) model, the words
+	appearing around a target word serve as input.  That input is projected
+	linearly onto the hidden layer and the network then attempts to predict
+	the target word on output.  Training is achieved through
+	back-propagation.  The word vectors are encoded in the weights of the
+	first synaptic layer, ``syn0''.  The weights of the second synaptic
+	layer (``syn1neg'', in the case of negative sampling) are typically
+	discarded.  In the other model, called skip-gram, target and context
+	words swap places, so that the target word now serves as input, while
+	the network attempts to predict the context words on output.
 
-For simplicity only the word2vec CBOW word embedding with a single set
-of hyperparameters is considered.  Specifically, a CBOW model with a
-hidden layer of size 100 is trained using negative sampling with 5
-negative samples, a window size of $10$, a minimum frequency of
-$128$, and $10$ passes through the corpus.  Sub-sampling was not used so
-that the influence of word frequency could be more clearly discerned.
-Similar experimental results were obtained using hierarchical softmax,
-but these are omitted for succinctness.  The relatively high
-low-frequency cut-off is chosen to ensure that word vectors, in all but
-degenerate cases, receive a sufficient number of gradient updates to be
-meaningful.  This frequency cut-off results in a vocabulary of $81117$
-words (only unigrams were considered).
+For simplicity only the word2vec CBOW word embedding with a single set of
+hyperparameters is considered.  Specifically, a CBOW model with a hidden layer
+of size 100 is trained using negative sampling with 5 negative samples, a
+window size of $10$, a minimum frequency of $128$, and $10$ passes through the
+corpus.  Sub-sampling was not used so that the influence of word frequency
+could be more clearly discerned.  Similar experimental results were obtained
+using hierarchical softmax, but these are omitted for succinctness.  The
+relatively high low-frequency cut-off is chosen to ensure that word vectors, in
+all but degenerate cases, receive a sufficient number of gradient updates to be
+meaningful.  This frequency cut-off results in a vocabulary of $81117$ words
+(only unigrams were considered).
 
-The most recent revision of word2vec was used.\footnote{SVN revision 42,
-  see \url{http://word2vec.googlecode.com/svn/trunk/}} The source code
-for performing the experiments is made available on
+The most recent revision of word2vec was used.\footnote{SVN revision 42, see
+\url{http://word2vec.googlecode.com/svn/trunk/}} The source code for performing
+the experiments is made available on
 GitHub.\footnote{\url{https://github.com/benjaminwilson/word2vec-norm-experiments}}
@@
 In the experiments detailed below, we modify the corpus in a controlled
-manner by introducing tokens into the corpus via a replacement
+manner by introducing pseudowords into the corpus via a replacement
 procedure.  For the frequency experiment, the procedure is as follows.
@@
 geometric distribution, and that occurrence of the word \word{cat} is
-replaced with the token \word{CAT\_i}.  In this way, the word \word{cat}
-is replaced throughout the corpus by a family of tokens with varying
+replaced with the pseudoword \word{CAT\_i}.  In this way, the word \word{cat}
+is replaced throughout the corpus by a family of pseudowords with varying
 frequencies but approximately the same co-occurrence distribution as
-\word{cat}.  That is, all these tokens are used in roughly the same
+\word{cat}.  That is, all these pseudowords are used in roughly the same
 contexts as the original word.
 
-The geometric distribution is truncated to limit the number of tokens
+The geometric distribution is truncated to limit the number of pseudowords
 inserted into the corpus.  For any choice $0 < p < 1$ and maximum value
@@
 the word embedding.  Using the replacement procedure, we introduce a
-small number of families of tokens into the corpus.  The tokens in each
+small number of families of pseudowords into the corpus.  The pseudowords in each
 family vary in frequency but, replacing a single word, all share a
@@
 word frequency in isolation, everything else being kept equal.  We
-consider two types of tokens.
+consider two types of pseudowords.
 
-\begin{subsection}{Tokens derived from existing words}\label{WFVEexisting}
+\begin{subsection}{Pseudowords derived from existing words}\label{WFVEexisting}
 We choose uniformly at random a small number of words from the
 unmodified vocabulary for our experiment.  In order that the inserted
-tokens do not have too low a frequency, only words which occur at least
+pseudowords do not have too low a frequency, only words which occur at least
 10 thousand times are chosen.  We also include the high-frequency
@@
 performed for each of these words, using a geometric decay rate of $p =
-\tfrac{1}{2}$, and maximum value $n=20$, so that the 1st token is
+\tfrac{1}{2}$, and maximum value $n=20$, so that the 1st pseudoword is
 inserted with a probability of about $0.5$, the 2nd with a probability of
 about $0.25$, and so on.  This value of $p$ is one of a range of values
-that ensure that, for each word, multiple tokens will be inserted with a
+that ensure that, for each word, multiple pseudowords will be inserted with a
 frequency sufficient to survive the low-frequency cut-off of
 128.  A maximum value $n=20$ suffices for this choice of $p$, since
-$2^{20 + \log_2{128}}$ exceeds the maximum frequency of any token in
+$2^{20 + \log_2{128}}$ exceeds the maximum frequency of any word in
 the corpus.  Figure~\ref{fig:word-frequency-experiment-text-cat}
 illustrates the effect of these modifications on a sample text, with a
-family of tokens \word{CAT\_i}, derived from the word \word{cat}.
+family of pseudowords \word{CAT\_i}, derived from the word \word{cat}.
 Notice that all occurrences of the word \word{cat} have been replaced
-with the tokens \word{CAT\_i}.
+with the pseudowords \word{CAT\_i}.
 
@@
           experiment as per Section~\ref{WFVEexisting}, where the word
-          \word{cat} is replaced with tokens \word{CAT\_i} using the
+          \word{cat} is replaced with pseudowords \word{CAT\_i} using the
           truncated geometric distribution (\ref{distro}) with
@@
 \label{WFVEmeaningless}
-Whereas the tokens introduced above all replace an existing word that
+Whereas the pseudowords introduced above all replace an existing word that
 carries a meaning, we now include for comparison a high-frequency,
@@
 these modifications on a sample text, where a higher relative frequency
-of $0.05$ is used instead for illustrative purposes.  Although the probability of inserting token
-\word{VOID\_1} is twice that of inserting token \word{VOID\_2} and four
-times that of inserting token \word{VOID\_3}, these three tokens all
-appear once in this small piece of text.
+of $0.05$ is used instead for illustrative purposes.
+%Although the probability of inserting the pseudoword
+%\word{VOID\_1} is twice that of inserting the pseudoword \word{VOID\_2} and four
+%times that of inserting the pseudoword \word{VOID\_3}, these three pseudowords all
+%appear once in this small piece of text.
 
@@
           instead of the word \word{cat} now the meaningless word
-          \word{VOID} is replaced with tokens \word{VOID\_i}.  For
+          \word{VOID} is replaced with pseudowords \word{VOID\_i}.  For
           illustrative purposes, the meaningless word \word{VOID} was
@@
 Figure~\ref{word-frequency-experiment-heatmap} shows the cosine
-similarity of pairs of vectors representing some of the tokens used in
+similarity of pairs of vectors representing some of the pseudowords used in
 this experiment.  Recall that the cosine similarity measures the extent
 to which two vectors have the same direction, taking a maximum value of
-$1$ and a minimum value of $-1$.  The number of different tokens
+$1$ and a minimum value of $-1$.  The number of different pseudowords
 associated with an experiment word is the number of times that its
-frequency can be halved and remains above the low-frequency cut-off of
+frequency can be halved and remain above the low-frequency cut-off of
 $128$.
@@
 distributions.  We therefore expect that the vectors representing the
-tokens derived from any of these words can be properly learned from the
+pseudowords derived from any of these words can be properly learned from the
 modified corpus, even though their combined frequency equals the
@@
 Figure~\ref{word-frequency-experiment-heatmap} clearly demonstrate that
-the direction of these vectors is constant for each family of tokens.
+the direction of these vectors is constant for each family of pseudowords.
 In other words, the direction of these vectors is only determined by the
@@
 vector representing \word{VOID\_i} or \word{THE\_i} increases with $i$,
-for the tokens occur increasingly less often.  This is consistent with
+for the pseudowords occur increasingly less often.  This is consistent with
 Figure~\ref{word-frequency-experiment-heatmap}, where it is apparent
@@
 true for higher values of $i$.  The same trend is apparent, although
-less extremely so, for the word \word{the}.  The vectors for the tokens
+less extremely so, for the word \word{the}.  The vectors for the pseudowords
 \word{THE\_i} share a common direction for values of $i$ as high as
@@
 deviations are artifacts of poor learning due to an insufficient number
-of occurrences of these tokens.
+of occurrences of these pseudowords.
 Table~\ref{word-frequency-experiment-cosine-similarity}, showing the
-most similar words to each token \word{THE\_i}, supports this further.
-In the discussions below, the tokens \word{VOID\_i} and \word{THE\_i}
+most similar words to each pseudoword \word{THE\_i}, supports this further.
+In the discussions below, the pseudowords \word{VOID\_i} and \word{THE\_i}
 whose vectors have a cosine similarity less than $0.8$ with
@@
 completeness.  Each line corresponds to a single word, and the points on
-each line indicate the frequency and vector length of the tokens derived
+each line indicate the frequency and vector length of the pseudowords derived
 from that word.  For example, the six points on the line corresponding
 to the word \word{protestant} are labeled, from right to left, by the
-tokens \word{PROTESTANT\_1}, \word{PROTESTANT\_2}, \dots,
+pseudowords \word{PROTESTANT\_1}, \word{PROTESTANT\_2}, \dots,
 \word{PROTESTANT\_6}.  Again, the number of points on the line is
@@
 frequency of the word \word{protestant} can be halved at most $6$ times
-so that the frequency of the last token is still above the low-frequency
+so that the frequency of the last pseudoword is still above the low-frequency
 cut-off.  Because all the points on a line share the same co-occurrence
@@
 approximately $10^6$.  We have no explanation of this.  Also observe
-that the lengths of the vectors representing the meaningless tokens
+that the lengths of the vectors representing the meaningless pseudowords
 \word{VOID\_i} are approximately constant (about $2.5$).  Since we
@@
 of the word vector of \word{VOID} irrespective of its frequency.  In
-particular, the token \word{VOID\_1} may be taken as an approximation.
+particular, the vector of the pseudoword \word{VOID\_1} may be taken as an approximation.
 
@@
 	\caption{ Heatmap of the cosine similarity of the vectors
-          representing some of the tokens used in the word frequency
+          representing some of the pseudowords used in the word frequency
           experiment.  The words other than \word{the} and \word{VOID}
@@
 	\caption{ Words in the original vocabulary most similar to the
-          tokens \word{THE\_i}, and their cosine similarity with the
-          most frequent such token, \word{THE\_1}.  It is apparent from
+          pseudowords \word{THE\_i}, and their cosine similarity with the
+          most frequent such pseudoword, \word{THE\_1}.  It is apparent from
           the nearest neighbour list that the vectors of the
-          low-frequency tokens have not been adequately trained.  }
+          low-frequency pseudowords have not been adequately trained.  }
 	\label{word-frequency-experiment-cosine-similarity}
@@
 	\centering{\includegraphics[scale=0.6]{word-frequency-experiment-graph}}
-	\caption{ Vector length vs.\ frequency for tokens derived from a
-          few words chosen at random.  For each word, tokens of varying
+	\caption{ Vector length vs.\ frequency for pseudowords derived from a
+          few words chosen at random.  For each word, pseudowords of varying
           frequency but with the co-occurrence distribution of that word
@@
 Section~\ref{replacement-procedure} is performed using the
-distribution~(\ref{distro2}) with $n=7$.  For every replacement token
-(e.g. \word{CAT\_i}), additional occurrences of this token are
+distribution~(\ref{distro2}) with $n=7$.  For every replacement pseudoword
+(e.g. \word{CAT\_i}), additional occurrences of this pseudoword are
 interspersed uniformly at random throughout the corpus, such that the
-final frequency of the replacement token is $2/n$ times that of the
+final frequency of the replacement pseudoword is $2/n$ times that of the
 original word \word{cat}.  For example, if the original word \word{cat}
@@
 word \word{cat} is removed from the corpus and replaced with a family of
-tokens \word{CAT\_i}, $1 \leqslant i \leqslant 7$.  These tokens all
+pseudowords \word{CAT\_i}, $1 \leqslant i \leqslant 7$.  These pseudowords all
 have the same frequency, but their co-occurrence distributions, while
 based on that of \word{cat}, have an increasing amount of noise.
-Specifically, the proportion of noise for the $i$th token is
+Specifically, the proportion of noise for the $i$th pseudoword is
 %
@@
 %
-which is evenly distributed.  The first token contains no noise at all,
-while the last token stands for pure noise.  The particular choice of
+which is evenly distributed.  The first pseudoword contains no noise at all,
+while the last pseudoword stands for pure noise.  The particular choice of
 $n$ assures a reasonable coverage of the interval $[0,1]$.  Other
@@
 dogs.  Notice that the word \word{cat} has been replaced entirely in the
-cats section by \word{CAT\_i} and, moreover, that these same tokens
+cats section by \word{CAT\_i} and, moreover, that these same pseudowords
 appear also in the dogs section. These occurrences (and additionally,
@@
           experiment, where the word \word{cat} was chosen for
-          replacement.  The tokens were generated using the distribution
+          replacement.  The pseudowords were generated using the distribution
           (\ref{distro2}) with $n=7$.}
@@
 Figure~\ref{fig:co-occurrence-noise-heatmap} shows the cosine similarity
-of pairs of vectors representing some of the tokens used in this
-experiment.  Remember that the first token ($i=1$) in a family is
+of pairs of vectors representing some of the pseudowords used in this
+experiment.  Remember that the first pseudoword ($i=1$) in a family is
 without noise in its co-occurrence distribution, while the last one
@@
 vectors within a family only moderately deviate from the original
-direction defined by the first token ($i=1$) when noise is added to the
+direction defined by the first pseudoword ($i=1$) when noise is added to the
 co-occurrence distribution.  For $1<i<7$, the deviation typically
-increases with the proportion of noise.  The vector of the last token
+increases with the proportion of noise.  The vector of the last pseudoword
 ($i=n$), associated with pure noise, is seen within each of the families
@@
 Figure~\ref{fig:co-occurrence-noise-heatmap} as the frequency of the experiment
-tokens is insufficient to sample the full variety of the contexts of VOID, i.e.
-all contexts (see Section \ref{WFVE-direction}).
+pseudowords is insufficient to sample the full variety of the contexts of
+\word{VOID}, i.e.  all contexts (see Section \ref{WFVE-direction}).
 
@@
 	\caption{ Heatmap of the cosine similarity of the vectors
-          representing some of the tokens used in the co-occurrence
+          representing some of the pseudowords used in the co-occurrence
           noise experiment (the words were chosen at random).  The
@@
           the vectors only moderately changes when noise is added to the
-          co-occurrence distribution.  The vector of the tokens
+          co-occurrence distribution.  The vector of the pseudowords
           associated with pure noise ($i=7$) is seen to be almost
@@
 	\caption{ Vector length vs.\ proportion of occurrences from the noise distribution
-          for words chosen for this experiment.  For each word, tokens
+          for words chosen for this experiment.  For each word, pseudowords
           of equal frequency but with increasing proportion of
@@
 More elaborate experiments could be carried out.  For instance, by
-introducing tokens into the corpus that mix, with varying proportions,
+introducing pseudowords into the corpus that mix, with varying proportions,
 the co-occurrence distributions of two words, the path between the word

### 2015-10-08 article_generate_cosine_similarity.py
misc small changes

@@ -9,7 +9,13 @@ from functions import *
 vectors_syn0_filename = sys.argv[1]
 word = sys.argv[2] # e.g. 'the'
 
-vectors, norms = row_normalise_dataframe(load_word2vec_binary(vectors_syn0_filename))
+def row_normalise_dataframe(df):
+    matrix = df.as_matrix() * 1.
+    norms = np.sqrt((matrix ** 2).sum(axis=1))
+    normed_mat = matrix / norms[:, np.newaxis]
+    return pd.DataFrame(normed_mat, index=df.index, columns=df.columns)
+
+vectors = row_normalise_dataframe(load_word2vec_binary(vectors_syn0_filename))
 
 # e.g. 'THE_1' ...
 tokens = [build_experiment_token(word, i) for i in range(1, max(word_freq_experiment_ratio, word_freq_experiment_power_max) + 1)]

### 2015-10-08 functions.py

@@ -1,6 +1,7 @@
 from __future__ import print_function
 import random
 import numpy as np
+from scipy.spatial.distance import cdist
 import pandas as pd
 import matplotlib.pyplot as plt
 from collections import defaultdict
@@ -174,3 +175,16 @@ def cosine_similarity_heatmap(test_vecs, ticks, **kwargs):
     _ = plt.yticks(np.arange(0.5, len(test_vecs.index), 1), ticks, fontsize=11)
     _ = plt.xticks(np.arange(0.5, len(test_vecs.index), 1), ticks, rotation=90, fontsize=11)
     plt.tight_layout()
+
+def by_distance_from(table, v, **params):
+    """
+    Return a Series, listing the distance of each row
+    from the vector given.
+    To specify different metrics, see docstring of scipy.spatial.distance.cdist
+    (default if Euclidean).
+    """
+    v = np.array(v).reshape(1, -1)
+    dist = pd.Series(
+        cdist(v, table, **params)[0], index=table.index, copy=True)
+    dist.sort()
+    return dist

### 2015-10-08 choose_experiment_words.py
small fixes to code

@@ -8,7 +8,7 @@ from parameters import *
 from functions import *
 
 random.seed(sys.argv[1])
-counts = read_word_counts(sys.stdin)
+counts = read_word_counts(sys.argv[2])
 
 frequent_words = [word for word in counts if counts[word] > experiment_word_occurrence_min]
 experiment_words = random.sample(frequent_words, number_of_experiment_words)

### 2015-10-08 functions.py

@@ -18,24 +18,26 @@ def count_words(f):
                 counts[word] += 1
     return counts
 
-def read_words(f):
+def read_words(filename):
     """
     given a CSV of word counts, returns the list of words.
     """
     words = []
-    for line in f:
-        word = line.split(',')[0]
-        words.append(word)
+    with file(filename) as f:
+        for line in f:
+            word = line.split(',')[0]
+            words.append(word)
     return words
 
-def read_word_counts(f):
+def read_word_counts(filename):
     """
     given a CSV of word counts, returns a word count dictionary.
     """
     counts = dict()
-    for line in f:
-        word, count = line.strip().split(',')
-        counts[word] = int(count)
+    with file(filename) as f:
+        for line in f:
+            word, count = line.strip().split(',')
+            counts[word] = int(count)
     return counts
 
 def build_experiment_token(word, sample):

### 2015-10-09 article/main.tex
add reference to wsd pseudoword

@@
         the co-occurrence distribution of the word.  The coefficients of
-        linearity depend upon the word.  The special vector in feature
+        linearity depend upon the word.  The special point in feature
         space, defined by the (artificial) word with pure noise in its
@@
 propose to modify the training corpus in a controlled fashion.  To
-this end, we insert new pseudowords into the corpus with varying frequencies
-and varying levels of noise in their co-occurrence distributions.  By
-modeling the frequency and co-occurrence distributions of these pseudowords
-on existing words in the corpus, we are able to study their effect on
-word vectors independently of one another.  We can thus study a family
-of pseudowords that all appear in the same context, but with different
-frequencies, or study a family of pseudowords that all have the same
-frequency, but appear in a different number of contexts.  Starting from
-the limited number of contexts in which a word appears in the original
-corpus, we can increase this number by interspersing the word in arbitrary
-contexts at random.  The word thus looses its significance in a
-controlled way.  Although we present our approach using the word2vec
-CBOW model, these and related experiments could equally well be carried
-out for other word embedding methods such as the word2vec skip-gram
-model \cite{DistRepns,EfficientEstimation}, GloVe
-\cite{pennington2014glove}, and SENNA \cite{collobert-2011}.
+this end, we insert new tokens into the corpus with varying frequencies and
+varying levels of noise in their co-occurrence distributions.  By modeling the
+frequency and co-occurrence distributions of these tokens, or 
+\textit{pseudowords}\footnote{We refer to these tokens as
+pseudowords, since their properties are modeled upon words in the
+lexicon and because our corpus modification approach is reminiscent of the
+pseudoword approach for generating labeled data for word sense disambiguation
+tasks in \cite{gale1992work}.}, on existing words in the corpus, we are able to
+study their effect on word vectors independently of one another.  We can thus
+study a family of pseudowords that all appear in the same context, but with
+different frequencies, or study a family of pseudowords that all have the same
+frequency, but appear in a different number of contexts.  Starting from the
+limited number of contexts in which a word appears in the original corpus, we
+can increase this number by interspersing the word in arbitrary contexts at
+random.  The word thus looses its significance in a controlled way.  Although
+we present our approach using the word2vec CBOW model, these and related
+experiments could equally well be carried out for other word embedding methods
+such as the word2vec skip-gram model \cite{DistRepns,EfficientEstimation},
+GloVe \cite{pennington2014glove}, and SENNA \cite{collobert-2011}.
 
@@
 perpendicular to it that represents a word with pure noise in its
-co-occurrence distribution.  Surprisingly, this special vector in the feature
-space, obtained by interspersing a pseudoword uniformly at random throughout
-the corpus with a frequency sufficiently large to sample all contexts, is
-non-zero.
+co-occurrence distribution.  Surprisingly, the special point in feature
+space, obtained by interspersing a pseudoword uniformly at random
+throughout the corpus with a frequency sufficiently large to sample all
+contexts, is non-zero.
 
@@
 \begin{section}{Related work}\label{related-work}
-In recent work on the evaluation of word embeddings, Schnabel, Labutov, Mimno
-and Joachims \cite{schnabelemnlp2015} trained logistic regression models to
-predict whether a word was rare or frequent given only the direction of its
-word vector.  The prediction accuracy was then measured as a function of word
-embedding method and of the threshold for word rarity.  It was found in the
-case of word2vec CBOW that word vector direction could be used to distinguish
-very rare words from all other words. The authors claimed further that word
-vector direction must encode frequency directly, and not indirectly via
-semantic information.  Our experimental finding that word vector direction is
-independent of word frequency contradicts this claim in the case of word2vec
-CBOW.
+In recent work on the evaluation of word embeddings, Schnabel et
+al.~\cite{schnabelemnlp2015} trained logistic regression models to
+predict whether a word was rare or frequent given only the direction of
+its word vector.  For various word embedding methods, the prediction
+accuracy was measured as a function of the threshold for word rarity.
+It was found in the case of word2vec CBOW that word vector direction
+could be used to distinguish very rare words from all other words. The
+authors claimed further that word vector direction must encode frequency
+directly, and not indirectly via semantic information.  Our experimental
+finding that word vector direction is independent of word frequency
+contradicts this claim in the case of word2vec CBOW.
 
@@
           \word{cat} is replaced with pseudowords \word{CAT\_i} using the
-          truncated geometric distribution (\ref{distro}) with
+          truncated geometric distribution~(\ref{distro}) with
           $p=\tfrac{1}{2}$ and $n=20$.}
@@
 
-\begin{subsection}{Tokens derived from an artificial, meaningless word}
+\begin{subsection}{Pseudowords derived from an artificial, meaningless word}
 \label{WFVEmeaningless}
@@
           experiment, where the word \word{cat} was chosen for
-          replacement.  The pseudowords were generated using the distribution
-          (\ref{distro2}) with $n=7$.}
+          replacement.  The pseudowords were generated using the
+          distribution~(\ref{distro2}) with $n=7$.}
 	\label{fig:co-occurrence-noise-experiment-text}
@@
 of pairs of vectors representing some of the pseudowords used in this
-experiment.  Remember that the first pseudoword ($i=1$) in a family is
+experiment.  Remember that the first pseudoword~($i=1$) in a family is
 without noise in its co-occurrence distribution, while the last one
@@
 This convergence to a common point is only indistinctly apparent in
-Figure~\ref{fig:co-occurrence-noise-heatmap} as the frequency of the experiment
+Figure~\ref{fig:co-occurrence-noise-heatmap}, as the frequency of the experiment
 pseudowords is insufficient to sample the full variety of the contexts of
-\word{VOID}, i.e.  all contexts (see Section \ref{WFVE-direction}).
+\word{VOID}, i.e., all contexts (see Section~\ref{WFVE-direction}).
 
@@
 transformed feature space.
-
-%\textcolor{red}{can it be argued that the feature space must be fundamentally Euclidean, given that the path marked by the vectorisations of intermediate words is a straight line?
-i.e., if such paths maybe considered as ``geodesics'' between words, and we show that these geodesics appear as straight lines when an Euclidean metric is imposed, then suggests that the Euclidean metric is in fact a good choice.
-what would happen if we chose another metric?}
-
 \end{section}

### 2015-10-09 article_generate_cosine_similarity.py
token -> pseudoword in cosine_similarity for the_i

@@ -25,7 +25,7 @@ non_experiment_idxs = [idx for idx in vectors.index if idx != idx.upper()] # all
 non_experiment_vectors = vectors.loc[non_experiment_idxs]
 
 print r'\begin{tabular}{l | c | l}'
-print r'token & similarity to \word{%s} & most similar words in unmodified corpus\\' % tokens[0].replace('_', '\_')
+print r'pseudoword & similarity to \word{%s} & most similar words in unmodified corpus\\' % tokens[0].replace('_', '\_')
 print '\hline'
 for token in tokens:
     cs = vectors.loc[token].dot(vectors.loc[tokens[0]])

### 2015-11-26 article/main.tex
fixes #8 (turn off subsampling)

@@
 
-Notice the inflection of the syn1neg curve in
-Figure~\ref{fig:word-frequency-experiment-graph} for \word{the} at
-approximately $10^6$.  We have no explanation of this.  Also observe
-that the lengths of the vectors representing the meaningless pseudowords
+Observe that the lengths of the vectors representing the meaningless pseudowords
 \word{VOID\_i} are approximately constant (about $2.5$).  Since we

### 2015-11-26 build_images_word_frequency.py
revise the principle for excluding WFVE vectors (CS with neighbour)

@@ -63,12 +63,16 @@ cosine_similarity_heatmap(example_vecs, ticks, figsize=(12, 10))
 plt.savefig('outputs/word-frequency-experiment-heatmap.eps')
 
 # for the experiment words, drop those word vectors that are not well trained
-# specifically, those whose cosine similarity with word_1 is < 0.8
+# specifically, drop word_i+1, word_i+2, .. if the cosine similarity of word_i and word_i+1 is less than 0.3
 for word in word_freq_experiment_words + [meaningless_token]:
     idxs = wf_experiment_tokens(word) # ordered: word_1, word_2 ..
     example_vecs = vectors_syn0.loc[idxs]
     cs = cosine_similarity(example_vecs)
-    poorly_trained = [idx for i, idx in enumerate(idxs) if cs[0,i] < 0.8]
+    poorly_trained = []
+    for i in range(1, len(example_vecs.index)):
+        if cs[i-1, i] < 0.3:
+            poorly_trained = idxs[i:]
+            break
     stats.drop(poorly_trained, axis=0, inplace=True)
     vectors_syn0.drop(poorly_trained, axis=0, inplace=True)
     vectors_syn1neg.drop(poorly_trained, axis=0, inplace=True)

### 2015-12-02 article/main.tex
article changes prior to AS rereading, updating after Schnabel disc

@@
 word vector length decreases as the level of noise in the co-occurrence
-distribution of the word is increased.  We show furthermore that the
-direction of the word vectors is independent of word frequency and only
-moderately depends upon co-occurrence noise.  When noise is added to the
+distribution of the word is increased.  In addition, we show that the direction
+of a word vector varies smoothly with word frequency and the level of
+co-occurrence noise.  When noise is added to the
 co-occurrence distribution of a word, the corresponding vector smoothly
@@
 \begin{section}{Related work}\label{related-work}
-In recent work on the evaluation of word embeddings, Schnabel et
-al.~\cite{schnabelemnlp2015} trained logistic regression models to
-predict whether a word was rare or frequent given only the direction of
-its word vector.  For various word embedding methods, the prediction
-accuracy was measured as a function of the threshold for word rarity.
-It was found in the case of word2vec CBOW that word vector direction
-could be used to distinguish very rare words from all other words. The
-authors claimed further that word vector direction must encode frequency
-directly, and not indirectly via semantic information.  Our experimental
-finding that word vector direction is independent of word frequency
-contradicts this claim in the case of word2vec CBOW.
-
 Our experimental finding that word vector length decreases with co-occurrence
@@
 
-Consider first the words other than \word{the} and \word{VOID}.  These
-words typically occur in a limited number of distinct contexts, and so a
-limited number of samples is required to capture their co-occurrence
-distributions.  We therefore expect that the vectors representing the
-pseudowords derived from any of these words can be properly learned from the
-modified corpus, even though their combined frequency equals the
-frequency of that word in the original corpus.  The red blocks in
-Figure~\ref{word-frequency-experiment-heatmap} clearly demonstrate that
-the direction of these vectors is constant for each family of pseudowords.
-In other words, the direction of these vectors is only determined by the
-original word and is independent of how often that word appears,
-provided sufficient samples have been processed for the word vector to
-settle in its equilibrium direction.
+Consider first the word vectors for the pseudowords associated to the word
+\word{the}.  Notice that the cosine similarity of the word vectors for
+\word{THE\_1} and \word{THE\_i} decreases monotonically with $i$, while the
+cosine similarity of the word vectors for \word{THE\_i} and \word{THE\_18}
+increases monotonically with $i$.  Indeed the direction of
+the word vector \word{THE\_i} changes systematically, interpolating between the
+directions of the word vectors of the highest-frequency pseudoword
+\word{THE\_1} and the lowest-frequency pseudoword \word{THE\_18}.  The same
+trend is apparent (though over shorter frequency ranges) for all the families of pseudowords
+other than that for \word{VOID}.
 
-The artificial word \word{VOID}, on the other hand, occurs by
-construction in every context.  A much higher number of samples is
-therefore required to capture its co-occurrence distribution, and
-thereby to learn its word vector.  The same is true, but to a lesser
-extent, for the stopword \word{the}.  The difficulty of learning the
-vector representing \word{VOID\_i} or \word{THE\_i} increases with $i$,
-for the pseudowords occur increasingly less often.  This is consistent with
-Figure~\ref{word-frequency-experiment-heatmap}, where it is apparent
-that the vectors representing \word{VOID\_i} for small $i$, say $i = 1,
-\dots, 5$, roughly share a common direction, while this is no longer
-true for higher values of $i$.  The same trend is apparent, although
-less extremely so, for the word \word{the}.  The vectors for the pseudowords
-\word{THE\_i} share a common direction for values of $i$ as high as
-$14$, before they start changing direction.  We conclude that these
-deviations are artifacts of poor learning due to an insufficient number
-of occurrences of these pseudowords.
-Table~\ref{word-frequency-experiment-cosine-similarity}, showing the
-most similar words to each pseudoword \word{THE\_i}, supports this further.
-In the discussions below, the pseudowords \word{VOID\_i} and \word{THE\_i}
-whose vectors have a cosine similarity less than $0.8$ with
-\word{VOID\_1} and \word{THE\_1}, respectively, are therefore discarded.
+Consider now the word vectors for pseudowords derived from the meaningless word
+\word{VOID}.  The word vectors for \word{VOID\_7}, \dots, \word{VOID\_13}
+are approximately orthogonal to one another, just as would be expected from
+randomly drawn vectors in a high dimensional space.
+As the pseudoword \word{VOID} occurs by
+construction in every context, a much higher number of samples is
+required to capture its co-occurrence distribution, and
+thereby to learn its word vector (the same is true, but to a lesser
+extent, for the stopword \word{the}).
+We conclude that the vectors corresponding to the lower frequency
+pseudowords \word{VOID\_7}, \dots, \word{VOID\_13} have not been trained on a
+sufficient number of samples to establish their proper direction.
+These vectors are excluded from further analysis.
+The vectors for \word{VOID\_1}, \dots, \word{VOID\_6}, on the other hand,
+exhibit the smooth change in word vector direction with word frequency described in
+the previous paragraph.
 
-From these results, we conclude that word frequency has no effect on the
-direction of a word vector, given a sufficient number of samples to
-learn the co-occurrence distribution of the word.
+In recent work on the evaluation of word embeddings, Schnabel et
+al.~\cite{schnabelemnlp2015} trained logistic regression models to
+predict whether a word was rare or frequent given only the direction of
+its word vector.  For various word embedding methods, the prediction
+accuracy was measured as a function of the threshold for word rarity.
+It was found in the case of word2vec CBOW that word vector direction
+could be used to distinguish very rare words from all other words.
+Figure~\ref{word-frequency-experiment-heatmap} is consistent with this finding,
+as it is apparent that word vector direction does change gradually with
+frequency.  Schnabel et al.\ claim further that word vector direction must
+encode word frequency directly, and not indirectly via semantic information.
+Figure~\ref{word-frequency-experiment-heatmap}, considered for any particular
+experiment word in isolation (e.g. \word{SQUAD}), demonstrates that the variance of word
+vector direction with word frequency is indeed independent
+of co-occurrence (semantic) information, and thereby provides further evidence
+for this claim.
 
@@
 \end{figure}
-
-\begin{table}
-	\input{cosine-similarity.tex}
-	\caption{ Words in the original vocabulary most similar to the
-          pseudowords \word{THE\_i}, and their cosine similarity with the
-          most frequent such pseudoword, \word{THE\_1}.  It is apparent from
-          the nearest neighbour list that the vectors of the
-          low-frequency pseudowords have not been adequately trained.  }
-	\label{word-frequency-experiment-cosine-similarity}
-\end{table}
 
@@
 
+\begin{section}{Acknowledgments}
+The authors thank Tobias Schnabel for helpful discussions.
+\end{section}
+
 \clearpage

### 2015-12-14 article/main.tex
incorporate adriaan corrections to second version

@@
 
-Consider first the word vectors for the pseudowords associated to the word
-\word{the}.  Notice that the cosine similarity of the word vectors for
+Consider first the vectors for the pseudowords associated to the word
+\word{the}.  Notice that the cosine similarity of the vectors for
 \word{THE\_1} and \word{THE\_i} decreases monotonically with $i$, while the
-cosine similarity of the word vectors for \word{THE\_i} and \word{THE\_18}
+cosine similarity of the vectors for \word{THE\_i} and \word{THE\_18}
 increases monotonically with $i$.  Indeed the direction of
-the word vector \word{THE\_i} changes systematically, interpolating between the
-directions of the word vectors of the highest-frequency pseudoword
+the vector \word{THE\_i} changes systematically, interpolating between the
+directions of the vectors of the highest-frequency pseudoword
 \word{THE\_1} and the lowest-frequency pseudoword \word{THE\_18}.  The same
@@
 
-Consider now the word vectors for pseudowords derived from the meaningless word
-\word{VOID}.  The word vectors for \word{VOID\_7}, \dots, \word{VOID\_13}
+Consider now the vectors for pseudowords derived from the meaningless word
+\word{VOID}.  The vectors for \word{VOID\_7}, \dots, \word{VOID\_13}
 are approximately orthogonal to one another, just as would be expected from
@@
 required to capture its co-occurrence distribution, and
-thereby to learn its word vector (the same is true, but to a lesser
+thereby to learn its vector (the same is true, but to a lesser
 extent, for the stopword \word{the}).
@@
 The vectors for \word{VOID\_1}, \dots, \word{VOID\_6}, on the other hand,
-exhibit the smooth change in word vector direction with word frequency described in
+exhibit the smooth change in vector direction with word frequency described in
 the previous paragraph.
